---
title: "Figure Generation Manuscript"
author: "Tine D'hamers"
date: "2025-12-15"
output: html_document
---

This script contains the code for the generation of the final figures in the manuscript. This is based on data/objects that are obtained in other scritps, and specified here in the notes.

```{r LibraryLoading}
library(patchwork)
library(miloR, 
        lib.loc = "/kyukon/data/gent/vo/002/gvo00224/TOBI/Projects/SingleCellPipeline/resultsCeline/seuratintegrate/Rpackages/R_4.2.1-foss-2022a")
library(ggbeeswarm)
library(gtools)
library(GGally)
```

```{r PlotAesthetics}
# Themes
source("../PlotThemes_sc_CVID.R")
colorsStim<-c(colorsStim, c("CD3/CD28" = colorsStim[["CD3/CD28_beads"]]))

options(bitmapType='cairo')
```

```{r ParameterSettings}
# Saving file
FiguresOut_Path<-"/scratch/gent/vo/002/gvo00224/TOBI/Projects/EProfile/Results/Exp_202211_HCSamples_CEV017-22/ManuscriptFigures"
##### PARAMETERS ###
FDR_thres<-0.1
```

# Figure 2: Clustering, neighbourhoods and annotation revision

```{r Figure2_UMAPS_ClusterDecomposition, fig.width=7, fig.height=7}
#### UMAP - 8H - COLOURED BASED ON CLUSTERS ####
plt2A<-DimPlot(object = IntData.B_8h.FilCus,
        reduction = "B_filtered_CCAInt.umap.rna",
        group.by = "B_filtered_CCAInt_clusters",
        raster = F, label = T, label.size = 7,
        pt.size = 0.5)+
    themeG+
    labs(x = "", y = "",
         title = "")+
    theme(panel.border = element_rect(colour = "grey60", fill=NA, linewidth=0.5),
          axis.ticks = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank())+
  NoLegend()+
  coord_fixed()+
  theme(plot.margin = unit(c(0.5,0.5,0.5,0.5), "cm"))+
  # Make sure squares are obtained
  theme(aspect.ratio=1)

#### UMAP - 30H - COLOURED BASED ON CLUSTERS ####
plt2B<-DimPlot(object = IntData.B_30h.FilCus,
        reduction = "B_filtered_CCAInt.umap.rna",
        group.by = "B_filtered_CCAInt_clusters",
        raster = F, label = T, label.size = 7,
        pt.size = 0.5)+
    themeG+
    labs(x = "", y = "",
         title = "")+
    theme(panel.border = element_rect(colour = "grey60", fill=NA, linewidth=0.5),
          axis.ticks = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank())+
  NoLegend()+
  coord_fixed()+
  theme(plot.margin = unit(c(0.5,0.5,0.5,0.5), "cm"))+
  # Make sure squares are obtained
  theme(aspect.ratio=1)

#### UMAP - 8H - COLOURED BASED ON STIMULATION ####
plt2C<-DimPlot(object = IntData.B_8h.FilCus,
        reduction = "B_filtered_CCAInt.umap.rna",
        group.by = "Stimulation",
        raster = F, label = F,
        pt.size = 0.5)+
    themeG+
    labs(x = "", y = "",
         title = "")+
    theme(panel.border = element_rect(colour = "grey60", fill=NA, linewidth=0.5),
          axis.ticks = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank())+
  scale_color_manual(values = colorsStim)+
  NoLegend()+
  coord_fixed()+
  theme(plot.margin = unit(c(0.5,0.5,0.5,0.5), "cm"))+
  # Make sure squares are obtained
  theme(aspect.ratio=1)

#### UMAP - 30H - COLOURED BASED ON CLUSTERS ####
plt2D<-DimPlot(object = IntData.B_30h.FilCus,
        reduction = "B_filtered_CCAInt.umap.rna",
        group.by = "Stimulation",
        raster = F, label = F,
        pt.size = 0.5)+
    themeG+
    labs(x = "", y = "",
         title = "")+
    theme(panel.border = element_rect(colour = "grey60", fill=NA, linewidth=0.5),
          axis.ticks = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank())+
  scale_color_manual(values = colorsStim)+
  NoLegend()+
  coord_fixed()+
  theme(plot.margin = unit(c(0.5,0.5,0.5,0.5), "cm"))+
  # Make sure squares are obtained
  theme(aspect.ratio=1)


(plt2A + plt2C) / (plt2B + plt2D)

ggsave(filename = file.path(FiguresOut_Path, "Fig2_UMAP_Clusters_Stimulations.pdf"),
       device = "pdf", dpi = 330,
       width = 10, height = 10)
ggsave(filename = file.path(FiguresOut_Path, "Fig2_UMAP_Clusters_Stimulations.png"),
       device = "png", dpi = 330,
       width = 10, height = 10)
```

```{r Figure2_Barplots_Clusters, fig.height=7, fig.width=7}
#### BARPLOT - 8H - PER CLUSTER ####
# Extract data for barplots
plt2E.data <- table(IntData.B_8h.FilClus$B_filtered_CCAInt_clusters, 
                    IntData.B_8h.FilClus$Stimulation)%>%
  as.data.frame()
names(plt2E.data) <- c("Cluster", "Stimulation", "Cells")

# Calculate total number of cells per cluster 
# Calculate the fraction per cluster with regard to the stimulation
plt2E.data<-plt2E.data%>%
  group_by(Cluster)%>%
  mutate(TotalCells = sum(Cells),
         Fraction = Cells/TotalCells*100,
         Cluster = factor(Cluster,
                          levels = levels(Cluster)%>%as.numeric()%>%sort(decreasing = T),
                          ordered = T),
         Stimulation = factor(Stimulation,
                              levels = c("ODN2006", "IFNgamma", "MegaCD40L", "CD3/CD28_beads", "Unstimulated"),
                              ordered = T))
# Plot the data using a barplot
plt2E<-ggplot(plt2E.data)+
  geom_col(mapping = aes(x = Fraction, y = Cluster,
                     fill = Stimulation))+
  # Add the total number of cells on the end of the bar
  geom_text(data = plt2E.data%>%
              group_by(Cluster)%>%
              # Only one number present for total cells for each of the clusters separately
              summarise(TotalCells = max(TotalCells)),
            angle = 0,
            x = 102, vjust = 0.5, hjust = 0,
            size = 4,
            aes(y = Cluster, label = TotalCells))+
  scale_fill_manual(values = colorsStim)+
  themeG+
  theme(axis.text = element_text(size = 10),
        axis.title = element_text(size = 15),
        panel.grid.major.x = element_line(colour = "lightgrey", 
                                          size = 0.5),
        axis.line = element_blank(),
        axis.ticks.y = element_blank())+
  xlim(0,110)+
  scale_x_continuous(breaks=seq(0, 110, 20),
                     limits = c(0,120),
                     expand = c(0,0))+
  labs(x="")

#### BARPLOT - 30H - PER CLUSTER ####
# Extract data for barplots
plt2F.data <- table(IntData.B_30h.FilClus$B_filtered_CCAInt_clusters, 
                    IntData.B_30h.FilClus$Stimulation)%>%
  as.data.frame()
names(plt2F.data) <- c("Cluster", "Stimulation", "Cells")

# Calculate total number of cells per cluster 
# Calculate the fraction per cluster with regard to the stimulation
plt2F.data<-plt2F.data%>%
  group_by(Cluster)%>%
  mutate(TotalCells = sum(Cells),
         Fraction = Cells/TotalCells*100,
         Cluster = factor(Cluster,
                          levels = levels(Cluster)%>%as.numeric()%>%sort(decreasing = T),
                          ordered = T),
         Stimulation = factor(Stimulation,
                              levels = c("ODN2006", "IFNgamma", "MegaCD40L", "CD3/CD28_beads", "Unstimulated"),
                              ordered = T))
# Plot the data using a barplot
plt2F<-ggplot(plt2F.data)+
  geom_col(mapping = aes(x = Fraction, y = Cluster,
                     fill = Stimulation))+
  # Add the total number of cells on the end of the bar
  geom_text(data = plt2F.data%>%
              group_by(Cluster)%>%
              # Only one number present for total cells for each of the clusters separately
              summarise(TotalCells = max(TotalCells)),
            angle = 0,
            x = 102, vjust = 0.5, hjust = 0,
            size = 4,
            aes(y = Cluster, label = TotalCells))+
  scale_fill_manual(values = colorsStim)+
  themeG+
  theme(axis.text = element_text(size = 10),
        axis.title = element_text(size = 15),
        panel.grid.major.x = element_line(colour = "lightgrey", 
                                          size = 0.5),
        axis.line = element_blank(),
        axis.ticks.y = element_blank())+
  xlim(0,110)+
  scale_x_continuous(breaks=seq(0, 110, 20),
                     limits = c(0,120),
                     expand = c(0,0))

plt2E / plt2F & NoLegend()

ggsave(filename = file.path(FiguresOut_Path, "Fig2_Barplots_Clusters_Stimulations.pdf"),
       device = "pdf", dpi = 330,
       width = 5, height = 10)
ggsave(filename = file.path(FiguresOut_Path, "Fig2_Barplots_Clusters_Stimulations.png"),
       device = "png", dpi = 330,
       width = 5, height = 10)

as_ggplot(get_legend(p = plt2E))
ggsave(filename = file.path(FiguresOut_Path, "Fig2_Legend_Stimulations.pdf"),
       device = "pdf", dpi = 330,
       width = 3, height=3)
```

The MiloR object necessary to visualize the plots below are generated in the `013_MiloR_DA_FilteredBCells.Rmd` script.

```{r Figure2_MiloAbundancyPlots_8h_Supplementary, fig.height=3, fig.width=10}
pIFN_8h<-plotNhoodGraphDA(x = milo.B_8h.FilClus$milo_graph, alpha = 0.1,
                   milo_res = DA.B_8h.FilClus%>%filter(Stimulation == "IFNgamma"), layout = "B_filtered_CCAInt.umap.rna"%>%toupper(),
                   size_range = c(0.01,1.5), node_stroke = 0.2)+
    labs(subtitle = paste0("IFN","\U03B3","/R-848"))+
    guides(size=guide_legend("Size of Nhoods"), 
           edge_width = guide_legend("Overlap Nhoods"), 
           fill = guide_colourbar("LogFC", theme = theme(legend.key.width  = unit(0.5, "lines"))))+
  coord_fixed()+
  theme(plot.margin = unit(c(0.5,0.5,0.5,0.5), "cm"),
        panel.border = element_rect(colour = "grey60", fill=NA, linewidth=0.5),
        aspect.ratio = 1)
  

pCD40L_8h<-plotNhoodGraphDA(x = milo.B_8h.FilClus$milo_graph, alpha = 0.1,
                   milo_res = DA.B_8h.FilClus%>%filter(Stimulation == "MegaCD40L"), layout = "B_filtered_CCAInt.umap.rna"%>%toupper(),
                   size_range = c(0.01,1.5), node_stroke = 0.2)+
    labs(subtitle = "MegaCD40L")+
    guides(size=guide_legend("Size of Nhoods"), 
           edge_width = guide_legend("Overlap Nhoods"), 
           fill = guide_colourbar("LogFC", theme = theme(legend.key.width  = unit(0.5, "lines"))))+
  coord_fixed()+
  theme(plot.margin = unit(c(0.5,0.5,0.5,0.5), "cm"),
        panel.border = element_rect(colour = "grey60", fill=NA, linewidth=0.5),
        aspect.ratio = 1)

pCD3CD28_8h<-plotNhoodGraphDA(x = milo.B_8h.FilClus$milo_graph, alpha = 0.1,
                   milo_res = DA.B_8h.FilClus%>%filter(Stimulation == "antiCD3CD28"), layout = "B_filtered_CCAInt.umap.rna"%>%toupper(),
                   size_range = c(0.01,1.5), node_stroke = 0.2)+
    labs(subtitle = "Anti-CD3/CD28")+
    guides(size=guide_legend("Size of Nhoods"), 
           edge_width = guide_legend("Overlap Nhoods"), 
           fill = guide_colourbar("LogFC", theme = theme(legend.key.width  = unit(0.5, "lines"))))+
  coord_fixed()+
  theme(plot.margin = unit(c(0.5,0.5,0.5,0.5), "cm"),
        panel.border = element_rect(colour = "grey60", fill=NA, linewidth=0.5),
        aspect.ratio = 1)

pODN_8h<-plotNhoodGraphDA(x = milo.B_8h.FilClus$milo_graph, alpha = 0.1,
                   milo_res = DA.B_8h.FilClus%>%filter(Stimulation == "ODN2006"), layout = "B_filtered_CCAInt.umap.rna"%>%toupper(),
                   size_range = c(0.01,1.5), node_stroke = 0.2)+
    labs(subtitle = "ODN2006")+
    guides(size=guide_legend("Size of Nhoods"), 
           edge_width = guide_legend("Overlap Nhoods"), 
           fill = guide_colourbar("LogFC", theme = theme(legend.key.width  = unit(0.5, "lines"))))+
  coord_fixed()+
  theme(plot.margin = unit(c(0.5,0.5,0.5,0.5), "cm"),
        panel.border = element_rect(colour = "grey60", fill=NA, linewidth=0.5),
        aspect.ratio = 1)

(pODN_8h + pIFN_8h + pCD40L_8h + pCD3CD28_8h) + plot_layout(ncol = 4) &
  NoLegend() &
  theme(plot.subtitle = element_text(hjust = 0.5, size = 11)) 

ggsave(filename = file.path(FiguresOut_Path, "Fig2_MiloR_UMAPs.png"),
       device = "png", dpi = 330,
       width = 8, height=3)

# Extract and save the legend
as_ggplot(get_legend(p = pIFN_8h))
ggsave(filename = file.path(FiguresOut_Path, "Fig2_Legend_MiloR.pdf"),
       device = "pdf", dpi = 330,
       width = 2, height=5)
```


```{r Figure2_SelectionOfNhoods_8h}
shapeValues = c("TRUE" = 19, "FALSE" = 1)
sizeValues = c("TRUE" = 1, "FALSE" = 0.5)

# Read in the necessary data
plotBee_ClusAnnot.data<-read.xlsx(xlsxFile = file.path(params$path_Results,
                                                         "MiloR",
                                                         paste(params$date2, "CutOffs_Selection_Nhoods_Cells", "8h.xlsx", sep = "_")),
                                  sheet = "OverviewData")
# Factorize the clusters
plotBee_ClusAnnot.data<-plotBee_ClusAnnot.data%>%
  filter(B_filtered_CCAInt_clusters != "Mixed")%>%
  mutate(B_filtered_CCAInt_clusters = factor(B_filtered_CCAInt_clusters,
                                             levels = unique(B_filtered_CCAInt_clusters)%>%mixedsort(decreasing = T),
                                             ordered = T))
# Plot Beeswarm plots in which specific neighbourhoods are selected
plotDAbeeswarm_SubsetIndication(plotBee_ClusAnnot.data%>%filter(Stimulation == "ODN2006"), 
                                group.by  = "B_filtered_CCAInt_clusters", alpha = FDR_thres, ToInclude = "ToInclude",
                                sizeValues = sizeValues, shapeValues = shapeValues)+
  labs(subtitle = "ODN2006", y = "", x = "Original clusters")+
plotDAbeeswarm_SubsetIndication(plotBee_ClusAnnot.data%>%filter(Stimulation == "IFNgamma"), 
                                group.by  = "B_filtered_CCAInt_clusters", alpha = FDR_thres, ToInclude = "ToInclude",
                                sizeValues = sizeValues, shapeValues = shapeValues)+
  labs(subtitle = bquote("IFN-"*gamma*"/R-848"),
       x = "", y = "")+
plotDAbeeswarm_SubsetIndication(plotBee_ClusAnnot.data%>%filter(Stimulation == "MegaCD40L"), 
                                group.by  = "B_filtered_CCAInt_clusters", alpha = FDR_thres, ToInclude = "ToInclude",
                                sizeValues = sizeValues, shapeValues = shapeValues)+
  labs(subtitle = "MegaCD40L", x =  "Original clusters", y = "LogFC")+
plotDAbeeswarm_SubsetIndication(plotBee_ClusAnnot.data%>%filter(Stimulation == "antiCD3CD28"), 
                                group.by  = "B_filtered_CCAInt_clusters", alpha = FDR_thres, ToInclude = "ToInclude",
                                sizeValues = sizeValues, shapeValues = shapeValues)+
  labs(subtitle = "Anti-CD3/CD28", y = "LogFC", x = "")  & ylim(-10,10)&
  themeG &
  theme(panel.grid.major = element_line(color = adjustcolor( "lightgrey", alpha.f = 0.1),
                                          linewidth = 1),
        axis.line = element_blank(),
        axis.ticks = element_blank(),
        plot.subtitle = element_text(hjust = 0.5, size = 11,
                                     color = "black"),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 12)) &  
  NoLegend()

ggsave(filename = file.path(FiguresOut_Path, "Fig2_SelectedNhoods_8h.png"),
       device = "png", dpi = 330,
       width = 10, height=10)
```

```{r Figure2_RevisedAnnotation_Barplot, fig.width = 3, fig.height=7}
#### STACKED BAR PLOT - EXCLUDING HC ####
plt2I<-ggplot(data = UpdAnnot_Cells_Stim_Unstim.Fractions%>%
         filter(Frac_MetaD != 0)%>%
         group_by(NBH_Annot, MetaD_Annot, TimePoint)%>%
         summarise(TotalCells_NBH = sum(TotalCells_NBH),
                   TotalCells_MetaD = sum(TotalCells_MetaD))%>%
         mutate(Frac_MetaD = TotalCells_NBH/TotalCells_MetaD*100)%>%
         mutate(NBH_Annot = factor(NBH_Annot,
                                   levels = c("Undefined", "Unstimulated", "ODN2006", "IFNgamma", "MegaCD40L", "antiCD3CD28"),
                                   labels = c("Undefined","Unstimulated","Stimulated", "Stimulated", "Stimulated", "Stimulated" ),
                                   ordered = T),
                TimePoint = factor(TimePoint,
                                   levels = c("30h", "8h"),
                                   ordered = T)),
       mapping = aes(x = Frac_MetaD, 
                     y = TimePoint,
                     fill = NBH_Annot))+
  themeG+
  theme(panel.grid.major = element_line(color = "grey95",
                                        size = 0.5,
                                        linetype = "solid"),
        axis.line.x = element_blank(),
        axis.ticks =  element_blank())+
  facet_grid(MetaD_Annot~.,
             switch = "y")+
  geom_col(width = 0.6)+
  theme(panel.grid.major.x = element_line(color = "grey95",
                                        size = 0.5,
                                        linetype = "solid"),
        panel.grid.major.y = element_blank(),
        legend.position = "bottom",
        legend.title = element_text(hjust = 0.5),
        legend.title.position = "top",
        axis.line = element_blank(), 
        strip.text= element_text(size = 11),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 12))+
  labs(y = "Incubation Time",
       x = "Fraction",
       fill = "Revised annotation")+
  scale_fill_manual(values = c("Stimulated" = "#49449E",
                               "Unstimulated" = "#8A2F2C",
                               "Undefined" = "grey80"))+
  scale_x_continuous(limits=c(-10,110), breaks=seq(0,100, by = 20),
                     expand = c(0,0))+
  scale_y_discrete(position = "right")+
  guides(fill=guide_legend(ncol=2))

plt2I

ggsave(plot = plt2I,
       filename = file.path(FiguresOut_Path, "Fig2_RevisedAnnotation.pdf"),
       device = "pdf", dpi = 330,
       width = 5, height=8)
```

# Figure 3: Differential gene expression

For this figure, all datasets are obtained in the `014_DEGenes_SelectedNhoods.Rmd` script.

```{r Figure3_UpsetPlots_8h, fig.width=5, fig.height=10}
#### FIGURE 3A ####
### PLOTTING THE DATA USING UPSET PLOT OF COMPLEX HEATMAP ###
# Generate upset plot
upPlot_8h.Fixed.Meta_Rotate<-UpSet(
   t(UpsetData_SIGN_8h.Fixed.Meta),
   comb_order = order(comb_size(UpsetData_SIGN_8h.Fixed.Meta), decreasing=T),
   # Colors of the dots and bars below
   comb_col = "#595B5C", bg_pt_col = 'white',
   # Add plot title
   column_title = "8h of stimulation",
   column_names_rot = 45,
   # Lay out of the text
   column_title_gp = gpar(fontsize = 15),
   column_names_gp = gpar(fontsize = 12),
   # Update row annotation (entire set of DEgenes)
   top_annotation = HeatmapAnnotation(
              "DE genes" = anno_barplot(set_size(UpsetData_SIGN_8h.Fixed.Meta),
                                        border = FALSE,
                                        gp = gpar(fill = "#595B5C",
                                                  lty = "blank"), 
                                        add_numbers = T,
                                        numbers_gp = gpar(fontsize = 10),
                                        height = unit(4, "cm"),
                                        width = unit(4, "cm"),
                                        axis_param = list(gp = gpar(fontsize = 10))),
              annotation_name_rot = 90,
              annotation_name_side = "left",
              annotation_name_gp = gpar(fontsize = 12)),
   # Update column annotation on top of the Upset plot
   right_annotation = rowAnnotation(
      # Add barplot based on the direction of the DE genes
      "Intersection size" = anno_barplot(UpsetData_COMB_8h.Fixed.Meta.table,
                                         border = FALSE,
                                         # Store correct colors
                                         gp = gpar(fill = colDirection[colnames(
                                           UpsetData_COMB_8h.Fixed.Meta.table)],
                                                   lty = "blank"),
                                         add_numbers = T,
                                         width = unit(5, "cm"),
                                         height = unit(5,"cm"),
                                         axis_param = list(gp = gpar(fontsize = 10))),
      # Names to the left
      annotation_name_side = "bottom",
      # Rotation of the annotation at the left side
      annotation_name_rot = 0,
      annotation_name_gp = gpar(fontsize = 12))
   )

# Generate legend
lgd = Legend(
  labels    = colnames(UpsetData_COMB_8h.Fixed.Meta.table),
  legend_gp = gpar(fill = colDirection[colnames(UpsetData_COMB_8h.Fixed.Meta.table)]),
  title     = "Direction"
)

# Plot the final upsetplot
pdf(file.path(FiguresOut_Path, "Fig3_UpsetDEGenes_8h.pdf"), 
    width=5, height=6.5)
draw(upPlot_8h.Fixed.Meta_Rotate,
     annotation_legend_list = list(lgd),
     annotation_legend_side = "left")
dev.off()
```

```{r Figure3_UpsetPlots_30h, fig.width=5, fig.height=10}
#### FIGURE 3B ####
### PLOTTING THE DATA USING UPSET PLOT OF COMPLEX HEATMAP ###
# Generate upset plot
upPlot_30h.Fixed.Meta_Rotate<-UpSet(
   t(UpsetData_SIGN_30h.Fixed.Meta),
   comb_order = order(comb_size(UpsetData_SIGN_30h.Fixed.Meta), decreasing=T),
   # Colors of the dots and bars below
   comb_col = "#595B5C", bg_pt_col = 'white',
   # Add plot title
   column_title = "30h of stimulation",
   column_names_rot = 45,
   # Lay out of the text
   column_title_gp = gpar(fontsize = 15),
   column_names_gp = gpar(fontsize = 12),
   # Update row annotation (entire set of DEgenes)
   top_annotation = HeatmapAnnotation(
              "DE genes" = anno_barplot(set_size(UpsetData_SIGN_30h.Fixed.Meta),
                                        border = FALSE,
                                        gp = gpar(fill = "#595B5C",
                                                  lty = "blank"), 
                                        add_numbers = T,
                                        numbers_gp = gpar(fontsize = 10),
                                        height = unit(4, "cm"),
                                        width = unit(4, "cm"),
                                        axis_param = list(gp = gpar(fontsize = 10))),
              annotation_name_rot = 90,
              annotation_name_side = "left",
              annotation_name_gp = gpar(fontsize = 12)),
   # Update column annotation on top of the Upset plot
   right_annotation = rowAnnotation(
      # Add barplot based on the direction of the DE genes
      "Intersection size" = anno_barplot(UpsetData_COMB_30h.Fixed.Meta.table,
                                         border = FALSE,
                                         # Store correct colors
                                         gp = gpar(fill = colDirection[colnames(
                                           UpsetData_COMB_30h.Fixed.Meta.table)],
                                                   lty = "blank"),
                                         add_numbers = T,
                                         width = unit(5, "cm"),
                                         height = unit(5,"cm"),
                                         axis_param = list(gp = gpar(fontsize = 10))),
      # Names to the left
      annotation_name_side = "bottom",
      # Rotation of the annotation at the left side
      annotation_name_rot = 0,
      annotation_name_gp = gpar(fontsize = 12))
   )

# Generate legend
lgd = Legend(
  labels    = colnames(UpsetData_COMB_30h.Fixed.Meta.table),
  legend_gp = gpar(fill = colDirection[colnames(UpsetData_COMB_30h.Fixed.Meta.table)]),
  title     = "Direction"
)

# Plot the final upsetplot
pdf(file.path(FiguresOut_Path, "Fig3_UpsetDEGenes_30h.pdf"), 
    width=5, height=6.5)
draw(upPlot_30h.Fixed.Meta_Rotate,
     annotation_legend_list = list(lgd),
     annotation_legend_side = "left")
dev.off()
```

```{r CorrelationGeneExpression_SelectedGenes, fig.width=9, fig.height=8}
#### FIGURE 3C ####
CorHeatmap <- Heatmap(DEG_Stim_Unstim_AllTimes_FilClus.Fixed.Meta.CorMat.SelGenes,
        name = "Pearson correlation",
        rect_gp = gpar(type = "none"), 
        cluster_rows = FALSE, cluster_columns = FALSE,
        show_row_names = F, show_column_names = F,
        col = colorRamp2(c(0, 0.5, 1), c("white", "#93B5C6", "#044C99")),
        cell_fun = function(j, i, x, y, w, h, fill) {
          if(i<j) {
            grid.rect(x, y, w, h, gp = gpar(fill = fill, col = "white"))
            grid.text(sprintf("%.2f", DEG_Stim_Unstim_AllTimes_FilClus.Fixed.Meta.CorMat.SelGenes[i, j]), 
                      x, y, gp = gpar(fontsize = 15, col = "white"))
            }
          },
        right_annotation = rowAnnotation(set=anno_text(colnames(DEG_Stim_Unstim_AllTimes_FilClus.Fixed.Meta.CorMat.SelGenes),
                                                       gp = gpar(fontsize = 15))),
        top_annotation = HeatmapAnnotation(set=anno_text(colnames(DEG_Stim_Unstim_AllTimes_FilClus.Fixed.Meta.CorMat.SelGenes),
                                                         location = 0, just="left", rot = 45,
                                                         gp = gpar(fontsize = 15))))

CorHeatmap

OverlapHeatmap <- Heatmap(MatOverlap,
        name = "Absolute overlap",
        rect_gp = gpar(type = "none"), 
        row_order = names(DEG_Stim_Unstim_8hvs30_FilClus.Fixed.Meta),
        column_order = names(DEG_Stim_Unstim_8hvs30_FilClus.Fixed.Meta),
        col = colorRamp2(c(0, 300, 1100), c("white", "#AD7E55", "#6C3505")),
        row_names_side = "left", column_names_side = "top",
        cell_fun = function(j, i, x, y, w, h, fill) {
          if(i>=j){
            grid.rect(x, y, w, h, gp = gpar(fill = fill, col = "white"))
            grid.text(round(MatOverlap[i,j], digits = 1), x, y,
                      gp = gpar(fontsize = 15, col = "black"))
            }
          },
        show_row_names = F, show_column_names = F)

OverlapHeatmap

pdf(file.path(FiguresOut_Path, "Fig3_CorHeatmap_OverlapGenes.pdf"), 
    width=9, height=8)
draw(OverlapHeatmap + CorHeatmap, ht_gap = unit(-147, "mm"))
dev.off()
```

```{r Figure3_SelectionTopGenes_UpregulatedGenesOnly, fig.width=4, fig.height=7}
# PlotData
#### ROTATED FIGURE ####
# ggplot(topGenes_8h_30h.Data.Fixed.Meta_UP,
#        mapping = aes(x = reorder(Gene, desc(Gene)),
#                      y = reorder(Stimulation, desc(Stimulation))))+
#   facet_grid(TimePoint~.)+
#   geom_vline(mapping = aes(xintercept = Gene),
#              color = "grey80", 
#              alpha = 0.5)+
#   geom_hline(mapping = aes(yintercept = Stimulation),
#              color = "grey80",
#              alpha = 0.5)+
#   geom_point(mapping = aes(size = Score,
#                            fill = avg_log2FC),
#              colour = "grey50",
#              shape = 21)+
#   geom_point(data = topGenes_8h_30h.Data.Fixed.Meta_UP%>%
#                filter(InTop),
#              shape = 20, color = "black",
#              size = 0.1)+
#   scale_fill_gradient2(low = colDirection["DOWN"], 
#                         mid = "white", 
#                         high = colDirection["UP"],
#                         midpoint = 0)+
#   scale_size(range = c(1,6))+
#   themeG+
#   # scale_y_discrete(position = "bottom")+
#   theme(axis.text.x = element_text(angle = 45,
#                                    vjust = 1, hjust = 1,
#                                    size=10), 
#         axis.text.y = element_text(size = 10),
#         axis.line = element_blank(),
#         axis.ticks = element_blank(),
#         axis.title = element_blank())

ggplot(topGenes_8h_30h.Data.Fixed.Meta_UP,
       mapping = aes(y = Gene,
                     x = Stimulation))+
  facet_grid(.~TimePoint)+
  geom_hline(mapping = aes(yintercept = Gene),
             color = "grey80",
             alpha = 0.5)+
  geom_vline(mapping = aes(xintercept = Stimulation),
             color = "grey80",
             alpha = 0.5)+
  geom_point(mapping = aes(size = Score,
                           fill = avg_log2FC),
             colour = "grey50",
             shape = 21)+
  geom_point(data = topGenes_8h_30h.Data.Fixed.Meta_UP%>%
               filter(InTop),
             shape = 20, color = "black",
             size = 0.1)+
  scale_fill_gradient2(low = colDirection["DOWN"],
                        mid = "white",
                        high = colDirection["UP"],
                        midpoint = 0)+
  scale_size(range = c(1,6))+
  themeG+
  scale_x_discrete(position = "bottom")+
  theme(axis.text.x = element_text(angle = 45,
                                   vjust = 1, hjust = 1,
                                   size=10),
        axis.text.y = element_text(size = 10),
        axis.line = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank())

ggsave(filename = file.path(FiguresOut_Path, "Fig3_ExpressionLevels_Top5_PerTimeStim.pdf"),
       device = "pdf", dpi = 330,
       width = 5, height=6)
```

```{r Figure3_UpsetPlotGeneSets, fig.width=10, fig.height=4}
#### FIGURE 3D ####
# Only visualize the combinations with at least 5 gene sets in common
UpsetData_SIGN_GSE.Fixed.SelFig<-UpsetData_SIGN_GSE.Fixed[comb_size(UpsetData_SIGN_GSE.Fixed) >= 5]

# Extract the timepoint
colUpset.Time<-set_name(UpsetData_SIGN_GSE.Fixed.SelFig)%>%
  grepl(pattern = "30")
colUpset.Time[colUpset.Time]<-"30"
colUpset.Time[colUpset.Time == "FALSE"]<-"8"
# Extract the expression direction
colUpset.Exp<-set_name(UpsetData_SIGN_GSE.Fixed.SelFig)%>%
  grepl(pattern = "UP")
colUpset.Exp[colUpset.Exp]<-"UP"
colUpset.Exp[colUpset.Exp == "FALSE"]<-"DOWN"

#### ROTATED FIGURE ####
# UpSet(t(UpsetData_SIGN_GSE.Fixed.SelFig), 
#       comb_order = order(comb_size(UpsetData_SIGN_GSE.Fixed.SelFig), decreasing = T),
#       # Lay out of the text
#       column_names_gp = gpar(fontsize = 10),
#       comb_col = "#595B5C", bg_pt_col = "white",
#       top_annotation = HeatmapAnnotation(
#                 "GSE" = anno_barplot(set_size(UpsetData_SIGN_GSE.Fixed.SelFig),
#                                           border = FALSE,
#                                           gp = gpar(fill = "#595B5C",
#                                                     lty = "blank"), 
#                                           add_numbers = T,
#                                           width = unit(1, "cm"),
#                                           height = unit(3, "cm")),
#                 TimePoint=colUpset.Time,
#                 Direction=colUpset.Exp,
#                 col = list("TimePoint"=colorsTime,
#                            "Direction"=colDirection)),
#       right_annotation = rowAnnotation(
#                 "Intersection size" = anno_barplot(comb_size(UpsetData_SIGN_GSE.Fixed.SelFig),
#                                                    border = FALSE,
#                                                    gp = gpar(fill = "#595B5C",
#                                                              lty = "blank"), 
#                                                    add_numbers = T,
#                                                    height = unit(1, "cm"),
#                                                    width = unit(2, "cm")),
#                 annotation_name_side = "bottom",
#                 annotation_name_rot = 90))

pdf(file.path(FiguresOut_Path, "Fig3_UpsetGSEA.pdf"), 
    width=10, height=4)
# Visualize the upsetplot
UpSet(UpsetData_SIGN_GSE.Fixed.SelFig, 
      comb_order = order(comb_size(UpsetData_SIGN_GSE.Fixed.SelFig), decreasing = T),
      # Lay out of the text
      column_names_gp = gpar(fontsize = 10),
      comb_col = "#595B5C", bg_pt_col = "white",
      right_annotation = rowAnnotation(
                "GSE" = anno_barplot(set_size(UpsetData_SIGN_GSE.Fixed.SelFig),
                                          border = FALSE,
                                          gp = gpar(fill = "#595B5C",
                                                    lty = "blank"), 
                                          add_numbers = T,
                                          width = unit(3, "cm"),
                                          height = unit(1, "cm")),
                TimePoint=colUpset.Time,
                Direction=colUpset.Exp,
                col = list("TimePoint"=colorsTime,
                           "Direction"=colDirection)),
      top_annotation = HeatmapAnnotation(
                "Intersection size" = anno_barplot(comb_size(UpsetData_SIGN_GSE.Fixed.SelFig),
                                                   border = FALSE,
                                                   gp = gpar(fill = "#595B5C",
                                                             lty = "blank"), 
                                                   add_numbers = T,
                                                   height = unit(3, "cm"),
                                                   width = unit(1, "cm")),
                annotation_name_side = "left",
                annotation_name_rot = 0))
dev.off()
```
# Supplementary figures

## Correlation in cluster fractions per stimulation

```{r SuplFigure_Barplots_ClustersPerStimulation, fig.height=7, fig.width=7}
#### BARPLOT - 8H - PER CLUSTER ####
# Extract data for barplots
plt2E.data.Rev <- table(IntData.B_8h.FilClus$Stimulation,
                        IntData.B_8h.FilClus$B_filtered_CCAInt_clusters)%>%
  as.data.frame()
names(plt2E.data.Rev) <- c("Stimulation", "Cluster", "Cells")

# Calculate total number of cells per cluster 
# Calculate the fraction per cluster with regard to the stimulation
plt2E.data.Rev<-plt2E.data.Rev%>%
  group_by(Stimulation)%>%
  mutate(TotalCells = sum(Cells),
         Fraction = Cells/TotalCells*100,
         Cluster = factor(Cluster,
                          levels = levels(Cluster)%>%as.numeric()%>%sort(decreasing = T),
                          ordered = T),
         Stimulation = factor(Stimulation,
                              levels = c("ODN2006", "IFNgamma", "MegaCD40L", "CD3/CD28_beads", "Unstimulated"),
                              ordered = T))
# Plot the data using a barplot
plt2E.Rev<-ggplot(plt2E.data.Rev)+
  geom_col(mapping = aes(x = Fraction, y = Stimulation),
           fill = "white", color = "black")+
  # Add the total number of cells on the end of the bar
  geom_text(data = plt2E.data.Rev%>%
              group_by(Stimulation)%>%
              # Only one number present for total cells for each of the clusters separately
              summarise(TotalCells = max(TotalCells)),
            angle = 0,
            x = 102, vjust = 0.5, hjust = 0,
            size = 4,
            aes(y = Stimulation, label = TotalCells))+
  themeG+
  theme(axis.text = element_text(size = 10),
        axis.title = element_text(size = 15),
        panel.grid.major.x = element_line(colour = "lightgrey", 
                                          size = 0.5),
        axis.line = element_blank(),
        axis.ticks.y = element_blank())+
  xlim(0,110)+
  scale_x_continuous(breaks=seq(0, 110, 20),
                     limits = c(0,120),
                     expand = c(0,0))+
  labs(x="", subtitle="8h")

#### BARPLOT - 30H - PER CLUSTER ####
# Extract data for barplots
plt2F.data.Rev <- table(IntData.B_30h.FilClus$Stimulation,
                        IntData.B_30h.FilClus$B_filtered_CCAInt_clusters)%>%
  as.data.frame()
names(plt2F.data.Rev) <- c("Stimulation", "Cluster", "Cells")

# Calculate total number of cells per cluster 
# Calculate the fraction per cluster with regard to the stimulation
plt2F.data.Rev<-plt2F.data.Rev%>%
  group_by(Stimulation)%>%
  mutate(TotalCells = sum(Cells),
         Fraction = Cells/TotalCells*100,
         Cluster = factor(Cluster,
                          levels = levels(Cluster)%>%as.numeric()%>%sort(decreasing = T),
                          ordered = T),
         Stimulation = factor(Stimulation,
                              levels = c("ODN2006", "IFNgamma", "MegaCD40L", "CD3/CD28_beads", "Unstimulated"),
                              ordered = T))
# Plot the data using a barplot
plt2F.Rev<-ggplot(plt2F.data.Rev)+
  geom_col(mapping = aes(x = Fraction, y = Stimulation),
           fill = "white", color = "black")+
  # Add the total number of cells on the end of the bar
  geom_text(data = plt2F.data.Rev%>%
              group_by(Stimulation)%>%
              # Only one number present for total cells for each of the clusters separately
              summarise(TotalCells = max(TotalCells)),
            angle = 0,
            x = 102, vjust = 0.5, hjust = 0,
            size = 4,
            aes(y = Stimulation, label = TotalCells))+
  themeG+
  theme(axis.text = element_text(size = 10),
        axis.title = element_text(size = 15),
        panel.grid.major.x = element_line(colour = "lightgrey", 
                                          size = 0.5),
        axis.line = element_blank(),
        axis.ticks.y = element_blank())+
  xlim(0,110)+
  scale_x_continuous(breaks=seq(0, 110, 20),
                     limits = c(0,120),
                     expand = c(0,0))+
  labs(x="", subtitle = "30h")

plt2E.Rev / plt2F.Rev & NoLegend()


```

```{r CorrelationPlot_Fractions_ClustersPerStimulation, fig.height=10, fig.width=10}
#### SCATTER PLOT + LABELING ####
plot_lower_corrPlot.ggpairs<-function(data, mapping){
  
  # Extract the names of the columns present on the x and y axis
  x_name<-as_label(mapping$x)
  y_name<-as_label(mapping$y)
  
  # Plotting the data using a scatter plot
  ggplot(data, mapping = aes(x = .data[[x_name]],
                            y = .data[[y_name]]))+
    geom_abline(slope = 1, intercept = 0,
              color = "grey50")+
    geom_point(size = 1)+
    theme_minimal()+
    scale_y_continuous(limits=c(0,50), breaks=seq(0, 50, by = 5))+
    scale_x_continuous(limits=c(0,50), breaks=seq(0, 50, by = 5))+
    theme(axis.text.x = element_text(angle=45))
}

#### SCATTERPLOT ####
plot_diag_corrPlot.ggpairs<-function(data, mapping){
  # Extract the names of the columns present on the x and y axis
  x_name<-as_label(mapping$x)
  
  # Plotting the data
  ggplot(data = data, 
         mapping = aes(y = .data[[x_name]], 
                       x = Cluster))+
    geom_col(width = 0.5)+
    theme_minimal()+
    theme(axis.text.x = element_text(angle=45))
}

#### CORRELATION CALCULATION ####
plot_upper_corrPlot.ggpairs <- function(data, mapping){
  
  # Extract the names of the columns present on the x and y axis
  x_name<-as_label(mapping$x)
  y_name<-as_label(mapping$y)

  
  # Store correlation
  tmp_cor<-cor(x = data[[x_name]], y = data[[y_name]],
             method = "spearman")
  
  # Plot the output
  ggplot() + 
    annotate(geom = "text",
             x = 0.5, y = 0.5,
             label = sprintf("Corr: \n %.3f", tmp_cor),
             size = 6)+
    theme_void()
  }

# Prepare plotting data
plt2E.data.Rev.Corr<-plt2E.data.Rev%>%
  select(Stimulation, Fraction, Cluster)%>%
  pivot_wider(names_from = "Stimulation",
              values_from = "Fraction")%>%
  mutate(Cluster = factor(Cluster,
                          levels = mixedsort(unique(Cluster), decreasing = TRUE),
                          ordered = TRUE))%>%
  select(Cluster, ODN2006, IFNgamma, MegaCD40L, `CD3/CD28_beads`, Unstimulated)

plt2F.data.Rev.Corr<-plt2F.data.Rev%>%
  select(Stimulation, Fraction, Cluster)%>%
  pivot_wider(names_from = "Stimulation",
              values_from = "Fraction")%>%
  mutate(Cluster = factor(Cluster,
                          levels = mixedsort(unique(Cluster), decreasing = TRUE),
                          ordered = TRUE))%>%
  select(Cluster, ODN2006, IFNgamma, MegaCD40L, `CD3/CD28_beads`, Unstimulated)
  

ggpairs(plt2E.data.Rev.Corr,
        columns = 2:6,
        lower = list(continuous = GGally::wrap(plot_lower_corrPlot.ggpairs)),
        diag = list(continuous = GGally::wrap(plot_diag_corrPlot.ggpairs)),
        upper = list(continuous = GGally::wrap(plot_upper_corrPlot.ggpairs)),
        title = "8h")+
  theme(axis.text = element_text(size = 10),
        axis.text.x = element_text(angle = 270, vjust = 0.5, hjust = 0),
        strip.background = element_blank(),
        strip.text = element_text(size = 17))

ggsave(filename = file.path(FiguresOut_Path, "SuplFig_ClusterFractions_8h_CorrelationPlot.pdf"),
       device = "pdf", dpi = 330,
       width = 10, height=10, bg = "white")

ggpairs(plt2F.data.Rev.Corr,
        columns = 2:6,
        lower = list(continuous = GGally::wrap(plot_lower_corrPlot.ggpairs)),
        diag = list(continuous = GGally::wrap(plot_diag_corrPlot.ggpairs)),
        upper = list(continuous = GGally::wrap(plot_upper_corrPlot.ggpairs)),
        title = "30h")+
  theme(axis.text = element_text(size = 10),
        axis.text.x = element_text(angle = 270, vjust = 0.5, hjust = 0),
        strip.background = element_blank(),
        strip.text = element_text(size = 17))

ggsave(filename = file.path(FiguresOut_Path, "SuplFig_ClusterFractions_30h_CorrelationPlot.pdf"),
       device = "pdf", dpi = 330,
       width = 10, height=10, bg = "white")
```
## Milo abundancy - selection Nhoods - 30h

```{r Figure2_MiloAbundancyPlots_30h_Supplementary, fig.height=7, fig.width=7}
pIFN_30h<-plotNhoodGraphDA(x = milo.B_30h.FilClus$milo_graph, alpha = 0.1,
                   milo_res = DA.B_30h.FilClus%>%filter(Stimulation == "IFNgamma"), layout = "B_filtered_CCAInt.umap.rna"%>%toupper(),
                   size_range = c(0.01,1.5), node_stroke = 0.2)+
    labs(subtitle = paste0("IFN","\U03B3","/R-848"))+
    guides(size=guide_legend("Size of Nhoods"), 
           edge_width = guide_legend("Overlap Nhoods"), 
           fill = guide_colourbar("LogFC", theme = theme(legend.key.width  = unit(0.5, "lines"))))+
  coord_fixed()+
  theme(plot.margin = unit(c(0.5,0.5,0.5,0.5), "cm"),
        panel.border = element_rect(colour = "grey60", fill=NA, linewidth=0.5),
        aspect.ratio = 1)
  

pCD40L_30h<-plotNhoodGraphDA(x = milo.B_30h.FilClus$milo_graph, alpha = 0.1,
                   milo_res = DA.B_30h.FilClus%>%filter(Stimulation == "MegaCD40L"), layout = "B_filtered_CCAInt.umap.rna"%>%toupper(),
                   size_range = c(0.01,1.5), node_stroke = 0.2)+
    labs(subtitle = "MegaCD40L")+
    guides(size=guide_legend("Size of Nhoods"), 
           edge_width = guide_legend("Overlap Nhoods"), 
           fill = guide_colourbar("LogFC", theme = theme(legend.key.width  = unit(0.5, "lines"))))+
  coord_fixed()+
  theme(plot.margin = unit(c(0.5,0.5,0.5,0.5), "cm"),
        panel.border = element_rect(colour = "grey60", fill=NA, linewidth=0.5),
        aspect.ratio = 1)

pCD3CD28_30h<-plotNhoodGraphDA(x = milo.B_30h.FilClus$milo_graph, alpha = 0.1,
                   milo_res = DA.B_30h.FilClus%>%filter(Stimulation == "antiCD3CD28"), layout = "B_filtered_CCAInt.umap.rna"%>%toupper(),
                   size_range = c(0.01,1.5), node_stroke = 0.2)+
    labs(subtitle = "Anti-CD3/CD28")+
    guides(size=guide_legend("Size of Nhoods"), 
           edge_width = guide_legend("Overlap Nhoods"), 
           fill = guide_colourbar("LogFC", theme = theme(legend.key.width  = unit(0.5, "lines"))))+
  coord_fixed()+
  theme(plot.margin = unit(c(0.5,0.5,0.5,0.5), "cm"),
        panel.border = element_rect(colour = "grey60", fill=NA, linewidth=0.5),
        aspect.ratio = 1)

pODN_30h<-plotNhoodGraphDA(x = milo.B_30h.FilClus$milo_graph, alpha = 0.1,
                   milo_res = DA.B_30h.FilClus%>%filter(Stimulation == "ODN2006"), layout = "B_filtered_CCAInt.umap.rna"%>%toupper(),
                   size_range = c(0.01,1.5), node_stroke = 0.2)+
    labs(subtitle = "ODN2006")+
    guides(size=guide_legend("Size of Nhoods"), 
           edge_width = guide_legend("Overlap Nhoods"), 
           fill = guide_colourbar("LogFC", theme = theme(legend.key.width  = unit(0.5, "lines"))))+
  coord_fixed()+
  theme(plot.margin = unit(c(0.5,0.5,0.5,0.5), "cm"),
        panel.border = element_rect(colour = "grey60", fill=NA, linewidth=0.5),
        aspect.ratio = 1)

(pODN_30h + pIFN_30h + pCD40L_30h + pCD3CD28_30h) + plot_layout(ncol = 4) &
  NoLegend() &
  theme(plot.subtitle = element_text(hjust = 0.5, size = 11)) 

ggsave(filename = file.path(FiguresOut_Path, "SuplFig_MiloR_UMAPs_30h.png"),
       device = "png", dpi = 330,
       width = 8, height=3)

# Extract and save the legend
as_ggplot(get_legend(p = pIFN_30h))
ggsave(filename = file.path(FiguresOut_Path, "SuplFig_Legend_MiloR_30h.pdf"),
       device = "pdf", dpi = 330,
       width = 2, height=5)
```

Included the mixed NBHs as well to show how this looks like. These are excluded in the figure of 8h of stimulation.

```{r Figure2_SelectionOfNhoods_30h_Supplementary}
# Read in the necessary data
plotBee_ClusAnnot.data.30h<-read.xlsx(xlsxFile = file.path(params$path_Results,
                                                       "MiloR",
                                                       paste(params$date2, "CutOffs_Selection_Nhoods_Cells", "30h.xlsx", sep = "_")),
                                  sheet = "OverviewData")
# Factorize the clusters
plotBee_ClusAnnot.data.30h<-plotBee_ClusAnnot.data.30h%>%
  mutate(B_filtered_CCAInt_clusters = factor(B_filtered_CCAInt_clusters,
                                             levels = unique(B_filtered_CCAInt_clusters)%>%mixedsort(decreasing = T),
                                             ordered = T))
# Plot Beeswarm plots in which specific neighbourhoods are selected
plotDAbeeswarm_SubsetIndication(plotBee_ClusAnnot.data.30h%>%filter(Stimulation == "ODN2006"), 
                                group.by  = "B_filtered_CCAInt_clusters", alpha = FDR_thres, ToInclude = "ToInclude",
                                sizeValues = sizeValues, shapeValues = shapeValues)+
  labs(subtitle = "ODN2006", y = "", x = "Original clusters")+
plotDAbeeswarm_SubsetIndication(plotBee_ClusAnnot.data.30h%>%filter(Stimulation == "IFNgamma"), 
                                group.by  = "B_filtered_CCAInt_clusters", alpha = FDR_thres, ToInclude = "ToInclude",
                                sizeValues = sizeValues, shapeValues = shapeValues)+
  labs(subtitle = bquote("IFN-"*gamma*"/R-848"),
       x = "", y = "")+
plotDAbeeswarm_SubsetIndication(plotBee_ClusAnnot.data.30h%>%filter(Stimulation == "MegaCD40L"), 
                                group.by  = "B_filtered_CCAInt_clusters", alpha = FDR_thres, ToInclude = "ToInclude",
                                sizeValues = sizeValues, shapeValues = shapeValues)+
  labs(subtitle = "MegaCD40L", x =  "Original clusters", y = "LogFC")+
plotDAbeeswarm_SubsetIndication(plotBee_ClusAnnot.data.30h%>%filter(Stimulation == "antiCD3CD28"), 
                                group.by  = "B_filtered_CCAInt_clusters", alpha = FDR_thres, ToInclude = "ToInclude",
                                sizeValues = sizeValues, shapeValues = shapeValues)+
  labs(subtitle = "Anti-CD3/CD28", y = "LogFC", x = "")  & ylim(-10,10)&
  themeG &
  theme(panel.grid.major = element_line(color = adjustcolor( "lightgrey", alpha.f = 0.1),
                                          linewidth = 1),
        axis.line = element_blank(),
        axis.ticks = element_blank(),
        plot.subtitle = element_text(hjust = 0.5, size = 11,
                                     color = "black"),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 12)) &  
  NoLegend()

ggsave(filename = file.path(FiguresOut_Path, "SuplFig_SelectedNhoods_30h.png"),
       device = "png", dpi = 330,
       width = 10, height=10)
```

## Revised annotation - HC information included

```{r Figure2_Scatterplot_RevisedAnnotation_HCInformation}
#### SCATTER PLOT ####
ggplot(data = UpdAnnot_Cells_Stim_Unstim.Fractions%>%
         filter(Frac_MetaD != 0)%>%
         mutate(NBH_Annot = factor(NBH_Annot,
                                   levels = c("ODN2006", "IFNgamma", "MegaCD40L", "antiCD3CD28", "Unstimulated", "Undefined"),
                                   labels = c("Stimulated", "Stimulated", "Stimulated", "Stimulated", "Unstimulated", "Undefined"))),
       mapping = aes(y = Frac_MetaD, 
                     x = TimePoint,
                     color = HC))+
  scale_y_continuous(limits=c(-10,110), breaks=seq(0,100, by = 20))+
  theme(panel.grid.major = element_line(color = "grey95",
                                        size = 0.5,
                                        linetype = "solid"),
        axis.line.x = element_blank())+
  stat_summary(mapping = aes(group = TimePoint),
               fun.data = mean_sdl, 
               geom = "errorbar",
               width = 0.1,
               color = "grey60")+
  stat_summary(mapping = aes(group = TimePoint),
               fun = mean,
               geom = "point",
               shape = 4,
               color = "grey60")+
  facet_grid(cols = vars(MetaD_Annot),
             rows = vars(NBH_Annot))+
  geom_point(alpha = 0.6)+
  themeG+
  scale_colour_manual(values = colorsHC)+
  theme(panel.grid.major.y = element_line(color = "grey95",
                                        size = 0.5,
                                        linetype = "solid"),
        axis.line = element_blank(),
        axis.ticks = element_blank())+
  labs(x = "",
       y = "Fraction")+
  scale_y_continuous(expand = c(0,0))

ggsave(filename = file.path(FiguresOut_Path, "SuplFig_RevisedAnnotation_HC.pdf"),
       device = "pdf", dpi = 330,
       width = 8, height=5)
```

## Density plots of revised annotation
```{r DensityPlots_RevisedAnnot, fig.height=17, fig.width=7}
#### 8H - ORIGINAL ANNOTATION ####
Idents(IntData.B_8h.FilClus)<-'Stimulation'
Dens.plotlist <- list()

# Generate the density plots and save them in the list
for (stim in c("ODN2006", "IFNgamma", "MegaCD40L", "CD3/CD28_beads", "Unstimulated")){
  Dens.plotlist[[stim]] <- DimPlot(IntData.B_8h.FilClus, reduction = "B_filtered_CCAInt.umap.rna",
                                 group.by = "Stimulation", cols = rep("lightgrey", times = 5))+
    themeG+
    theme(legend.position = "none")+
    stat_density2d(data = subset(IntData.B_8h.FilClus, idents = stim)@reductions[["B_filtered_CCAInt.umap.rna"]]@cell.embeddings,
                   mapping = aes(x = BfilteredCCAIntumaprna_1,
                                 y = BfilteredCCAIntumaprna_2,
                                 fill = ..level..),
                   geom = "polygon", alpha = 0.6)+
    scale_fill_gradient(low = "lightblue",high = "darkblue", guide = "colourbar",
                        breaks = c(0,0.05, 0.1))+
    labs(title = stim, x = "UMAP1", y = "UMAP2")+
    theme(axis.line = element_blank(),
          axis.ticks = element_blank(),
          axis.text = element_blank(),
          axis.title = element_blank())+
  NoLegend()+
  coord_fixed()+
  ylim(-7,7)+
  xlim(-8,8)+
  theme(plot.margin = unit(c(0.5,0.5,0.5,0.5), "cm"),
        # Make sure squares are obtained
        aspect.ratio=1,
        panel.border = element_rect(colour = "grey60", fill=NA, linewidth=0.5),
        axis.ticks = element_blank(),
        axis.title = element_blank(),
        axis.text = element_blank(),
        plot.title = element_text(hjust = 0.5, size = 20))
}

# Update titles of stimulatory conditions if needed
Dens.plotlist$IFNgamma<-Dens.plotlist[["IFNgamma"]]+
  labs(title = paste0("IFN-","\U03B3","/R-848"))
Dens.plotlist$`CD3/CD28_beads`<-Dens.plotlist$`CD3/CD28_beads`+
  labs(title = "Anti-CD3/CD28")

# Join plots
plt_OG_8h<-ggarrange(plotlist = Dens.plotlist, ncol = 1, nrow = 5)
plt_OG_8h<-annotate_figure(plt_OG_8h, 
                            top = text_grob("Metadata-based annotation - 8h",
                                            color = "red", 
                                            face = "bold", size = 14))
plt_OG_8h

#### 8H - REVISED ANNOTATION ####
Idents(IntData.B_8h.FilClus)<-'RevisedStim'
Dens.plotlist <- list()

# Generate the density plots and save them in the list
for (stim in c("ODN2006", "IFNgamma", "MegaCD40L", "antiCD3CD28", "Unstimulated", "Undefined")){
  Dens.plotlist[[stim]] <- DimPlot(IntData.B_8h.FilClus, reduction = "B_filtered_CCAInt.umap.rna",
                                 group.by = "RevisedStim", cols = rep("lightgrey", times = 6))+
    themeG+
    theme(legend.position = "none")+
    stat_density2d(data = subset(IntData.B_8h.FilClus, idents = stim)@reductions[["B_filtered_CCAInt.umap.rna"]]@cell.embeddings,
                   mapping = aes(x = BfilteredCCAIntumaprna_1,
                                 y = BfilteredCCAIntumaprna_2,
                                 fill = ..level..),
                   geom = "polygon", alpha = 0.6)+
    scale_fill_gradient(low = "lightblue",high = "darkblue", guide = "colourbar",
                        breaks = c(0,0.05, 0.1))+
    labs(title = stim, x = "UMAP1", y = "UMAP2")+
    theme(axis.line = element_blank(),
          axis.ticks = element_blank(),
          axis.text = element_blank(),
          axis.title = element_blank())+
    NoLegend()+
    coord_fixed()+
    ylim(-7,7)+
    xlim(-8,8)+
    theme(plot.margin = unit(c(0.5,0.5,0.5,0.5), "cm"),
          # Make sure squares are obtained
          aspect.ratio=1,
          panel.border = element_rect(colour = "grey60", fill=NA, linewidth=0.5),
          axis.ticks = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          plot.title = element_text(hjust = 0.5, size = 20))
}

# Update titles of stimulatory conditions if needed
Dens.plotlist$IFNgamma<-Dens.plotlist[["IFNgamma"]]+
  labs(title = paste0("IFN-","\U03B3","/R-848"))
Dens.plotlist$antiCD3CD28<-Dens.plotlist$antiCD3CD28+
  labs(title = "Anti-CD3/CD28")

# Join plots
plt_Rev_8h<-ggarrange(plotlist = Dens.plotlist, ncol = 1, nrow = 7)
plt_Rev_8h<-annotate_figure(plt_Rev_8h, 
                            top = text_grob("Nhood-based annotation - 8h",
                                            color = "red", 
                                            face = "bold", size = 14))
plt_Rev_8h

#### 30H - ORIGINAL ANNOTATION ####
Idents(IntData.B_30h.FilClus)<-'Stimulation'
Dens.plotlist <- list()

# Generate the density plots and save them in the list
for (stim in c("ODN2006", "IFNgamma", "MegaCD40L", "CD3/CD28_beads", "Unstimulated")){
  Dens.plotlist[[stim]] <- DimPlot(IntData.B_30h.FilClus, reduction = "B_filtered_CCAInt.umap.rna",
                                 group.by = "Stimulation", cols = rep("lightgrey", times = 5))+
    themeG+
    theme(legend.position = "none")+
    stat_density2d(data = subset(IntData.B_30h.FilClus, idents = stim)@reductions[["B_filtered_CCAInt.umap.rna"]]@cell.embeddings,
                   mapping = aes(x = BfilteredCCAIntumaprna_1,
                                 y = BfilteredCCAIntumaprna_2,
                                 fill = ..level..),
                   geom = "polygon", alpha = 0.6)+
    scale_fill_gradient(low = "lightblue",high = "darkblue", guide = "colourbar",
                        breaks = c(0,0.05, 0.1))+
    labs(title = stim, x = "UMAP1", y = "UMAP2")+
    theme(axis.line = element_blank(),
          axis.ticks = element_blank(),
          axis.text = element_blank(),
          axis.title = element_blank())+
    NoLegend()+
    coord_fixed()+
    ylim(-7,7)+
    xlim(-8,8)+
    theme(plot.margin = unit(c(0.5,0.5,0.5,0.5), "cm"),
          # Make sure squares are obtained
          aspect.ratio=1,
          panel.border = element_rect(colour = "grey60", fill=NA, linewidth=0.5),
          axis.ticks = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          plot.title = element_text(hjust = 0.5, size = 20))
}

# Update titles of stimulatory conditions if needed
Dens.plotlist$IFNgamma<-Dens.plotlist[["IFNgamma"]]+
  labs(title = paste0("IFN-","\U03B3","/R-848"))
Dens.plotlist$`CD3/CD28_beads`<-Dens.plotlist$`CD3/CD28_beads`+
  labs(title = "Anti-CD3/CD28")

# Join plots
plt_OG_30h<-ggarrange(plotlist = Dens.plotlist, ncol = 1, nrow = 5)
plt_OG_30h<-annotate_figure(plt_OG_30h, 
                            top = text_grob("Metadata-based annotation - 30h",
                                            color = "red", 
                                            face = "bold", size = 14))
plt_OG_30h

#### 30H - REVISED ANNOTATION ####
Idents(IntData.B_30h.FilClus)<-'RevisedStim'
Dens.plotlist <- list()

# Generate the density plots and save them in the list
for (stim in c("ODN2006", "IFNgamma", "MegaCD40L", "antiCD3CD28", "Unstimulated", "Undefined")){
  Dens.plotlist[[stim]] <- DimPlot(IntData.B_30h.FilClus, reduction = "B_filtered_CCAInt.umap.rna",
                                 group.by = "RevisedStim", cols = rep("lightgrey", times = 6))+
    themeG+
    theme(legend.position = "none")+
    stat_density2d(data = subset(IntData.B_30h.FilClus, idents = stim)@reductions[["B_filtered_CCAInt.umap.rna"]]@cell.embeddings,
                   mapping = aes(x = BfilteredCCAIntumaprna_1,
                                 y = BfilteredCCAIntumaprna_2,
                                 fill = ..level..),
                   geom = "polygon", alpha = 0.6)+
    scale_fill_gradient(low = "lightblue",high = "darkblue", guide = "colourbar",
                        breaks = c(0,0.05, 0.1))+
    labs(title = stim, x = "UMAP1", y = "UMAP2")+
    theme(axis.line = element_blank(),
          axis.ticks = element_blank(),
          axis.text = element_blank(),
          axis.title = element_blank())+
    NoLegend()+
    coord_fixed()+
    ylim(-7,7)+
    xlim(-8,8)+
    theme(plot.margin = unit(c(0.5,0.5,0.5,0.5), "cm"),
          # Make sure squares are obtained
          aspect.ratio=1,
          panel.border = element_rect(colour = "grey60", fill=NA, linewidth=0.5),
          axis.ticks = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank(),
          plot.title = element_text(hjust = 0.5, size = 20))
}

# Update titles of stimulatory conditions if needed
Dens.plotlist$IFNgamma<-Dens.plotlist[["IFNgamma"]]+
  labs(title = paste0("IFN-","\U03B3","/R-848"))
Dens.plotlist$antiCD3CD28<-Dens.plotlist$antiCD3CD28+
  labs(title = "Anti-CD3/CD28")

# Join plots
plt_Rev_30h<-ggarrange(plotlist = Dens.plotlist, ncol = 1, nrow = 7)
plt_Rev_30h<-annotate_figure(plt_Rev_30h, 
                            top = text_grob("Nhood-based annotation - 30h",
                                            color = "red", 
                                            face = "bold", size = 14))
plt_Rev_30h

# Save plots
ggsave(plot = plt_OG_8h,
      filename = file.path(FiguresOut_Path, "SuplFig_DensPlots_8h_MetaDataAnnot.png"),
       device = "png", dpi = 330,
       width = 5, height=25, bg = "white")
ggsave(plot = plt_OG_30h,
      filename = file.path(FiguresOut_Path, "SuplFig_DensPlots_30h_MetaDataAnnot.png"),
       device = "png", dpi = 330,
       width = 5, height=25, bg = "white")
ggsave(plot = plt_Rev_8h,
      filename = file.path(FiguresOut_Path, "SuplFig_DensPlots_8h_NhoodAnnot.png"),
       device = "png", dpi = 330,
       width = 5, height=35, bg = "white")
ggsave(plot = plt_Rev_30h,
      filename = file.path(FiguresOut_Path, "SuplFig_DensPlots_30h_NhoodAnnot.png"),
       device = "png", dpi = 330,
       width = 5, height=35, bg = "white")
```

## Correlation plots - Gene expression
```{r CorrelationPlots_GeneExpression, fig.width=20, fig.height=20}
DEG_Stim_Unstim_AllTimes_FilClus.Fixed.Meta.DF_reordered<-DEG_Stim_Unstim_AllTimes_FilClus.Fixed.Meta.DF%>%
  select(Gene, ODN2006_8h, IFNgamma_8h, MegaCD40L_8h, CD3CD28_8h,
         ODN2006_30h, IFNgamma_30h, MegaCD40L_30h, CD3CD28_30h,)

pltCorrDens<-ggpairs(DEG_Stim_Unstim_AllTimes_FilClus.Fixed.Meta.DF_reordered,
        columns = 2:dim(DEG_Stim_Unstim_AllTimes_FilClus.Fixed.Meta.DF)[2],
        lower = list(continuous = GGally::wrap(plot_lower_scatter_SelGenes.ggpairs, thres = log2FC_thres,
                                               DEG_Selected_DOWN=DEG_Stim_Unstim_8hvs30_FilClus.Fixed.Meta_DOWN,
                                               DEG_Selected_UP=DEG_Stim_Unstim_8hvs30_FilClus.Fixed.Meta_UP)),
        diag = list(continuous = GGally::wrap(plot_diag_density_SelGenes.ggpairs,
                                              thres = log2FC_thres,
                                              DEG_Selected_DOWN=DEG_Stim_Unstim_8hvs30_FilClus.Fixed.Meta_DOWN,
                                              DEG_Selected_UP=DEG_Stim_Unstim_8hvs30_FilClus.Fixed.Meta_UP)),
        upper = list(continuous = GGally::wrap(plot_upper_correlation_SelGenes.ggpairs,
                                               DEG_Selected_DOWN=DEG_Stim_Unstim_8hvs30_FilClus.Fixed.Meta_DOWN,
                                               DEG_Selected_UP=DEG_Stim_Unstim_8hvs30_FilClus.Fixed.Meta_UP)))
pltCorrDens+
  theme(axis.text = element_text(size = 12),
        axis.text.x = element_text(angle = 270, vjust = 0.5, hjust = 0),
        strip.background = element_blank(),
        strip.text = element_text(size = 17))

ggsave(filename = file.path(FiguresOut_Path, "SuplFig_ExpressionCorrelationPlots.pdf"),
       device = "pdf", dpi = 330,
       width = 15, height=20, bg = "white")
```

## Overlap in DE genes

```{r HeatmapOverlapDEGenes}
# Place in heatmap, without reordering of the columns
pdf(file.path(FiguresOut_Path, "SuplFig_Heatmap_Overlap_UpRegulatedGenes.pdf"), 
    width=9, height=8)
Heatmap(MatOverlap_UP,
        row_order = names(DEG_Stim_Unstim_8hvs30_FilClus.Fixed.Meta_UP),
        column_order = names(DEG_Stim_Unstim_8hvs30_FilClus.Fixed.Meta_UP),
        col = colorRamp2(c(0, 50, 100), c("white", "#8BD597", "darkgreen")),
        row_names_side = "left", column_names_side = "top",
        row_names_gp = gpar(fontsize = 15), column_names_gp = gpar(fontsize = 15),
        column_names_rot = 45,
        rect_gp = gpar(col = "white", lwd = 1.5),
        cell_fun = function(j, i, x, y, w, h, col) { 
          # add text to each grid
          grid.text(round(MatOverlap_UP[i, j], digits = 1), x, y,
                    gp = gpar(fontsize = 15, col = "black"))},
        name = "Absolute overlap")
dev.off()

# Place in heatmap, without reordering of the columns
pdf(file.path(FiguresOut_Path, "SuplFig_Heatmap_Overlap_DownRegulatedGenes.pdf"), 
    width=9, height=8)
Heatmap(MatOverlap_DOWN,
        row_order = names(DEG_Stim_Unstim_8hvs30_FilClus.Fixed.Meta_DOWN),
        column_order = names(DEG_Stim_Unstim_8hvs30_FilClus.Fixed.Meta_DOWN),
        row_names_gp = gpar(fontsize = 15), column_names_gp = gpar(fontsize = 15),
        col = colorRamp2(c(0, 350, 700), c("white", "#EF798C", "darkred")),
        rect_gp = gpar(col = "white", lwd = 1.5),
        column_names_rot = 45,
        row_names_side = "left", column_names_side = "top",
        cell_fun = function(j, i, x, y, w, h, col) { 
          # add text to each grid
          grid.text(round(MatOverlap_DOWN[i, j], digits = 1), x, y,
                    gp = gpar(fontsize = 15, col = "black"))},
        name = "Absolute overlap")
dev.off()
```

## Expression values of unique subsets
```{r ExpLevels_UniqueSubsets}
# PlotData
ggplot(topGenes_8h_30h_unique.Data.Fixed.Meta,
       mapping = aes(x = Stimulation,
                     y = Gene))+
  facet_grid(.~TimePoint)+
  geom_hline(mapping = aes(yintercept = Gene),
             color = "grey80", 
             alpha = 0.5)+
  geom_vline(mapping = aes(xintercept = Stimulation),
             color = "grey80",
             alpha = 0.5)+
  geom_point(mapping = aes(size = abs(Score),
                           fill = avg_log2FC),
             colour = "grey50",
             shape = 21)+
  geom_point(data = topGenes_8h_30h_unique.Data.Fixed.Meta%>%
               filter(InTop==TRUE),
             shape = 20, color = "black",
             size = 0.1)+
  scale_fill_gradient2(low = colDirection["DOWN"], 
                        mid = "white", 
                        high = colDirection["UP"],
                        midpoint = 0)+
  scale_size(range = c(1,6))+
  themeG+
  theme(axis.text.x = element_text(angle = 45,
                                   vjust = 1, hjust = 0, size = 10),
        axis.text.y = element_text(size = 10),
        axis.line = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank())+
  scale_x_discrete(position = "top")+
  ggtitle("Top genes based on  hard cut-off - Unique subsets only")

ggsave(filename = file.path(FiguresOut_Path, "SuplFig_ExpressionLevels_DEG_UniqueSubsets.pdf"),
       device = "pdf", dpi = 330,
       width = 6, height=10, bg = "white")
```

## GSEA showing top 10 GSE

```{r GSEA_Top10, fig.height=15, fig.width=15}
# Plot data
(ggplot(data = dotPlot.Data.Fixed.Meta[!is.na(dotPlot.Data.Fixed.Meta$SelectIn),])+
   geom_tile(mapping = aes(x = SelectIn,
                           y = GeneSet),
             na.rm = T, fill = "grey65")+
    themeG+theme(axis.title = element_blank())+
    scale_x_discrete(position = "top"))+
(ggplot(data = dotPlot.Data.Fixed.Meta)+
  geom_vline(mapping = aes(xintercept = Stimulation),
             color = "grey80", 
             alpha = 0.5)+
  geom_hline(mapping = aes(yintercept = GeneSet),
             color = "grey80", 
             alpha = 0.5)+
  geom_point(mapping = aes(x = Stimulation,
                     y = GeneSet,
                     size = Overlap,
                     fill  = -log10(AdjPval)),
             shape = 21, color = "grey50")+
  scale_fill_gradientn(colors = c("white", "lightblue", "darkblue", "#04074E"),
                        breaks = c(0, 10, 20, 30),
                        values = c(0, 0.05, 0.50, 0.95, 1))+
  scale_size(range = c(0,6))+
  themeG+
  theme(axis.text.x = element_text(angle = 45,
                                   vjust = 1, hjust = 0), 
        axis.line = element_blank(),
        axis.ticks = element_blank())+
  scale_x_discrete(position = "top")+
  facet_grid(.~TimePoint)+
  theme(axis.text.y = element_blank(),
        axis.title = element_blank()))+
  plot_layout(widths = c(1, 6))+ 
  plot_annotation(title = "Hard Cutoffs for Nhood selection")

# Save plot
ggsave(filename = file.path(FiguresOut_Path, "SuplFig_Top10_GSEA.pdf"),
       device = "pdf", dpi = 330,
       width = 12, height=15, bg = "white")
```