---
title: "Analysis Data QC scCITE"
author: "Tine D'hamers"
date: "2023-03-09"
output:
  html_document:
    theme: cosmo
    highlight: textmate
    df_print: paged
    toc: true
    toc_depth: 2
params:
  projectName:
    value: 'CVIDPatients_scCITESeq_ShallowDepth'
  sampleName:
    value: 'CEV023' 
  pathFiles:
    value: '/scratch/gent/vo/000/gvo00027/TOBI/Projects/CVID/Results/scCITE/Exp_Nov2023_PatientSamples_CEV23-26/ShallowDepthSeq_202311/NF_OutputData/CEV023/cellranger/CVIDPatients_scCITESeq_ShallowDepth_CEV023/outs'
  pathFiles_freemuxlet:
    value: '/scratch/gent/vo/000/gvo00027/TOBI/Projects/CVID/Results/scCITE/Exp_Nov2023_PatientSamples_CEV23-26/ShallowDepthSeq_202311/NF_OutputData/CEV023/freemuxlet/CEV023_freemuxletout'
  annotation_file:
    value: '/kyukon/data/gent/vo/000/gvo00027/TOBI/CVID/Data/scCITE/Exp_Nov2023_PatientSamples_CEV23-26/ShallowDepthSeq_202311/NF_OutputData/SampleInformation/20240104_SampleAnnotation_CVIDPatients.txt'
  features: 
    value: "CD80.1,CD86.1,CD40.1,CD3,CD3.1,CD8,CD8a,CD27.1,CD27.2,IgM,IgD,CD45RB,CD45RA,IgG1-k,IgG2a-k,IgG2b-k,IgG-Fc,CD38.1,CD38.2,CD19.1"
  outputDir:
    value: "/scratch/gent/vo/000/gvo00027/TOBI/Projects/CVID/Results/scCITE/Exp_Nov2023_PatientSamples_CEV23-26/ShallowDepthSeq_202311/QCReports"
---

# Sample `r params$sampleName`

## Initialise the working directory

```{r Variables}
# Specify the name of the sample and save it as the variable ‘sampleName’
projectName <- params$projectName
sampleName <- params$sampleName
# Output files of Cellranger process
pathFiles <- params$pathFiles
# Output files of freemuxlet process
pathFiles_freemuxlet <- params$pathFiles_freemuxlet
# Annotation file contains sample information
annotation_file <- params$annotation_file
genesets_visualise <- "/data/gent/vo/000/gvo00027/SingleCell10X/seuratintegrate/GeneSets/h.all.v7.1.symbols.gmt"
marker_genes_visualise <- "/data/gent/vo/000/gvo00027/SingleCell10X/seuratintegrate/GeneSets/markers.csv"
featuresADT <- unlist(strsplit(params$features,","))
```

### Create output dirs

```{r CreateOutDir,message=FALSE,warning=FALSE,tidy=TRUE}
##add some subfolders
output_dir_results <- file.path(params$outputDir, sampleName)
output_dir_results_qc <- paste0(output_dir_results,"/QC")
output_dir_robjects <- paste0(output_dir_results,"/Robjects")
dir.create(output_dir_results,recursive=TRUE)
dir.create(output_dir_results_qc,recursive=TRUE)
dir.create(output_dir_robjects,recursive = TRUE)
```

## Load packages

```{r SetupLibraries, include=FALSE}
library(tidyverse)
library(dplyr)
library(RColorBrewer)
library(Rtsne)
library(sva)
library(Seurat)
library(cowplot)
library(gridExtra)
library(scater)
library(patchwork)
library(biomaRt)
library(tximport)
library('openxlsx')
library("SeuratDisk",
          lib.loc = "/kyukon/data/gent/vo/000/gvo00027/SingleCell10X/seuratintegrate/Rpackages/R_4.2.1-foss-2022a/")
library(ggpubr)
#library(DoubletFinder)
```

```{r LoadPlotScript}
source("/data/gent/vo/000/gvo00027/SingleCell10X/nextflowpipeline/NF_scRNAseq/plotColors.R")
source("/kyukon/data/gent/vo/000/gvo00027/TOBI/CVID/Scripts/PlotThemes_sc_CVID.R")
source("/kyukon/data/gent/vo/000/gvo00027/TOBI/CVID/Scripts/Functions_sc_CVID.R")
```

## Read in data

```{r AnnotationFile}
# Annotation regarding the samples 
annotation <- read_tsv(annotation_file)
# Link correct color to correct annotation
# NOT NECESSARY TO ADJUST--> see PlotThemes_sc_CVID.R script
# names(colorsStim)<-annotation$Stimulation%>%unique()%>%sort()
# colorsPatients<-colorsPatients[1:length(annotation$PatientID%>%unique())]
# names(colorsPatients)<-annotation$PatientID%>%unique()%>%sort()

# Extract information only related to the pool of interest
annotation_sample <- annotation %>% 
  filter(SampleID == sampleName)
```

```{r ReadCellRangerData}
# Directory contains files:
# barcodes.tsv.gz
# features.tsv.gz
# matrix.mtx.gz
cellranger_data <- Read10X(data.dir = paste0(pathFiles,
                                             "/per_sample_outs/",
                                             projectName, "_", sampleName,
                                             "/count/sample_filtered_feature_bc_matrix"),
                           # Remove tailing "-1" if present in all cell barcodes
                           strip.suffix=TRUE)

```

## Create Seurat Object

### Create seurat

```{r CreateSeurat}
# Cells in columns (cell barcodes)
# Features in rows (genes)
seuratObj <- CreateSeuratObject(counts = cellranger_data$`Gene Expression`,
                                project = projectName)
```

The sample has `r nrow(seuratObj)` genes and `r ncol(seuratObj)` cells.

### Add annotation

```{r AnnotationSeurat}
# Add extra information to samples in metadata of Seurat object
seuratObj[["Full_samplename"]] <- paste0(projectName,"_",sampleName)
seuratObj[["samplename"]] <- sampleName
```

The sample has `r table(seuratObj@meta.data$orig.ident)` cells of sample `r rownames(table(seuratObj@meta.data$orig.ident))`

## QC with Seurat

### Create QC table

```{r CreateQCTable}
# Building data frame
qc_df <- tibble(qc_stats = "number of cells prefiltered",
                values= table(seuratObj@meta.data$orig.ident)[1])
```

### Calculated mitochondrial genes

Mitochondrial (MT) genes are present in the index, so quantification of MT genes is possible. The percentage of MT genes needs to be low. High percentages of MT genes can be an indication of damaged cells.

```{r MTGenes, message=FALSE, warning=FALSE,tidy=T}
#Calculate percent mitochondrial genes (MT for human, mt for mouse/Zebrafish)
seuratObj[["percent_mito"]] <- PercentageFeatureSet(object = seuratObj, 
                                                    pattern = "^MT")
# Add to QCTable
qc_df <- qc_df %>% 
  bind_rows(list(qc_stats = "mean mitochondrial percentage prefiltered", 
                 values = mean(seuratObj@meta.data$percent_mito)), 
            list(qc_stats = "median mitochondrial percentage prefiltered",
                 values = median(seuratObj@meta.data$percent_mito)),
            list(qc_stats = "mean nFeature RNA prefiltered", 
                 values = mean(seuratObj@meta.data$nFeature_RNA)), 
            list(qc_stats = "median nFeature RNA prefiltered", 
                 values= median(seuratObj@meta.data$nFeature_RNA)),
            list(qc_stats = "mean nCount RNA prefiltered", 
                 values = mean(seuratObj@meta.data$nCount_RNA)), 
            list(qc_stats = "median nCount RNA prefiltered", 
                 values= median(seuratObj@meta.data$nCount_RNA)))
```

### Plot QC features {.tabset .tabset-fade .tabset-pills}

#### All features

```{r PlotAllFeatures, message=FALSE, warning=FALSE,tidy=T}
# Use Violin plot function from external script
VlnfeaturePlot(seuratObj, c("nFeature_RNA","nCount_RNA","percent_mito"))+
  ylim(0,NA)+
  theme(axis.text.x = element_blank())+
  ylab("Value")
# Save as pdf
ggsave(filename = "Violin_features_beforenorm_adjTrans.pdf", 
       path= output_dir_results_qc, 
       width = 20, device = "pdf")
# Save as jpeg
ggsave(filename = "Violin_features_beforenorm_adjTrans.jpeg", 
       path= output_dir_results_qc, 
       width = 20, device = "jpeg")
```

#### nCount vs mito%

Every dot represents one cell. Per cell, the total number of RNA molecules detected in that cell is presented on the horizontal axis and the percentage of mitochondrial genes detected in that cell is presented on the vertical axis. Cells characterized by low percentages in mitochondrial genes and high amounts of RNA molecules are preferred. Cells with low number of RNA molecules and high percentages of mitochondrial genes are most likely damaged cells.

```{r ScatternCountMito, message=FALSE, warning=FALSE,tidy=T}
scatterplotseurat(seuratObj, xfeature = "nCount_RNA", 
                  yfeature = "percent_mito")+
  ggtitle("%mito vs. RNA counts")+
  theme_celine
# Save as pdf
ggsave(filename = "Scatter_nCount_mito.pdf", 
       path= output_dir_results_qc, 
       width = 10, height = 7, device = "pdf")
# Save as jpeg
ggsave(filename = "Scatter_nCount_mito.jpeg", 
       path= output_dir_results_qc, 
       width = 10, height = 7, device = "jpeg")
```

#### nCounts vs nFeatures

Every dot represents one cell. Per cell, the total number of RNA molecules detected in that cell is presented on the horizontal axis and the total number of unique genes detected in that cell is presented on the vertical axis.

```{r ScatternCountsnFeatures, message=FALSE, warning=FALSE,tidy=T}
scatterplotseurat(seuratObj, xfeature = "nCount_RNA", 
                  yfeature = "nFeature_RNA")+
  ggtitle("Unique RNAs vs. RNA counts")+
  theme_celine
# Save as pdf
ggsave(filename = "Scatter_nCount_nUMI.pdf", 
       path= output_dir_results_qc, 
       width = 10, height = 7, device = "pdf")
# Save as jpeg
ggsave(filename = "Scatter_nCount_nUMI.jpeg", 
       path= output_dir_results_qc, 
       width = 10, height = 7, device = "jpeg")
```

## Preprocessing

### Filter Cells

```{r FilterCells}
# Thresholds for cells of interest (intact, not damaged)
seuratObj <- subset(seuratObj, 
                    subset = nFeature_RNA > 200 & percent_mito < 25)
```

```{r qcUpdate1}
qc_df <- qc_df %>% 
  bind_rows(list(qc_stats = "mean mitochondrial percentage postfiltered", 
                 values = mean(seuratObj@meta.data$percent_mito)), 
            list(qc_stats = "median mitochondrial percentage postfiltered",
                 values = median(seuratObj@meta.data$percent_mito)),
            list(qc_stats = "mean nFeature RNA postfiltered", 
                 values= mean(seuratObj@meta.data$nFeature_RNA)), 
            list(qc_stats = "median nFeature RNA postfiltered", 
                 values= median(seuratObj@meta.data$nFeature_RNA)),
            list(qc_stats = "mean nCount RNA postfiltered", 
                 values= mean(seuratObj@meta.data$nCount_RNA)), 
            list(qc_stats = "median nCount RNA postfiltered", 
                 values= median(seuratObj@meta.data$nCount_RNA)),
            list(qc_stats = "number of cells postfiltered",
                 values = table(seuratObj@meta.data$orig.ident)[1]))
```

### Normalise RNA data

Based on the normalization method of choice using the `NormalizeData` [Seurat function](https://satijalab.org/seurat/reference/normalizedata), different transformations of the data are possible:

-   **LogNormalize**: Feature counts for each cell are divided by the total counts for that cell and multiplied by the scale.factor. This is then natural-log transformed using log1p.
-   **CLR**: Applies a centered log ratio transformation
-   **RC**: Relative counts: Feature counts for each cell are divided by the total counts for that cell and multiplied by the scale.factor. No log-transformation is applied. For counts per million (CPM) set scale.factor = 1e6

Note: Using a log transformation can result in negative values

```{r NormaliseData, message=FALSE, warning=FALSE,tidy=T}
# Normalisation of the count data
# Stored in seuratObj@assays$RNA@data
seuratObj <- NormalizeData(object = seuratObj, 
                           normalization.method = "LogNormalize", 
                           scale.factor = 10000)

# Columns are the different cells
# Total number of RNA molecules detected (use normalized data)
seuratObj[["nUMI"]] <- colSums(as.matrix(seuratObj[['RNA']]@data))

# Apply function on columns of matrix (argument MARGIN=2)
# Same as nFeature_RNA (look for number of unique genes expressed per cell)
seuratObj[["nGene"]] <- apply(as.matrix(seuratObj[['RNA']]@data),
                              2, function(x){sum(x>0)})

# Plot violin plot
VlnfeaturePlot(seuratObj, features=c("nUMI","nGene"))+
  ylim(min(seuratObj[["nUMI"]], seuratObj[["nGene"]]),
       max(seuratObj[["nUMI"]], seuratObj[["nGene"]]))+
  ylab("Value")+
  theme(axis.text.x = element_blank())
# Save as pdf
ggsave(filename = "Violin_nUMI_nGene_afternormalisation.pdf", 
       path= output_dir_results_qc, 
       width = 20, device = "pdf")
# Save as jpeg
ggsave(filename = "Violin_nUMI_nGene_afternormalisation.jpeg", 
       path= output_dir_results_qc, 
       width = 20, device = "jpeg")

qc_df <- qc_df %>% 
  bind_rows(list(qc_stats = "mean UMI postnormalised", 
                 values = mean(seuratObj@meta.data$nUMI)), 
            list(qc_stats = "median UMI postnormalised", 
                 values = median(seuratObj@meta.data$nUMI)), 
            list(qc_stats = "mean gene postnormalised", 
                 values = mean(seuratObj@meta.data$nGene)), 
            list(qc_stats = "median gene postnormalised", 
                 values = median(seuratObj@meta.data$nGene)))

```

### Determine highly variable genes (used for downstream clustering)

To find variable genes, the [Seurat function](https://satijalab.org/seurat/reference/findvariablefeatures) `FindVariableFeatures` can make use of three different selection methods to identify features that are outliers on a 'mean variability plot'.: 

- **vst**: First, fits a line to the relationship of log(variance) and log(mean) using local polynomial regression (loess). Then standardizes the feature values using the observed mean and expected variance (given by the fitted line). Feature variance is then calculated on the standardized values after clipping to a maximum (see clip.max parameter). 
- **mean.var.plot** (mvp): First, uses a function to calculate average expression (mean.function) and dispersion (dispersion.function) for each feature. Next, divides features into num.bin (deafult 20) bins based on their average expression, and calculates z-scores for dispersion within each bin. The purpose of this is to identify variable features while controlling for the strong relationship between variability and average expression. 
- **dispersion** (disp): selects the genes with the highest dispersion values

```{r HighlyVariableGenes, message=F, warning=F, tidy=T}
seuratObj <- FindVariableFeatures(object = seuratObj, 
                                  selection.method = "vst", 
                                  # Top variable features to select
                                  nfeatures = 2000)
# Information is saved in seuratObj@assays[["RNA"]]@meta.features

### Plot variable features with and without labels
top10 <- head(x = VariableFeatures(object = seuratObj), 10)
plot1 <- VariableFeaturePlot(object = seuratObj)
# Labeling top 10 variable genes in plot
plot2 <- LabelPoints(plot = plot1, 
                     points = top10, repel = TRUE)+
  theme_celine

print(plot2)
# Save as pdf
ggsave(filename = "Highly_Variable_Genes.pdf", 
       path= output_dir_results_qc, 
       width = 10, height = 7, device = "pdf")
# Save as jpeg
ggsave(filename = "Highly_Variable_Genes.jpeg", 
       path= output_dir_results_qc, 
       width = 10, height = 7, device = "jpeg")
```

Found `r length(VariableFeatures(seuratObj))` high variable genes!

### Scaling RNA data

The [Seurat function](https://satijalab.org/seurat/reference/scaledata) performs a gene-level scaling in which the counts of every feature will be centered to have a mean of 0 and scaled using the standard deviation as scaling factor. The number of UMI's can consequently show negative counts since features or genes showing counts less than the mean of all the counts for that gene will result in negative values afters centering. The number of unique genes will always be positive since we are calculating the number of genes with a count value higher than 0.

```{r Scaling, message=FALSE, warning=FALSE,tidy=T}
seuratObj <- ScaleData(object = seuratObj, features = rownames(seuratObj))

##### Check per group #####
# Columns are the different cells
seuratObj[['nUMI_scaled']]<-colSums(as.matrix(seuratObj[['RNA']]@scale.data))
seuratObj[['nGene_scaled']]<-apply(as.matrix(seuratObj[['RNA']]@scale.data),
                                   2,function(x){sum(x>0)})

VlnfeaturePlot(seuratObj, features=c("nUMI_scaled","nGene_scaled"))+
  ylim(min(seuratObj[["nUMI_scaled"]], seuratObj[["nGene_scaled"]]),
       max(seuratObj[["nUMI_scaled"]], seuratObj[["nGene_scaled"]]))+
  theme(axis.text.x = element_blank())
# Save as pdf
ggsave(filename = "Violin_nUMI_nGene_afsterscaling.pdf", 
       path= output_dir_results_qc, 
       width = 20, device = "pdf")
# Save as jpeg
ggsave(filename = "Violin_nUMI_nGene_afsterscaling.jpeg", 
       path= output_dir_results_qc, 
       width = 20, device = "jpeg")

qc_df <- qc_df %>% bind_rows(list(qc_stats = "mean UMI postscaling", 
                                  values = mean(seuratObj@meta.data$nUMI_scaled)), 
                             list(qc_stats = "median UMI postscaling", 
                                  values = median(seuratObj@meta.data$nUMI_scaled)),
                             list(qc_stats = "mean gene postscaling", 
                                  values = mean(seuratObj@meta.data$nGene_scaled)),
                             list(qc_stats = "median gene postscaling", 
                                  values = median(seuratObj@meta.data$nGene_scaled)))
```

### Correlation UMIs and number of unique genes

```{r}
scatterplotseurat(seuratObj, xfeature = 'nGene',
                  yfeature = 'nUMI')+
  ggtitle("Before Scaling")+
  theme_celine+
  scatterplotseurat(seuratObj, xfeature = 'nGene_scaled',
                    yfeature = 'nUMI_scaled')+
  ggtitle("After scaling")+
  theme_celine
# Save as pdf
ggsave(filename = "Scatter_nUMI_nGene_correlation.pdf", 
       path= output_dir_results_qc, 
       width = 20, device = "pdf")
# Save as jpeg
ggsave(filename = "Scatter_nUMI_nGene_correlation.jpeg", 
       path= output_dir_results_qc, 
       width = 20, device = "jpeg")
```

```{r}
density_plot<-rbind(
  data.frame(nUMI = seuratObj[['nUMI']],
             nGene = seuratObj[["nGene"]],
             scaled = "Before scaling")%>%
  pivot_longer(cols = c(nUMI,nGene),
               names_to = "Molecule",
               values_to = "Counts"),
  data.frame(nUMI_scaled = seuratObj[['nUMI_scaled']],
             nGene_scaled = seuratObj[["nGene_scaled"]],
             scaled = "After scaling")%>%
  pivot_longer(cols = c(nUMI_scaled,nGene_scaled),
               names_to = "Molecule",
               values_to = "Counts")
)

colors<-rep(colorsHashtags[c(1,4)], times = 2)
names(colors)<-c("Before scaling", "After scaling", "Before scaling", "After scaling")

ggplot(data=density_plot%>%
         filter(Molecule == "nUMI" |
                  Molecule == "nUMI_scaled"),
       mapping = aes(Counts, fill = scaled))+
  # Before scaling
  geom_density(alpha = 0.5)+
  ylab("Density")+
  xlab("nUMI")+
  xlim(min(density_plot$Counts),NA)+
  theme_celine+
  scale_fill_manual(values = colors,
                    name = "", 
                    breaks = "")+
  ggtitle("")+
  theme(plot.title=element_text(hjust=0.5),
        legend.position = "right")+
  theme_celine+
  # After scaling
  ggplot(data=density_plot%>%
         filter(Molecule == "nGene" |
                  Molecule == "nGene_scaled"),
       mapping = aes(Counts, fill = scaled))+
  geom_density(alpha = 0.5)+
  ylab("")+
  xlab("nGene")+
  theme_celine+
  scale_fill_manual(values = colors,
                    name = "", 
                    breaks = c("Before scaling","After scaling"), 
                    labels = c("Before", "After"))+
  ggtitle("")

# Save as pdf
ggsave(filename = "Density_nUMI_nGene.pdf", 
       path= output_dir_results_qc, 
       width = 10, height = 10, device = "pdf")
# Save as jpeg
ggsave(filename = "Density_nUMI_nGene.jpeg", 
       path= output_dir_results_qc, 
       width = 10, height = 10, device = "jpeg")

rm(density_plot, colors)
```

## Demultiplexing

### Extract cells with gene expression and antibody capture data

```{r JointCellNames}
# Look for cell both present in the gene expression and antibody capture data
joint.cellnames <- intersect(rownames(seuratObj@meta.data),
                             colnames(cellranger_data$`Antibody Capture`))
```

### Demultiplexing based on SNPs

```{r Freemuxlet}
freemuxlet_dt  <- read_tsv(paste0(pathFiles_freemuxlet,"/",
                                  sampleName,"_freemuxlet.pooled.clust1.samples.gz"))

freemuxlet_dt <- freemuxlet_dt %>% mutate(doublets = case_when(
  # Extract the type of droplet
  DROPLET.TYPE == "SNG" ~ "singlet",
  DROPLET.TYPE == "DBL" ~ "double",
  DROPLET.TYPE == "AMB" ~ "non-defined",
), sample = case_when(
  # Extract the best sample guess
  BEST.GUESS == "1,1" ~ "sample1",
  BEST.GUESS == "0,0" ~ "sample2",
  BEST.GUESS == "2,2" ~ "sample3",
  BEST.GUESS == "3,3" ~ "sample4",
  BEST.GUESS == "4,4" ~ "sample5",
  BEST.GUESS == "5,5" ~ "sample6"
), cellbarcode = sub("-.*","", BARCODE)) %>% 
  dplyr::select(cellbarcode, doublets, sample) %>% 
  column_to_rownames("cellbarcode")

# Add extra information in the meta data of the seurat object
seuratObj@meta.data["FM_doublets"]<-freemuxlet_dt[joint.cellnames,"doublets"]
seuratObj@meta.data["FM_sample"]<-freemuxlet_dt[joint.cellnames,"sample"]

# Information related to doublets, singlets or non-defined GEM
table(seuratObj$FM_doublets)

qc_df<- qc_df %>% 
  bind_rows(list(qc_stats="doublet FM", 
                 values = table(seuratObj$FM_doublets)["double"]),
            list(qc_stats = "negative FM", 
                 values= table(seuratObj$FM_doublets)["non-defined"]),
            list(qc_stats = "singlet FM", 
                 values = table(seuratObj$FM_doublets)["singlet"]))

# Information related to samples
table(seuratObj$FM_sample)

qc_df<- qc_df %>% bind_rows(lapply(1:4,
                   function(i){
                     sample<-paste0("sample",i)
                     list(qc_stats=paste0(sample, ' FM'),
                          values = table(seuratObj$FM_sample)[sample])}))
```

### Demultiplexing based on hashes

```{r DemulHashing}
# For this experiment, 6-8 samples were pooled in 1 pooled sample 
# Making use of 6-8 different hashtags per pooled sample
hashtagsUsed<-annotation_sample$Hash%>%sort()
hashtagsTotal<-annotation$Hash%>%unique()%>%sort()
names(colorsHashtags)<-hashtagsTotal

# Hashtags in rows, cells in columns
seuratObj[["HTO"]]  <- CreateAssayObject(counts = as.matrix(cellranger_data$`Antibody Capture`[hashtagsUsed,joint.cellnames]))
seuratObj <- NormalizeData(seuratObj, assay = "HTO", 
                           normalization.method = "CLR")

# Calculate number of cells lacking hashtag
qc_df <- qc_df %>%
  bind_rows(list(qc_stats = "non hashed cells",
                 values = sum(seuratObj@meta.data[["nFeature_HTO"]]==0)),
            list(qc_stats = "hashed cells",
                 values = sum(seuratObj@meta.data[["nFeature_HTO"]]!=0)))

qc_df<-qc_df %>% 
  bind_rows(list(qc_stats = "percentage non hashed cells",
                 values = qc_df$values[qc_df$qc_stats == "non hashed cells"]/
                   qc_df$values[qc_df$qc_stats == "number of cells postfiltered"]*100),
            list(qc_stats = "percentage hashed cells",
                 values = qc_df$values[qc_df$qc_stats == "hashed cells"]/
                   qc_df$values[qc_df$qc_stats == "number of cells postfiltered"]*100))

# Select for cells with hashtag to make demultiplexing step possible
seuratObj<-subset(seuratObj,
                  subset = nFeature_HTO!=0)

seuratObj <- HTODemux(seuratObj, assay = "HTO", positive.quantile = 0.99)
```

```{r CreateOutDirHashing}
output_dir_results_hashing <- paste0(output_dir_results,"/Hashing")
dir.create(output_dir_results_hashing, recursive=TRUE)
```

```{r DemulMetrics}
table(seuratObj$HTO_classification.global)
# Add classification based on HTO
qc_df<- qc_df %>% 
  bind_rows(list(qc_stats="doublet HTO", 
                 values = table(seuratObj$HTO_classification.global)["Doublet"]),
            list(qc_stats = "negative HTO", 
                 values= table(seuratObj$HTO_classification.global)["Negative"]),
            list(qc_stats = "singlet HTO", 
                 values = table(seuratObj$HTO_classification.global)["Singlet"]))

# Add information on hashtags
qc_df<- qc_df %>% 
  bind_rows(lapply(hashtagsTotal,
                   function(i){
                     list(qc_stats=paste(i, 'HTO'),
                          values = table(seuratObj$HTO_maxID)[i])}))

Idents(seuratObj) <- "HTO_maxID"
RidgePlot(seuratObj, assay = "HTO", 
          features = rownames(cellranger_data$`Antibody Capture`[hashtagsUsed,joint.cellnames]), 
          ncol = 2,
          cols = alpha(colorsHashtags,0.5),
          # Show hashtag with highest expression level in the end
          # No peak visible at expression level 0 if correct demultiplexing
          sort = "increasing")
# Save as pdf
ggsave(filename = "Distribution_hashtags.pdf", 
       path= output_dir_results_hashing, 
       width = 10, height = 15, device = "pdf")
# Save as jpeg
ggsave(filename = "Distribution_hashtags.jpeg", 
       path= output_dir_results_hashing, 
       width = 10, height = 15, device = "jpeg")

# Change default assay to avoid warnings
DefaultAssay(seuratObj) <- 'HTO'
plot<-lapply(1:(length(hashtagsUsed)-1),
             function(i){
               # Adjust colors based on the hashtags of interest
               colors<-rep("lightgrey", 8)
               names(colors)<-hashtagsUsed
               colors[i]<-colorsHashtags[i]
               colors[i+1]<-colorsHashtags[i+1]
               FeatureScatter(seuratObj, 
                                        feature1=hashtagsUsed[i+1],
                                        feature2 = hashtagsUsed[i])+
                 scale_color_manual(values = alpha(colors, 0.7),
                                    breaks = c(hashtagsUsed[i], hashtagsUsed[i+1]),
                                    name = "")+
                 theme_celine})
ggarrange(plotlist = plot)
# Save as pdf
ggsave(filename = "Scatter_hashtags.pdf", 
       path= output_dir_results_hashing, 
       width = 20, height = 20, device = "pdf")
# Save as jpeg
ggsave(filename = "Scatter_hashtags.jpeg", 
       path= output_dir_results_hashing, 
       width = 20, height = 20, device = "jpeg")
DefaultAssay(seuratObj) <- 'RNA'
```

```{r VlnPlotHTO}
Idents(seuratObj) <- "HTO_classification.global"
VlnPlot(seuratObj, features = "nCount_RNA", 
        pt.size = 0.1, log = TRUE)+
  theme_celine+
  theme(axis.text.x = element_text(angle=0, vjust=5, hjust=0.5),
                              strip.background = element_blank(),
                              axis.ticks.x = element_blank(),
                              axis.line.x = element_blank())+
  NoLegend()
# Save as pdf
ggsave(filename = "Violin_hashtags_doublets.pdf", 
       path= output_dir_results_hashing, 
       width = 20, device = "pdf")
# Save as jpeg
ggsave(filename = "Violin_hashtags_doublets.jpeg", 
       path= output_dir_results_hashing, 
       width = 20, device = "jpeg")
```

```{r HeatmapHTO}
HTOHeatmap(seuratObj, assay = "HTO", ncells = 5000)+
  scale_fill_gradientn(colors = c("#ECCFF2", "#418DF1", "#5F4699"))
# Save as pdf
ggsave(filename = "Heatmap_hashtags.pdf", 
       path= output_dir_results_hashing, 
       width = 20, device = "pdf")
# Save as jpeg
ggsave(filename = "Heatmap_hashtags.jpeg", 
       path= output_dir_results_hashing, 
       width = 20, device = "jpeg")
```

### Overlap Hashing - Freemuxlet (SNPs)

```{r CompareDemuxFreemux}
HTO_FM_compare <- as.data.frame(seuratObj@meta.data) %>%
  dplyr::select(HTO_classification.global, 
                FM_doublets, 
                HTO_classification, 
                FM_sample)

classes_HTO_FM <- HTO_FM_compare %>% 
  group_by(HTO_classification, FM_sample) %>% 
  summarise(counts_combos = n()) %>% 
  group_by(HTO_classification) %>% 
  top_n(1, wt = counts_combos) %>% 
  filter(HTO_classification %in% hashtagsUsed)

# Add the patientID 
# This must be comparable to the individuals found by Freemuxlet
classes_HTO_FM<-left_join(x = classes_HTO_FM, 
                          annotation_sample[c("Hash", "PatientID")],
                          by = c("HTO_classification" = "Hash"))

write.table(x = classes_HTO_FM,
          file = paste0(output_dir_results_hashing,"/HTO_FMO_counts.txt"),
          sep = "\t",
          col.names = TRUE,
          quote = FALSE)

print(classes_HTO_FM)
```

```{r CompareDemuxFreemux2}
# Look for the same annotation based on freemuxlet and hashtags
HTO_FM_compare <- HTO_FM_compare %>%
  mutate(FMHTO_classification = case_when(
    # If doublet seen in either hashtags or freemuxlet --> Doublet
    HTO_classification.global == "Doublet" | FM_doublets == "double"~ "Doublet",
    # Hashtags match with freemuxlet --> correct hashtag
    # Or Hashtag is identified but freemuxlet not --> only relying on hashtag
    (HTO_classification == "Hashtag1" & FM_sample %in% classes_HTO_FM[classes_HTO_FM$HTO_classification == "Hashtag1", "FM_sample"]) |
      (HTO_classification == "Hashtag1" & FM_doublets == "non-defined") ~ "Hashtag1",
    (HTO_classification == "Hashtag2" & FM_sample %in% classes_HTO_FM[classes_HTO_FM$HTO_classification == "Hashtag2", "FM_sample"]) | 
      (HTO_classification == "Hashtag2" & FM_doublets == "non-defined") ~ "Hashtag2",
    (HTO_classification == "Hashtag3" & FM_sample %in% classes_HTO_FM[classes_HTO_FM$HTO_classification == "Hashtag3", "FM_sample"]) |
      (HTO_classification == "Hashtag3" & FM_doublets == "non-defined") ~ "Hashtag3",
    (HTO_classification == "Hashtag4" & FM_sample %in% classes_HTO_FM[classes_HTO_FM$HTO_classification == "Hashtag4", "FM_sample"]) |
      (HTO_classification == "Hashtag4" & FM_doublets == "non-defined") ~ "Hashtag4",
    (HTO_classification == "Hashtag5" & FM_sample %in% classes_HTO_FM[classes_HTO_FM$HTO_classification == "Hashtag5", "FM_sample"]) |
      (HTO_classification == "Hashtag5" & FM_doublets == "non-defined") ~ "Hashtag5",
    (HTO_classification == "Hashtag6" & FM_sample %in% classes_HTO_FM[classes_HTO_FM$HTO_classification == "Hashtag6", "FM_sample"]) |
      (HTO_classification == "Hashtag6" & FM_doublets == "non-defined") ~ "Hashtag6",
    (HTO_classification == "Hashtag7" & FM_sample %in% classes_HTO_FM[classes_HTO_FM$HTO_classification == "Hashtag7", "FM_sample"]) |
      (HTO_classification == "Hashtag7" & FM_doublets == "non-defined") ~ "Hashtag7",
    (HTO_classification == "Hashtag8" & FM_sample %in% classes_HTO_FM[classes_HTO_FM$HTO_classification == "Hashtag8", "FM_sample"]) |
      (HTO_classification == "Hashtag8" & FM_doublets == "non-defined") ~ "Hashtag8",
    # Both hashtag and freemuxlet can not identify sample
    HTO_classification.global == "Negative" & FM_doublets == "non-defined" ~ "TrueNegative",
    # Only hashtags are undefined --> only relying on freemuxlet
    HTO_classification.global == "Negative" & FM_sample == "sample1" ~ 
      unique(classes_HTO_FM[classes_HTO_FM$FM_sample == "sample1", "PatientID"]$PatientID),
    HTO_classification.global == "Negative" & FM_sample == "sample2" ~ 
      unique(classes_HTO_FM[classes_HTO_FM$FM_sample == "sample2", "PatientID"]$PatientID),
    HTO_classification.global == "Negative" & FM_sample == "sample3" ~ 
      unique(classes_HTO_FM[classes_HTO_FM$FM_sample == "sample3", "PatientID"]$PatientID),
    HTO_classification.global == "Negative" & FM_sample == "sample4" ~ 
      unique(classes_HTO_FM[classes_HTO_FM$FM_sample == "sample4", "PatientID"]$PatientID),
    HTO_classification.global == "Negative" & FM_sample == "sample5" ~ 
      unique(classes_HTO_FM[classes_HTO_FM$FM_sample == "sample5", "PatientID"]$PatientID),
    HTO_classification.global == "Negative" & FM_sample == "sample6" ~ 
      unique(classes_HTO_FM[classes_HTO_FM$FM_sample == "sample6", "PatientID"]$PatientID),
    # Otherwise seen as unclassified
    TRUE~ "unclassified"
  ))
```

Per pool of samples, at least one healthy donor is unique. This means that for a certain healthy donor, only one stimulated sample is present in the pool. Consequently, it is possible to link the correct sample with the correct hashtag based on the output and identification of freemuxlet.

```{r UniqueDonors}
# Find the unique donor or donors
Unique<-annotation_sample$PatientID%>%table==1
Unique<-names(annotation_sample$PatientID%>%table)[Unique]


HTO_FM_compare<-HTO_FM_compare%>%
  mutate(FMHTO_classification = case_when(
    # If donor is unique in pool then use correct hashtag
    "P1" %in% Unique & FMHTO_classification == "P1" ~ (classes_HTO_FM%>%filter(PatientID=="P1"))$HTO_classification[1],
    "P2" %in% Unique & FMHTO_classification == "P2" ~ (classes_HTO_FM%>%filter(PatientID=="P2"))$HTO_classification[1],
    "P3" %in% Unique & FMHTO_classification == "P3" ~ (classes_HTO_FM%>%filter(PatientID=="P3"))$HTO_classification[1],
    "P4" %in% Unique & FMHTO_classification == "P4" ~ (classes_HTO_FM%>%filter(PatientID=="P4"))$HTO_classification[1],
    "P5" %in% Unique & FMHTO_classification == "P5" ~ (classes_HTO_FM%>%filter(PatientID=="P5"))$HTO_classification[1],
    "P6" %in% Unique & FMHTO_classification == "P6" ~ (classes_HTO_FM%>%filter(PatientID=="P6"))$HTO_classification[1],
    # Otherwise
    TRUE ~ FMHTO_classification)
  )

rm(Unique)

seuratObj@meta.data[["HTO_FM_consensus"]] <- HTO_FM_compare$FMHTO_classification
Idents(seuratObj) <- "HTO_FM_consensus"

table(seuratObj$HTO_FM_consensus)

qc_df<- qc_df %>% 
  bind_rows(lapply(hashtagsTotal,
                   function(i){
                     list(qc_stats=paste0(i, ' consensus 1'),
                          values = table(seuratObj$HTO_FM_consensus)[i])}))%>%
  bind_rows(list(qc_stats="doublet consensus", 
                 values = table(seuratObj$HTO_FM_consensus)["Doublet"]),
            list(qc_stats="true negatives consensus",
                 values = table(seuratObj$HTO_FM_consensus)["TrueNegative"]),
            # Contradiction between freemuxlet and hashtag (will be removed)
            list(qc_stats="unidentified consensus", 
                 values = table(seuratObj$HTO_FM_consensus)["unclassified"]),
            lapply(annotation$PatientID%>%unique()%>%sort(),
                   function(i){
                     list(qc_stats=paste0(i, ' consensus 1'),
                          values = table(seuratObj$HTO_FM_consensus)[i])}))
```

### Demultiplexing of the unidentified samples

The samples showing contradictions between the output of freemuxlet and the hashtags based on sample annotation are lost. But the samples that could not be identified based on the hashtags but were identified based on their SNPs (freemuxlet) can be used in an additional demultiplexing step. Filtering out the already identified cells reduces the complexity of the regression problem which could result in new findings based on the hashtags.

```{r ExtraDemulStep, eval = FALSE}
##### This code chunk is not evaluated because every sample is unique in the pool #####

# Filter out the cells only identified based on freemuxlet
seuratObj_unident <- subset(seuratObj, 
                            idents =  grep("P",unique(seuratObj$HTO_FM_consensus), value=TRUE))
# Demultiplexing of these cells based on their hashtags
seuratObj_unident <- HTODemux(seuratObj_unident, assay = "HTO", 
                              positive.quantile = 0.99,
                              init=2)
# Heatmap
HTOHeatmap(seuratObj_unident, assay = "HTO", ncells = 5000)+
  scale_fill_gradientn(colors = c("#ECCFF2", "#418DF1", "#5F4699"))
# Save as pdf
ggsave(filename = "Heatmap_hashtags_unidentified.pdf", 
       path= output_dir_results_hashing, 
       width = 20, device = "pdf")
# Save as jpeg
ggsave(filename = "Heatmap_hashtags_unidentified.jpeg", 
       path= output_dir_results_hashing, 
       width = 20, device = "jpeg")
```


```{r CompareDemuxFreemuxExtra, eval = F}
##### This code chunk is not evaluated because every sample is unique in the pool #####
HTO_FM_compare2 <- as.data.frame(seuratObj_unident@meta.data) %>%
  dplyr::select(HTO_classification.global, 
                FM_doublets, 
                HTO_classification, 
                FM_sample)

classes_HTO_FM2 <- HTO_FM_compare2 %>% 
  group_by(HTO_classification, FM_sample) %>% 
  summarise(counts_combos = n()) %>% 
  group_by(HTO_classification) %>% 
  top_n(1, wt = counts_combos) %>% 
  filter(HTO_classification %in% hashtagsUsed)

# Add the patientID and previously defined freemuxlet samples
# This must be comparable to the individuals found by Freemuxlet
classes_HTO_FM_part<-classes_HTO_FM[c("HTO_classification", "FM_sample","PatientID")]
colnames(classes_HTO_FM_part)[2]<-c("FM_sample1")
classes_HTO_FM2<- left_join(x = classes_HTO_FM2, 
                           classes_HTO_FM_part, 
                           by = c("HTO_classification" = "HTO_classification"))
classes_HTO_FM2<-classes_HTO_FM2[classes_HTO_FM2$FM_sample==classes_HTO_FM2$FM_sample1,
                                 c("HTO_classification",
                                   "FM_sample",
                                   "counts_combos",
                                   "PatientID")]

write.table(x = classes_HTO_FM2,
          file = paste0(output_dir_results_hashing,"/HTO_FMO_counts2.txt"),
          sep = "\t",
          col.names = TRUE,
          quote = FALSE)

print(classes_HTO_FM)
```

```{r CompareDemuxFreemuxExtra2, eval = F}
##### This code chunk is not evaluated because every sample is unique in the pool #####

# Look for the same annotation based on freemuxlet and hashtags
HTO_FM_compare2 <- HTO_FM_compare2 %>%
  mutate(FMHTO_classification = case_when(
    # If doublet seen in either hashtags or freemuxlet --> Doublet
    HTO_classification.global == "Doublet" | FM_doublets == "double"~ "Doublet",
    # Hashtags match with freemuxlet --> correct hashtag
    HTO_classification == "Hashtag1" & FM_sample %in% classes_HTO_FM2[classes_HTO_FM2$HTO_classification == "Hashtag1", "FM_sample"] ~ "Hashtag1",
    HTO_classification == "Hashtag2" & FM_sample %in% classes_HTO_FM2[classes_HTO_FM2$HTO_classification == "Hashtag2", "FM_sample"] ~ "Hashtag2",
    HTO_classification == "Hashtag3" & FM_sample %in% classes_HTO_FM2[classes_HTO_FM2$HTO_classification == "Hashtag3", "FM_sample"] ~ "Hashtag3",
    HTO_classification == "Hashtag4" & FM_sample %in% classes_HTO_FM2[classes_HTO_FM2$HTO_classification == "Hashtag4", "FM_sample"] ~ "Hashtag4",
    HTO_classification == "Hashtag5" & FM_sample %in% classes_HTO_FM2[classes_HTO_FM2$HTO_classification == "Hashtag5", "FM_sample"] ~ "Hashtag5",
    HTO_classification == "Hashtag6" & FM_sample %in% classes_HTO_FM2[classes_HTO_FM2$HTO_classification == "Hashtag6", "FM_sample"] ~ "Hashtag6",
    HTO_classification == "Hashtag7" & FM_sample %in% classes_HTO_FM2[classes_HTO_FM2$HTO_classification == "Hashtag7", "FM_sample"] ~ "Hashtag7",
    HTO_classification == "Hashtag8" & FM_sample %in% classes_HTO_FM2[classes_HTO_FM2$HTO_classification == "Hashtag8", "FM_sample"] ~ "Hashtag8",
    # Both hashtag and freemuxlet can not identify sample
    HTO_classification.global == "Negative" & FM_doublets == "non-defined" ~ "TrueNegative",
    # Otherwise seen as unclassified
    TRUE~ "unclassified"
  ))

seuratObj_unident@meta.data[["HTO_FM_consensus"]] <- HTO_FM_compare2$FMHTO_classification
Idents(seuratObj_unident) <- "HTO_FM_consensus"

table(seuratObj_unident$HTO_FM_consensus)

qc_df<- qc_df %>% 
  bind_rows(lapply(hashtagsTotal,
                   function(i){
                     list(qc_stats=paste0(i, ' consensus 2 (extra)'),
                          values = table(seuratObj_unident$HTO_FM_consensus)[i])}))

```

## Final set of cells

### Extract singlets

```{r Singlets}
# Idents are changed to the hashtag of healthy control ID that were found
# We extract only singlets + doublets that contain cells of the same donor/sample (same SNP and hashtag)

# Singlets obtained after first demultiplexing step
seuratObj_singlet1 <- subset(seuratObj, idents = hashtagsUsed)

# Not all hashtags are still present in the seurat object
# seuratObj_singlet2 <- subset(seuratObj_unident, idents = grep("Hashtag",
#                                                               levels(seuratObj_unident),
#                                                               value=TRUE))

# Merge all the identified singlets together in one seurat object
# seuratObj_singlet <- merge(seuratObj_singlet1,
#                            seuratObj_singlet2)

seuratObj_singlet<-seuratObj_singlet1
seuratObj_singlet <- NormalizeData(seuratObj_singlet, 
                                   normalization.method = 'CLR', margin = 2)
seuratObj_singlet <- FindVariableFeatures(seuratObj_singlet, 
                                          selection.method = "mean.var.plot")
seuratObj_singlet  <- ScaleData(seuratObj_singlet, 
                                features = VariableFeatures(seuratObj_singlet))
seuratObj_singlet <- RunPCA(seuratObj_singlet, 
                            features = VariableFeatures(seuratObj_singlet))
```

### Filter out remaining doublets

Based on the demultiplexing performed above, two types of multiplets/doublets were removed:

-   Doublets containing cells from different donors based on the genotyping demultiplexing with freemuxlet
-   Doublet containing cells coded with different hashtags (originating from another donor or condition)

With the [Seurat function](https://github.com/chris-mcginnis-ucsf/DoubletFinder) DoubletFinder, it is also possible to filter out the doublets that remain: doublets from same donor (not identified via freemuxlet) and tagged with the same hashtag. This function uses the gene expression to identify if a droplet contains multiple cells and needs to be deleted from the data. The input arguments are: 

- seu = Seuratobject (after NormalizeData, FindVariableGenes, ScaleData, RunPCA) 
- PCs = number of statistically-significant principal components 
- pN = Number of generated artificial doublets, expressed as a proportion of the merged real-artificial data 
- pK = PC neighborhood size used to compute pANN (this value depends on the scRNA-seq dataset that is used and should be optimized per data file) 
- nExp = pANN threshold used to make final doublet/singlet predictions

```{r DoubletFinder, eval = FALSE}
# Define the expected number of doublet cells
nExp <- round(ncol(seuratObj_singlet) * 0.03)
print(nExp)
# Find doublets/multiplets
seuratObj_singlet <- doubletFinder_v3(seuratObj_singlet,
                                      pN = 0.25, pK = 0.09, 
                                      nExp = nExp, PCs = 1:10)
# Classifcation of cells based on DoubletFinder 
ClassName<-grep("DF.classification", seuratObj_singlet@meta.data%>%colnames(),
                value = TRUE)
Idents(seuratObj_singlet)<-ClassName
table(Idents(seuratObj_singlet))

qc_df <- qc_df %>%
  bind_rows(list(qc_stats = "doublets DoubletFinder",
                 values = table(Idents(seuratObj_singlet))["Doublet"]))

# Only use the 'real' singlets
seuratObj_singlet<-subset(seuratObj_singlet,
                          idents = 'Singlet')
```

```{r qcStatUpdate}
qc_df <- qc_df %>%
  bind_rows(list(qc_stats = "number of singlet cells",
                 values = dim(seuratObj_singlet)[2]))
qc_df <- qc_df %>%
  bind_rows(list(qc_stats = "percentage remaining cells",
                 values = qc_df$values[qc_df$qc_stats == "number of singlet cells"]/
                   qc_df$values[qc_df$qc_stats == "number of cells prefiltered"]*100))
```

### Add annotation

```{r Annotation}
annotation_cells <- left_join(as.data.frame(seuratObj_singlet$HTO_FM_consensus),
                              annotation_sample,
                              by=c("seuratObj_singlet$HTO_FM_consensus"="Hash"))

# Add annotation information in meta data of seurat object
seuratObj_singlet@meta.data[["PatientID"]]<- annotation_cells$PatientID
seuratObj_singlet@meta.data[["PIDID"]]<- annotation_cells$PIDID
seuratObj_singlet@meta.data[["Stimulation"]]<- annotation_cells$Stimulation
seuratObj_singlet@meta.data[["TimePoint"]]<- annotation_cells$TimePoint
seuratObj_singlet@meta.data[["Sorter"]]<- annotation_cells$Sorter
seuratObj_singlet@meta.data[["PatStim"]]<-paste(annotation_cells$PatientID,
                                                annotation_cells$Stimulation, 
                                                sep = "_")
```

### Check hashtags on singlets

```{r FeatureScatterSinglet}
DefaultAssay(seuratObj_singlet) <- 'HTO'
plot<-lapply(1:(length(hashtagsUsed)-1),
             function(i){
               # Adjust colors based on the hashtags of interest
               colors<-rep("grey", 8)
               names(colors)<-hashtagsUsed
               colors[i]<-colorsHashtags[i]
               colors[i+1]<-colorsHashtags[i+1]
               FeatureScatter(seuratObj_singlet, 
                                        feature1 = hashtagsUsed[i+1],
                                        feature2 = hashtagsUsed[i])+
                 scale_color_manual(values = alpha(colors, 0.7),
                                    breaks = c(hashtagsUsed[i], hashtagsUsed[i+1]),
                                    name = "")+
                 theme_celine})
ggarrange(plotlist = plot)
# Save as pdf
ggsave(filename = "Scatter_singlets_hashtags.pdf", 
       path= output_dir_results_hashing, 
       width = 20, height = 20, device = "pdf")
# Save as jpeg
ggsave(filename = "Scatter_singlets_hashtags.jpeg", 
       path= output_dir_results_hashing, 
       width = 20, height = 20, device = "jpeg")
DefaultAssay(seuratObj_singlet) <- 'RNA'
```

### Save seurat object

```{r SaveSeurat}
saveRDS(seuratObj_singlet, 
        file = paste0(output_dir_robjects,"/seuratObj_",
                      sampleName,"_singlets.rds"))
```

## Dimensionality reduction

### Visualisation TSNE

```{r TSNE}
# We select the top 10 PCs for clustering and tSNE based on PCElbowPlot
seuratObj_singlet <- FindNeighbors(seuratObj_singlet, 
                                   reduction = "pca", dims = 1:10)
seuratObj_singlet <- FindClusters(seuratObj_singlet, 
                                  resolution = 0.6, verbose = FALSE)
seuratObj_singlet <- RunTSNE(seuratObj_singlet, 
                             reduction = "pca", dims = 1:10)

# Projecting singlet identities on TSNE visualization
# Colors based HTO's of samples
DimPlot(seuratObj_singlet, 
        group.by = "HTO_classification")+
  ggtitle("Hashtags")+
  scale_color_manual(values = colorsHashtags,
                     breaks = hashtagsUsed)+
  theme_celine+
  guides(color = guide_legend(override.aes = list(size=1.5)))
# Save as pdf
ggsave(filename = "TSNE_HTOClassification.pdf", 
       path=output_dir_results_hashing, 
       width = 10, height = 10, device = "pdf") 
# Save as jpeg
ggsave(filename = "TSNE_HTOClassification.jpeg", 
       path=output_dir_results_hashing, 
       width = 10, height = 10, device = "jpeg")

# Colors based on Patients
DimPlot(seuratObj_singlet,
        group.by="PatientID")+
  ggtitle("Healthy control")+
  scale_color_manual(values = colorsPatients,
                     breaks = sort(unique(annotation_sample$PatientID)))+
  theme_celine+
  guides(color = guide_legend(override.aes = list(size=1.5)))

# Save as pdf
ggsave(filename = "TSNE_PatientClassification.pdf", 
       path=output_dir_results_hashing, 
       width = 10, height = 10, device = "pdf") 
# Save as jpeg
ggsave(filename = "TSNE_PatientClassification.jpeg", 
       path=output_dir_results_hashing, 
       width = 10, height = 10, device = "jpeg") 

# Colors based on Stimulation
DimPlot(seuratObj_singlet,
        group.by = "Stimulation")+
  ggtitle("Stimulation mix")+
  scale_color_manual(values = colorsStim,
                     breaks = sort(unique(annotation_sample$Stimulation)))+
  theme_celine+
  guides(color = guide_legend(override.aes = list(size=1.5)))
# Save as pdf
ggsave(filename = "TSNE_StimulationClassification.pdf", 
       path=output_dir_results_hashing, 
       width = 10, height = 10, device = "pdf") 
# Save as jpeg
ggsave(filename = "TSNE_StimulationClassification.jpeg", 
       path=output_dir_results_hashing, 
       width = 10, height = 10, device = "jpeg") 
```

### Add ADT data for singlets

#### Combination RNA and ADT based UMAP

WNN UMAP is Weighted Nearest Neighbor algorithm used in UMAP. This dimensionality reduction algorithm takes into account both the information on RNA and ADT (antibody-derived tags).

```{r wwnUMAP}
# Create directory
output_dir_results_multimodal <- paste0(output_dir_results,"/ADTintegration")
dir.create(output_dir_results_multimodal,recursive=TRUE)

# Look for cell both present in the gene expression and antibody capture data
# Focusing on the singlets data set
joint.cellnames <- intersect(rownames(seuratObj_singlet@meta.data), colnames(cellranger_data$`Antibody Capture`))

seuratObj_singlet[["ADT"]]  <- CreateAssayObject(counts = as.matrix(cellranger_data$`Antibody Capture`[,joint.cellnames]))
DefaultAssay(seuratObj_singlet) <- 'ADT'
VariableFeatures(seuratObj_singlet) <- rownames(seuratObj_singlet[["ADT"]])

seuratObj_singlet <- NormalizeData(seuratObj_singlet, 
                                   normalization.method = 'CLR', margin = 2) %>%
  ScaleData() %>% 
  RunPCA(reduction.name = 'apca')

seuratObj_singlet  <- FindMultiModalNeighbors(seuratObj_singlet, 
                                              reduction.list = list("pca", "apca"),
                                              dims.list = list(1:30, 1:18),
                                              modality.weight.name = "RNA.weight"
)

seuratObj_singlet <- RunUMAP(seuratObj_singlet, 
                             nn.name = "weighted.nn",
                             # Dimensionality reduction based on RNA and ADT counts (weighted)
                             reduction.name = "wnn.umap", 
                             reduction.key = "wnnUMAP_")
seuratObj_singlet <- FindClusters(seuratObj_singlet, 
                                  graph.name = "wsnn", 
                                  algorithm = 3, 
                                  resolution = 2, 
                                  verbose = FALSE)

# Visualisation UMAP
# Colors based on clusters identified
DimPlot(seuratObj_singlet, 
        reduction = 'wnn.umap', 
        label = TRUE, 
        repel = TRUE, 
        label.size = 2.5) +
  theme_celine+
  ggtitle("UMAP clusters RNA + ADT")+
  NoLegend()

# Save as pdf
ggsave(filename = "wnnUMAP_RNAADT_integration.pdf", 
       path=output_dir_results_multimodal, 
       width = 10, height = 10, device = "pdf")
# Save as jpeg
ggsave(filename = "wnnUMAP_RNAADT_integration.jpeg", 
       path=output_dir_results_multimodal, 
       width = 10, height = 10, device = "jpeg")

# Colors based HTO's of samples
DimPlot(seuratObj_singlet, 
        reduction = "wnn.umap",
        group.by = "HTO_classification")+
  ggtitle("Hashtags")+
  scale_color_manual(values = colorsHashtags,
                     breaks = hashtagsUsed)+
  theme_celine+
  guides(color = guide_legend(override.aes = list(size=1.5)))

# Save as pdf
ggsave(filename = "wnnUMAP_RNAADT_HTOClassification_integration.pdf", 
       path=output_dir_results_hashing, 
       width = 10, height = 10, device = "pdf") 
# Save as jpeg
ggsave(filename = "wnnUMAP_RNAADT_HTOClassification_integration.jpeg", 
       path=output_dir_results_hashing, 
       width = 10, height = 10, device = "jpeg") 

# Colors based on healthy controls identified
DimPlot(seuratObj_singlet,
        reduction = 'wnn.umap',
        group.by = "PatientID")+
  ggtitle("Healthy control")+
  scale_color_manual(values = colorsPatients,
                     breaks = sort(unique(seuratObj_singlet$PatientID)))+
  theme_celine+
  guides(color = guide_legend(override.aes = list(size=1.5)))

# Save as pdf
ggsave(filename = "wnnUMAP_RNAADT_Patient_integration.pdf", 
       path=output_dir_results_multimodal, 
       width = 10, height = 10, device = "pdf")
# Save as jpeg
ggsave(filename = "wnnUMAP_RNAADT_Patient_integration.jpeg", 
       path=output_dir_results_multimodal, 
       width = 10, height = 10, device = "jpeg")

# Colors based on healthy controls identified
DimPlot(seuratObj_singlet,
        reduction = 'wnn.umap',
        group.by = "Stimulation")+
  ggtitle("Stimulation mix")+
  scale_color_manual(values = colorsStim,
                     breaks = sort(unique(seuratObj_singlet$Stimulation)))+
  theme_celine+
  guides(color = guide_legend(override.aes = list(size=1.5)))

# Save as pdf
ggsave(filename = "wnnUMAP_RNAADT_Stimulation_integration.pdf", 
       path=output_dir_results_multimodal, 
       width = 10, height = 10, device = "pdf")
# Save as jpeg
ggsave(filename = "wnnUMAP_RNAADT_Stimulation_integration.jpeg", 
       path=output_dir_results_multimodal, 
       width = 10, height = 10, device = "jpeg")
```

#### RNA only based UMAP

```{r UMAP_RNA}
seuratObj_singlet <- RunUMAP(seuratObj_singlet, 
                             reduction = 'pca', 
                             dims = 1:30, 
                             # Dimensionality reduction only on RNA counts
                             assay = 'RNA',
                             reduction.name = 'rna.umap', 
                             reduction.key = 'rnaUMAP_')

# Colors based on clusters identified
DimPlot(seuratObj_singlet, 
        reduction = 'rna.umap',  
        label = TRUE,
        repel = TRUE, 
        label.size = 2.5) + 
  theme_celine+
  NoLegend()+
  ggtitle("UMAP clusters RNA")

# Save as pdf
ggsave(filename = "rnaUMAP_RNAADT_integration.pdf", 
       path=output_dir_results_multimodal, 
       width = 10, height = 10, device = "pdf")
# Save as jpeg
ggsave(filename = "rnaUMAP_RNAADT_integration.jpeg", 
       path=output_dir_results_multimodal, 
       width = 10, height = 10, device = "jpeg")

# Colors based HTO's of samples
DimPlot(seuratObj_singlet, 
        reduction = "rna.umap",
        group.by = "HTO_classification")+
  ggtitle("Hashtags")+
  scale_color_manual(values = colorsHashtags,
                     breaks = hashtagsUsed)+
  theme_celine+
  guides(color = guide_legend(override.aes = list(size=1.5)))
# Save as pdf
ggsave(filename = "rnaUMAP_RNAADT_HTOClassification_integration.pdf", 
       path=output_dir_results_hashing, 
       width = 10, height = 10, device = "pdf")
# Save as jpeg
ggsave(filename = "rnaUMAP_RNAADT_HTOClassification_integration.jpeg", 
       path=output_dir_results_hashing, 
       width = 10, height = 10, device = "jpeg")

# Colors based on healthy control
DimPlot(seuratObj_singlet,
        reduction = 'rna.umap',
        group.by = "PatientID")+
  ggtitle("Healthy control")+
  scale_color_manual(values = colorsPatients,
                     breaks = sort(unique(seuratObj_singlet$PatientID)))+
  theme_celine+
  guides(color = guide_legend(override.aes = list(size=1.5)))

# Save as pdf
ggsave(filename = "rnaUMAP_RNAADT_Patient_integration.pdf", 
       path=output_dir_results_multimodal, 
       width = 10, height = 10, device = "pdf")
# Save as jpeg
ggsave(filename = "rnaUMAP_RNAADT_Patient_integration.jpeg", 
       path=output_dir_results_multimodal, 
       width = 10, height = 10, device = "jpeg")

# Colors based on stimulation mix used
DimPlot(seuratObj_singlet,
        reduction = 'rna.umap',
        group.by = "Stimulation")+
  ggtitle("Stimulation mix")+
  scale_color_manual(values = colorsStim,
                     breaks = sort(unique(seuratObj_singlet$Stimulation)))+
  theme_celine+
  guides(color = guide_legend(override.aes = list(size=1.5)))

# Save as pdf 
ggsave(filename = "rnaUMAP_RNAADT_Stimulation_integration.pdf", 
       path=output_dir_results_multimodal, 
       width = 10, height = 10, device = "pdf")
# Save as jpeg
ggsave(filename = "rnaUMAP_RNAADT_Stimulation_integration.jpeg", 
       path=output_dir_results_multimodal, 
       width = 10, height = 10, device = "jpeg")
```

#### ADT only based UMAP

```{r UMAP_ADT}
seuratObj_singlet <- RunUMAP(seuratObj_singlet, 
                             reduction = 'apca', 
                             dims = 1:18, 
                             # Dimensionality reduction only based on ADT counts
                             assay = 'ADT',
                             reduction.name = 'adt.umap', 
                             reduction.key = 'adtUMAP_')

# Coloring based clusters identified
DimPlot(seuratObj_singlet, 
        reduction = 'adt.umap',  
        label = TRUE,
        repel = TRUE, 
        label.size = 2.5)+
  ggtitle("UMAP clusters ADT")+
  theme_celine+
  theme(plot.title=element_text(hjust=0.5))+
  NoLegend()

# Save as pdf
ggsave(filename = "adtUMAP_RNAADT_integration.pdf",
       path=output_dir_results_multimodal, 
       width = 10, height = 10, device = "pdf")
# Save as jpeg
ggsave(filename = "adtUMAP_RNAADT_integration.jpeg",
       path=output_dir_results_multimodal, 
       width = 10, height = 10, device = "jpeg")

# Colors based HTO's of samples
DimPlot(seuratObj_singlet, 
        reduction = "adt.umap",
        group.by = "HTO_classification")+
  ggtitle("Hashtags")+
  scale_color_manual(values = colorsHashtags,
                     breaks = hashtagsUsed)+
  theme_celine+
  guides(color = guide_legend(override.aes = list(size=1.5)))

# Save as pdf
ggsave(filename = "adtUMAP_RNAADT_HTOClassification_integration.pdf", 
       path=output_dir_results_hashing, 
       width = 10, height = 10, device = "pdf") 
# Save as jpeg
ggsave(filename = "adtUMAP_RNAADT_HTOClassification_integration.jpeg", 
       path=output_dir_results_hashing, 
       width = 10, height = 10, device = "jpeg") 

# Coloring based on healthy control
DimPlot(seuratObj_singlet,
        reduction = 'adt.umap',
        group.by = "PatientID")+
  ggtitle("Healthy control")+
  scale_color_manual(values = colorsPatients,
                     breaks = sort(unique(seuratObj_singlet$PatientID)))+
  theme_celine+
  guides(color = guide_legend(override.aes = list(size=1.5)))

# Save as pdf
ggsave(filename = "adtUMAP_RNAADT_Patient_integration.pdf", 
       path=output_dir_results_multimodal, 
       width = 10, height = 10, device = "pdf")
# Save as jpeg
ggsave(filename = "adtUMAP_RNAADT_Patient_integration.jpeg", 
       path=output_dir_results_multimodal, 
       width = 10, height = 10, device = "jpeg")

# Colors based on stimulation mix used
DimPlot(seuratObj_singlet,
        reduction = 'adt.umap',
        group.by = "Stimulation")+
  ggtitle("Stimulation mixes")+
  scale_color_manual(values = colorsStim,
                     breaks = sort(unique(seuratObj_singlet$Stimulation)))+
  theme_celine+
  guides(color = guide_legend(override.aes = list(size=1.5)))

# Save as pdf
ggsave(filename = "adtUMAP_RNAADT_Stimulation_integration.pdf", 
       path=output_dir_results_multimodal, 
       width = 10, height = 10, device = "pdf")
# Save as jpeg
ggsave(filename = "adtUMAP_RNAADT_Stimulation_integration.jpeg", 
       path=output_dir_results_multimodal, 
       width = 10, height = 10, device = "jpeg")
```

### Feature plots

```{r FeaturePlot}
# Select the ADT, not hashtags
featuresADT<-paste0(seuratObj_singlet@assays[['ADT']]@key, featuresADT)

FeaturePlot(seuratObj_singlet, features = featuresADT,
            reduction = 'wnn.umap', 
            max.cutoff = 2,
            cols = c("lightgrey","darkred"), ncol = 3)

# Save as pdf
ggsave(filename = "wnnUMAP_featureplot.pdf", 
       path=output_dir_results_multimodal, 
       width = 20, height = 40, device = "pdf")
# Save as jpeg
ggsave(filename = "wnnUMAP_featureplot.jpeg", 
       path=output_dir_results_multimodal, 
       width = 20, height = 40, device = "jpeg")

VlnPlot(seuratObj_singlet, 
        features = "RNA.weight", 
        group.by = 'seurat_clusters', 
        sort = TRUE, pt.size = 0.1) +
  theme_bar+
  ggtitle("RNA weights")+
  NoLegend()

# Save as pdf
ggsave(filename = "wnnUMAP_vlnplot.pdf", 
       path=output_dir_results_multimodal, 
       width = 20, device = "pdf")
# Save as jpeg
ggsave(filename = "wnnUMAP_vlnplot.jpeg", 
       path=output_dir_results_multimodal, 
       width = 20, device = "jpeg")

# Switch back to RNA assay
DefaultAssay(seuratObj_singlet) <- 'RNA'
```

### Save seurat object

```{r SaveSinglets}
saveRDS(seuratObj_singlet, 
        file = paste0(output_dir_robjects,"/seuratObj_",
                      sampleName,"_afterintegrationADT.rds"))
```

### Visualise UMAP

```{r QC_visualisation, message=FALSE, warning=FALSE,tidy=T}
# UMI plot

drawplot(seuratObj_singlet, 'wnn.umap', 'nCount_RNA')+
  ylab("wnnUMAP_2")+
  xlab("wnnUMAP_1")+
  ggtitle("RNA counts")+
  theme_celine+
  theme(legend.title = element_blank())
# Save as pdf
ggsave(file = paste0(output_dir_results_qc,
                     "/wnn.umap_nCountRNA.pdf"), 
       width = 10, height = 7, device = "pdf")
# Save as jpeg
ggsave(file = paste0(output_dir_results_qc,
                     "/wnn.umap_nCountRNA.jpeg"), 
       width = 10, height = 7, device = "jpeg")

# Mito plot
drawplot(seuratObj_singlet,'wnn.umap','percent_mito')+
  ylab("wnnUMAP_2")+
  xlab("wnnUMAP_1")+
  ggtitle("Percentage mitochondrial genes")+
  theme_celine+
  theme(legend.title = element_blank())

# Save as pdf
ggsave(file = paste0(output_dir_results_qc,
                     "/wnn.umap_percentmito.pdf"), 
       width = 10, height = 7, device = "pdf")
# Save as jpeg
ggsave(file = paste0(output_dir_results_qc,
                     "/wnn.umap_percentmito.jpeg"), 
       width = 10, height = 7, device = "jpeg")
```

## Visualisation based on expression data

### FeaturePlot - Plot marker gene expression

```{r FeatureVlnGene, message=FALSE, warning=FALSE,tidy=T}
output_dir_genes <- paste0(output_dir_results,"/PerGene")
marker_list <-read_csv(marker_genes_visualise, 
                       col_names=c("gene"))
dir.create(output_dir_genes,recursive=TRUE)

for (featured_gene in unlist(marker_list)) {
  print(GeneReductionPlot(seuratObj_singlet, 
                          featured_gene, 
                          reductionType="wnn.umap")+
          ylab("wnnUMAP_2")+
          xlab("wnnUMAP_1")+
          theme_celine)
  # Save as pdf
  ggsave(file = paste0(output_dir_genes,"/UMAP_", featured_gene, ".pdf"), 
         width = 20, height = 20, device = "pdf")
  # Save as jpeg
  ggsave(file = paste0(output_dir_genes,"/UMAP_", featured_gene, ".jpeg"), 
         width = 20, height = 20, device = "jpeg")
}

for (featured_gene in unlist(marker_list)) {
  print(GeneVlnPlot(seuratObj_singlet, 
                    featured_gene, 
                    reductionType="wnn.umap")+
          ylab("Gene counts")+
          theme_bar)
  # Save as pdf
  ggsave(file = paste0(output_dir_genes,"/Violin_", featured_gene, ".pdf"), 
         width = 20, device = "pdf")
  # Save as jpeg
  ggsave(file = paste0(output_dir_genes,"/Violin_", featured_gene, ".jpeg"), 
         width = 20, device = "jpeg")
}

```

### Calculate differentially expressed genes

```{r DEGenes, warning=FALSE, message=FALSE, tidy=TRUE}
### Find markers for every cluster compared to all remaining cells, report only the positive ones
allMarkers <- FindAllMarkers(seuratObj_singlet, 
                             min.pct = 0.10, 
                             min.diff.pct = 0.25, 
                             logfc.threshold = 0.30, 
                             return.thresh = 0.01, 
                             only.pos = TRUE, assay="RNA")
table(allMarkers$cluster)
saveRDS(allMarkers, file = paste0(output_dir_results,
                                  "/markersGenes_",sampleName,".rds"))

### Create list with markers
totalNrClusters<-max(as.numeric(names(table(allMarkers$cluster))))
totalNrClustersPlusOne<-totalNrClusters+1
markersList<-list()

for(i in 1:totalNrClustersPlusOne){
  clusterNr<-i-1

  tmp<-allMarkers[allMarkers$cluster==clusterNr,]
  tmp$score<-tmp$pct.1/(tmp$pct.2+0.01)*tmp$avg_log2FC

  markersList[[i]]<-tmp[order(tmp$score, decreasing=TRUE),]
}
names(markersList)<-paste0("cluster",0:totalNrClusters)

### Write to Excel
write.xlsx(markersList, file = paste0(output_dir_results,
                                      "/allClustersGenes_",sampleName,
                                      ".xlsx"))
writeGMTfromDEresults(markersList,
                      paste0(output_dir_results,"/allClustersGenes_",
                             sampleName,".gmt"))
```

### Calculate differentially expressed proteins

```{r DEProteins, warning=FALSE, message=FALSE, tidy=TRUE}

### Find markers for every cluster compared to all remaining cells, report only the positive ones
allMarkers <- FindAllMarkers(seuratObj_singlet, 
                             min.pct = 0.10, 
                             min.diff.pct = 0.25, 
                             logfc.threshold = 0.30, 
                             return.thresh = 0.01, 
                             only.pos = TRUE, assay = "ADT")
table(allMarkers$cluster)
saveRDS(allMarkers, file = paste0(output_dir_results,"/markersProteins_",
                                  sampleName,".rds"))

### Create list with markers
totalNrClusters<-max(as.numeric(names(table(allMarkers$cluster))))
totalNrClustersPlusOne<-totalNrClusters+1
markersList<-list()

for(i in 1:totalNrClustersPlusOne){
  clusterNr<-i-1
  tmp<-allMarkers[allMarkers$cluster==clusterNr,]
  tmp$score<-tmp$pct.1/(tmp$pct.2+0.01)*tmp$avg_log2FC
  markersList[[i]]<-tmp[order(tmp$score, decreasing=TRUE),]
}
names(markersList)<-paste0("cluster",0:totalNrClusters)

### Write to Excel
write.xlsx(markersList, file = paste0(output_dir_results,"/allClustersProteins_",
                                      sampleName,".xlsx"))
writeGMTfromDEresults(markersList, paste0(output_dir_results,"/allClustersProteins_",
                                          sampleName,".gmt"))
```

### Cell Cycle Regression

```{r CellCycleReg}
s.genes <- cc.genes$s.genes[cc.genes$s.genes %in% 
                              rownames(seuratObj_singlet[["RNA"]])]
g2m.genes <- cc.genes$g2m.genes[cc.genes$g2m.genes %in%
                                  rownames(seuratObj_singlet[["RNA"]])]
seuratObj_singlet <- CellCycleScoring(seuratObj_singlet, 
                                      s.features = s.genes, 
                                      g2m.features = g2m.genes, set.ident = TRUE)
DimPlot(seuratObj_singlet, reduction= "wnn.umap")+
  scale_color_manual(values = c("#00D9C0", "#C592DD", "#F0B67F"))+
  theme_celine+
  guides(color = guide_legend(override.aes = list(size=1.5)))+
  ggtitle("Cell cycle regression")
# Save as pdf
ggsave(file = paste0(output_dir_genes,"/Cell_cycle_wnnumap.pdf"), 
       width = 20, device = "pdf")
# Save as jpeg
ggsave(file = paste0(output_dir_genes,"/Cell_cycle_wnnumap.jpeg"), 
       width = 20, device = "jpeg")
```

## A multimodal reference dataset

A reference data set, loades as a new seurat object, is used to label the cells of our pooled samples. Both main cell types (e.g. B cells, T cells) and subcell types (e.g. CD8+ T cells) are present in the data.

```{r MultimodalRef}
output_dir_labelcells <- paste0(output_dir_results,"/Cell_annotation")
dir.create(output_dir_labelcells,recursive=TRUE)
reference <- LoadH5Seurat("/data/gent/vo/000/gvo00027/SingleCell10X/seuratintegrate/multi.h5seurat")
DimPlot(object = reference, reduction = "wnn.umap", 
        group.by = "celltype.l1", label = TRUE, 
        label.size = 3, repel = TRUE) +
  theme_celine+
  NoLegend()+
  ggtitle("Cell Type 1")+
  ylab('wwnUMAP_2')+
  xlab('wnnUMAP_1')

DimPlot(object = reference, reduction = "wnn.umap", 
      group.by = "celltype.l2", label = TRUE, 
      label.size = 3, repel = TRUE) + 
  theme_celine+
  NoLegend()+
  ggtitle("Cell Type 2")+
  ylab('wwnUMAP_2')+
  xlab('wnnUMAP_1')
```

```{r MappingToRef}
seuratObj_singlet <- SCTransform(seuratObj_singlet, verbose = FALSE)

anchors <- FindTransferAnchors(
  reference = reference,
  query = seuratObj_singlet,
  normalization.method = "SCT",
  reference.reduction = "spca",
  dims = 1:50
)

seuratObj_singlet <- MapQuery(
  anchorset = anchors,
  query = seuratObj_singlet,
  reference = reference,
  refdata = list(
    celltype.l1 = "celltype.l1",
    celltype.l2 = "celltype.l2",
    predicted_ADT = "ADT"
  ),
  reference.reduction = "spca",
  reduction.model = "wnn.umap"
)



p1 <- DimPlot(seuratObj_singlet, reduction = "ref.umap", 
             group.by = "predicted.celltype.l1", label = TRUE, 
             label.size = 3, repel = TRUE) + 
  theme_celine+
  NoLegend()+
  ggtitle("Cell Type 1")
p2 <- DimPlot(seuratObj_singlet, reduction = "ref.umap", 
             group.by = "predicted.celltype.l2", label = TRUE, 
             label.size = 3 ,repel = TRUE) + 
  theme_celine+
  NoLegend()+
  ggtitle("Cell Type 2")

p3 <- DimPlot(seuratObj_singlet, reduction = "ref.umap",
        group.by = "Stimulation")+
  ggtitle("Stimulation mix")+
  scale_color_manual(values = colorsStim,
                     breaks = sort(unique(annotation_sample$Stimulation)))+
  theme_celine+
  guides(color = guide_legend(override.aes = list(size=1.5)))

ggarrange(plotlist = list(p1, p3, p2),
          widths = c(1,1.25))

# Save as pdf
ggsave(file = paste0(output_dir_labelcells,"/Celltypepred_refumap.pdf"), 
       height = 20, width = 40, device = "pdf")
# Save as jpeg
ggsave(file = paste0(output_dir_labelcells,"/Celltypepred_refumap.jpeg"), 
       height = 20, width = 40, device = "jpeg")

p1 <- DimPlot(seuratObj_singlet, reduction = "wnn.umap", 
             group.by = "predicted.celltype.l1", label = TRUE, 
             label.size = 3, repel = TRUE) + 
  theme_celine+
  NoLegend()+
  ggtitle("Cell Type 1")

p2 <- DimPlot(seuratObj_singlet, reduction = "wnn.umap", 
             group.by = "predicted.celltype.l2", label = TRUE, 
             label.size = 3 ,repel = TRUE) + 
  theme_celine+
  NoLegend()+
  ggtitle("Cell Type 2")

p3 <- DimPlot(seuratObj_singlet, reduction = "wnn.umap",
        group.by = "Stimulation")+
  ggtitle("Stimulation mix")+
  scale_color_manual(values = colorsStim,
                     breaks = sort(unique(annotation_sample$Stimulation)))+
  theme_celine+
  guides(color = guide_legend(override.aes = list(size=1.5)))

ggarrange(plotlist = list(p1, p3, p2),
          widths = c(1,1.25))

# Save as pdf
ggsave(file = paste0(output_dir_labelcells,"/Celltypepred_wnnumap.pdf"), 
       height = 10, width = 15, device = "pdf")
# Save asjpeg
ggsave(file = paste0(output_dir_labelcells,"/Celltypepred_wnnumap.jpeg"), 
       height = 10, width = 15, device = "jpeg")

# Check clustering on ADT and RNA UMAPS
p1 <- DimPlot(seuratObj_singlet, reduction = "rna.umap", 
             group.by = "predicted.celltype.l1", label = TRUE, 
             label.size = 3, repel = TRUE) + 
  theme_celine+
  NoLegend()+
  scale_color_manual(values = colorsCellType)+
  labs(title = "Cell Type 1",
       subtitle = "RNA only based UMAP")

p2 <- DimPlot(seuratObj_singlet, reduction = "adt.umap", 
             group.by = "predicted.celltype.l1", label = TRUE, 
             label.size = 3 ,repel = TRUE) + 
  theme_celine+
  NoLegend()+
  scale_color_manual(values = colorsCellType)+
  labs(title = "Cell Type 1",
       subtitle = "ADT only based UMAP")

ggarrange(plotlist = list(p1, p2),
          widths = c(1,1.25))
  
# Save as pdf
ggsave(file = paste0(output_dir_labelcells,"/Celltypepred_rnaadtumap.pdf"), 
       height = 10, width = 15, device = "pdf")
# Save asjpeg
ggsave(file = paste0(output_dir_labelcells,"/Celltypepred_rnaadtumap.jpeg"), 
       height = 10, width = 15, device = "jpeg")

DefaultAssay(seuratObj_singlet) <- "prediction.score.celltype.l2"

for (celltype in unique(rownames(seuratObj_singlet[["prediction.score.celltype.l2"]]))){
  FeaturePlot(seuratObj_singlet, features = celltype,  
            reduction = "ref.umap", cols = c("lightgrey", "darkred"), 
            split.by = "PatientID") & 
    theme(plot.title = element_text(size = 10,
                                    hjust = 0.5))+
    theme(plot.title = element_text(size = 12,color = "gray10"),
          plot.subtitle = element_text(size=8, color= "gray 30"),
          axis.text = element_text(size=8,color = "gray10"),
          axis.title = element_text(size=10,color = "gray10"),
          legend.text = element_text(size=8,color = "gray10"),
          legend.title = element_text(size=10,color = "gray10"),
          axis.line = element_line(color="gray50"),
          axis.ticks = element_line(color="gray50"))
  # Save as pdf
  ggsave(file = paste0(output_dir_labelcells,"/",celltype,"_refumap.pdf"), 
         height = 10, width = 15, device = "pdf")
  # Save as jpeg
  ggsave(file = paste0(output_dir_labelcells,"/",celltype,"_refumap.jpeg"), 
         height = 10, width = 15, device = "jpeg")
  
  FeaturePlot(seuratObj_singlet, features = celltype,  
              reduction = "wnn.umap", 
              cols = c("lightgrey", "darkred")) & 
    theme(plot.title = element_text(size = 10)) & 
    theme_celine
  
  # Save as pdf
  ggsave(file = paste0(output_dir_labelcells,"/", celltype, "_wnnumap.pdf"), 
         height = 15, width = 15, device = "pdf")
  # Save as jpeg
  ggsave(file = paste0(output_dir_labelcells,"/", celltype, "_wnnumap.jpeg"), 
         height = 15, width = 15, device = "jpeg")
}

```

The plot above shows the cells present in the pooled sample and mapped onto the reference dataset. This makes it possible to label the cells based on 
- Expressed genes: identified using the transcritpomic data of the RNA molecules in the cells 
- Expressed surface markers: identified by the oligonucleotide tagged antibodies in the CITE-seq panel

```{r VlnPlotSCT}
DefaultAssay(seuratObj_singlet) <- "SCT"
Idents(seuratObj_singlet) <- 'predicted.celltype.l2'
VlnPlot(seuratObj_singlet, features = "RNA.weight", 
        group.by = 'predicted.celltype.l2', sort = TRUE, pt.size = 0.1) +
  theme_bar+
  ggtitle("RNA weights")+
  NoLegend()+
  xlab("Cell Type 2")
  
# Save as pdf
ggsave(file = "RNAWeights_Labels.pdf", 
       height = 15, width = 15, device = "pdf")
# Save as jpeg
ggsave(file = "RNAWeights_Labels.jpeg", 
       height = 15, width = 15, device = "jpeg")
```

### Save seurat Object

```{r SaveFinalSeurat, message=FALSE, warning=FALSE,tidy=T}
saveRDS(seuratObj_singlet, file = paste0(output_dir_robjects,"/seuratObj_",
                                         sampleName,"_labeled.rds"))
```

## Fraction of cell types

```{r}
# Function table counts the number of cells per cell type
CellType1<-data.frame(cell = names(table(seuratObj_singlet$predicted.celltype.l1)),
                      cellNumber = table(seuratObj_singlet$predicted.celltype.l1)%>%as.numeric, 
                      cellPercentage = as.numeric(table(seuratObj_singlet$predicted.celltype.l1)/(dim(seuratObj_singlet)[2])*100),
                      cellType = "Type1")
rownames(CellType1)<-names(table(seuratObj_singlet$predicted.celltype.l1))
print(CellType1)

CellType2<-data.frame(cell = names(table(seuratObj_singlet$predicted.celltype.l2)),
                      cellNumber = table(seuratObj_singlet$predicted.celltype.l2)%>%as.numeric, 
                      cellPercentage = as.numeric(table(seuratObj_singlet$predicted.celltype.l2)/(dim(seuratObj_singlet)[2])*100),
                      cellType = "Type2")
print(CellType2)

CellTypes<-rbind(CellType1, CellType2)
write_tsv(CellTypes, paste0(output_dir_results, "/CellTypes.tsv"))

```

## QC report

```{r WriteQCReport}
write_tsv(qc_df,paste0(output_dir_results,"/qc_stats.tsv"))
```
