---
title: "012_MiloR_DA_FilteredBCells"
author: "Tine D'hamers"
date: "2025-06-27"
output: html_document
params:
  path_Results: "/kyukon/scratch/gent/vo/002/gvo00224/TOBI/Projects/EProfile/Results/Exp_202211_HCSamples_CEV017-22/002_NoBaseline/202506_Filtered_BCells"
  date: "20250625"
  date2: "20250820"
  date3: "20250926"
---

# Preparatory steps

## Load necessary libraries

```{r LoadLibraries}
# Single cell data manipulation
library(Seurat)
# library(scDblFinder)
# General data manipulation
library(tidyr)
library(dplyr)
library(tibble)
library(xlsx)
library(stringr)
# Plot generation
library(ggplot2)
library(ggpubr)
library(ggh4x)
library(ggbeeswarm)
library(RColorBrewer)
library(rstatix)
library(ComplexHeatmap)
library(circlize)
library(factoextra)
library(Rtsne)
library(ggridges)
library(uwot)
library(ggnewscale)
library(patchwork)
# Differential abundance analysis
library(miloR, 
        lib.loc = "/kyukon/data/gent/vo/002/gvo00224/TOBI/Projects/SingleCellPipeline/resultsCeline/seuratintegrate/Rpackages/R_4.2.1-foss-2022a")
# library(scater)
library(openxlsx)
```

## Load necessary function and aesthetics
```{r Functions_Aesthetics}
# Themes
source("../../PlotThemes_sc_CVID.R")
colorsStim<-c(colorsStim, c("CD3/CD28" = colorsStim[["CD3/CD28_beads"]]))

options(bitmapType='cairo')

source("../../Functions_sc_CVID.R")

# Coloring for the DE genes
colDirection<-c("DOWN" = "darkred",
                "UP" = "darkgreen",
                "MIXED" = "#D5BBB1",
                "NS" = "grey80")
grid.grey = "grey95"

# Variable storage
FDR_thres<-0.1
logFC_thres<-2
```

## Load dataset

Differential abundance analysis (DA) is performed on the filtered subset of the B cells stimulated for both 8h and 30h. These datasets are generated in the script `012_BCellFiltering_DEGenes.Rmd`.

```{r LoadDatasets}
IntData.B_8h.FilCus <- readRDS(file = file.path(params$path_Results,
                                                 "SeuratObjects",
                                                 paste0(params$date, "_seuratObj_NoBase_Integrated_Bclusters_Filtered_8h.rds")))
IntData.B_30h.FilCus <- readRDS(file = file.path(params$path_Results,
                                                       "SeuratObjects",
                                                       paste0(params$date, "_seuratObj_NoBase_Integrated_Bclusters_Filtered_30h.rds")))
```

# Differential abundance analysis with MiloR

## 8 hours of stimulation

### Data analysis using MiloR

```{r MiloR_DAanalysis_8h}
set.seed(2024)
# Generation of the milo object and graph
# Conform the vignette of MiloR
milo.B_8h.FilClus<-MiloDA(object = IntData.B_8h.FilCus, assay = "RNA", k = 50, d = 20, prop = 0.15,
                          # As recommended by Noah, perform KNN graph generation based on the dimensionality reduction technique
                          # containing the most information
                          RedDim = "B_filtered_CCAInt.rna", sampleID = "PatStim", 
                          design = c("PatStim", "Stimulation"), overlap = 5)
# Save the object for later use
saveRDS(milo.B_8h.FilClus, 
        file = file.path(params$path_Results, "MiloR",
                         paste(params$prefix_output, "MiloRobject_FilteredB_8h.rds", sep = "_")))

milo.B_8h.FilClus$size_histogram
# Set k such that you have a minimum of Sx5 counts per nhood, where S=total number of samples

# Statistical testing
milo.B_8h.FilClus$milo_design$Stimulation <- milo.B_8h.FilClus$milo_design$Stimulation%>%
  gsub(pattern = "CD3/CD28_beads", replacement = "antiCD3CD28")
contrasts<-c("StimulationIFNgamma - StimulationUnstimulated", "StimulationMegaCD40L - StimulationUnstimulated", 
             "StimulationantiCD3CD28 - StimulationUnstimulated", "StimulationODN2006 - StimulationUnstimulated")

# Extract results of differential abundance analysis
# Note: This part is a bit different from what is performed in 006_NoBase_visualization_doubletCheck_ClusterAnalysis_IntegrationV5.Rmd script
DA.B_8h.FilClus<-testNhoods(x = milo.B_8h.FilClus$milo_graph, design = ~0+Stimulation, design.df = milo.B_8h.FilClus$milo_design,
                            # Correct for multiple testing using a spatial FDR correction
                            fdr.weighting = "graph-overlap", 
                            model.contrasts = contrasts, reduced.dim = "B_filtered_CCAInt.rna"%>%toupper())

# Rename the columns of the DA results
names(DA.B_8h.FilClus)
DA.B_8h.FilClus<- DA.B_8h.FilClus%>%
  pivot_longer(cols = starts_with(("logFC")),
               names_to = "Stimulation",
               values_to = "logFC")%>%
  mutate(Stimulation = gsub(pattern = "logFC.Stimulation", 
                            replacement = "",
                            x = Stimulation),
         Stimulation = gsub(pattern = "...S.*", 
                            replacement = "",
                            x = Stimulation))
``` 

```{r ReadIn_MiloObjects_8h, eval=FALSE}
milo.B_8h.FilClus<-readRDS(file = file.path(params$path_Results, "MiloR",
                                            paste(params$prefix_output, "MiloRobject_FilteredB_8h.rds", sep = "_")))
```

### Neighbourhood graph plotting

```{r MiloDA_BCells_Filtered_Visualisation_8h, fig.height=5, fig.width=12}
# Visualization 
# Generation of plots
pIFN<-plotNhoodGraphDA(x = milo.B_8h.FilClus$milo_graph, alpha = 0.1,
                   milo_res = DA.B_8h.FilClus%>%filter(Stimulation == "IFNgamma"), layout = "B_filtered_CCAInt.umap.rna"%>%toupper(),
                   size_range = c(0.01,1.5), node_stroke = 0.2)+
    labs(subtitle = paste0("IFN","\U03B3","/R-848"))+
    guides(size=guide_legend("Size of Nhoods"), 
           edge_width = guide_legend("Overlap Nhoods"), 
           fill = guide_colourbar("LogFC", theme = theme(legend.key.width  = unit(0.5, "lines"))))

pCD40L<-plotNhoodGraphDA(x = milo.B_8h.FilClus$milo_graph, alpha = 0.1,
                   milo_res = DA.B_8h.FilClus%>%filter(Stimulation == "MegaCD40L"), layout = "B_filtered_CCAInt.umap.rna"%>%toupper(),
                   size_range = c(0.01,1.5), node_stroke = 0.2)+
    labs(subtitle = "MegaCD40L")+
    guides(size=guide_legend("Size of Nhoods"), 
           edge_width = guide_legend("Overlap Nhoods"), 
           fill = guide_colourbar("LogFC", theme = theme(legend.key.width  = unit(0.5, "lines"))))

pCD3CD28<-plotNhoodGraphDA(x = milo.B_8h.FilClus$milo_graph, alpha = 0.1,
                   milo_res = DA.B_8h.FilClus%>%filter(Stimulation == "antiCD3CD28"), layout = "B_filtered_CCAInt.umap.rna"%>%toupper(),
                   size_range = c(0.01,1.5), node_stroke = 0.2)+
    labs(subtitle = "Anti-CD3/CD28")+
    guides(size=guide_legend("Size of Nhoods"), 
           edge_width = guide_legend("Overlap Nhoods"), 
           fill = guide_colourbar("LogFC", theme = theme(legend.key.width  = unit(0.5, "lines"))))

pODN<-plotNhoodGraphDA(x = milo.B_8h.FilClus$milo_graph, alpha = 0.1,
                   milo_res = DA.B_8h.FilClus%>%filter(Stimulation == "ODN2006"), layout = "B_filtered_CCAInt.umap.rna"%>%toupper(),
                   size_range = c(0.01,1.5), node_stroke = 0.2)+
    labs(subtitle = "ODN2006")+
    guides(size=guide_legend("Size of Nhoods"), 
           edge_width = guide_legend("Overlap Nhoods"), 
           fill = guide_colourbar("LogFC", theme = theme(legend.key.width  = unit(0.5, "lines"))))

pUMAP<-DimPlot(object = IntData.B_8h.FilCus,
        reduction = "B_filtered_CCAInt.umap.rna", group.by = "Stimulation",
        pt.size = 0.001)+
    scale_color_manual(values = colorsStim)+
    labs(title=NULL, x = "UMAP1", y = "UMAP2", color = "Stimulation")+
    theme(axis.line = element_blank(), axis.text = element_blank(),
          axis.ticks = element_blank(), axis.title = element_blank())+
    guides(color = guide_legend(override.aes = list(size=1))) 
# Via patchwork, arranging of plots
pUMAP + plot_spacer() + ( (pIFN + plot_spacer() + pODN + plot_layout(widths = c(1,0,1))) / 
                            (pCD3CD28 + plot_spacer() + pCD40L + plot_layout(widths = c(1,0,1))))+
  plot_layout(guides = "collect", widths = c(1,0,1)) & 
  scale_fill_gradientn(limits = range(min(DA.B_8h.FilClus$logFC), max(DA.B_8h.FilClus$logFC)),
                       colours = c("#8C2F39", "white", "#2D0650"),
                       name = "LogFC") &
  labs(size = "Size of Nhood") &
  guides(fill = guide_colourbar("LogFC", theme = theme(legend.key.width  = unit(0.5, "lines")))) &
  theme(plot.subtitle = element_text(size = 10, color = "gray30"),
        legend.text =element_text(margin = margin(t = 1,l = 3), size = 8, color = "gray30"),
        legend.title = element_text(size = 10, color = "gray30"),
        plot.margin = margin(1,0,0,0)) &
  theme(legend.key.size = unit(1, "lines"),
        legend.spacing.y = unit(0.5, "lines"),
        legend.spacing.x = unit(0.5, "lines"),
        legend.margin = margin(1,0,0,1), 
        legend.key.spacing = unit(0.3, "lines"))
  
ggsave(filename = file.path(params$path_Results,
                            "MiloR",
                            paste(params$prefix_output, "DA_FilteredB_8h_RefUnstim.pdf",
                                  sep = "_")),
       bg = "white", width = 25, height = 14, device = "pdf", units = "cm",
       dpi = 400)

# Cleaning workspace
rm(pIFN, pODN,pCD3CD28, pCD40L, pUMAP)
```

### Beeswarm plots based on original clusters

```{r OriginalClustering_Nhoods_8h, fig.height=10, fig.width=10}
# Annotation of the Nhoods per stimulation linking to the original clusters
DA.B_8h.FilClus_ClustIFNG <- annotateNhoods(milo.B_8h.FilClus$milo_graph, 
                                            # Only possible per stimulation
                                            DA.B_8h.FilClus%>%
                                              filter(Stimulation == "IFNgamma"), 
                                            # Linking back to the clusters obtained after filtering and CCA integration
                                            coldata_col = "B_filtered_CCAInt_clusters")
# Convert to dataframe to use as input for the plotDAbeeswarm function
DA.B_8h.FilClus_ClustIFNG <- as.data.frame(DA.B_8h.FilClus_ClustIFNG)

# Check the outcome after Nhood annotation
## Interesting: the fractions of linked clusters to the Nhoods is given as well
head(DA.B_8h.FilClus_ClustIFNG)

# Annotation of the Nhoods per stimulation linking to the original clusters
plotBee_ClusAnnot <- lapply(unique(DA.B_8h.FilClus$Stimulation),
                                     FUN = function(layer){
                                       # Annotation of the Nhoods per stimulation linking to the original clusters
                                       DA.B_8h.FilClus_ClusAnnot.tmp <- annotateNhoods(milo.B_8h.FilClus$milo_graph,
                                                                                       # Only possible per stimulation
                                                                                      DA.B_8h.FilClus%>%
                                                                                        filter(Stimulation == layer), 
                                                                                      # Linking back to the clusters obtained after filtering and CCA integration
                                                                                      coldata_col = "B_filtered_CCAInt_clusters")
                                       # Highligh neighbourhoods that are a mix of clusters
                                       DA.B_8h.FilClus_ClusAnnot.tmp$B_filtered_CCAInt_clusters <- ifelse(
                                         DA.B_8h.FilClus_ClusAnnot.tmp$B_filtered_CCAInt_clusters_fraction < 0.7, "Mixed", 
                                         DA.B_8h.FilClus_ClusAnnot.tmp$B_filtered_CCAInt_clusters)
                                       # Convert to dataframe to use as input for the plotDAbeeswarm function
                                       DA.B_8h.FilClus_ClusAnnot.tmp <- as.data.frame(DA.B_8h.FilClus_ClusAnnot.tmp)%>%
                                         # Ordering the clusters to visualize properly in the graph
                                         mutate(B_filtered_CCAInt_clusters = factor(B_filtered_CCAInt_clusters,
                                                                                    levels = unique(DA.B_8h.FilClus_ClusAnnot.tmp$B_filtered_CCAInt_clusters) %>%
                                                                                      gtools::mixedsort(decreasing = T),
                                                                                    ordered = T))
                                       
                                       # Check the outcome
                                       head(DA.B_8h.FilClus_ClusAnnot.tmp)
                                       return(list(plotDAbeeswarm(DA.B_8h.FilClus_ClusAnnot.tmp,
                                                      group.by = "B_filtered_CCAInt_clusters", alpha = FDR_thres)+
                                         themeG+
                                         theme(panel.grid.major.y = element_line(color = adjustcolor( "lightgrey", alpha.f = 0.1),
                                                                                 linewidth = 1)),
                                         DA.B_8h.FilClus_ClusAnnot.tmp))})
# Change names of the list
names(plotBee_ClusAnnot) <- unique(DA.B_8h.FilClus$Stimulation)
# Change names of the lists in the list, aka data and plot layer
for (layer in names(plotBee_ClusAnnot)){
  names(plotBee_ClusAnnot[[layer]])<-c("plot", "data")
}

# Plotting result
(plotBee_ClusAnnot$IFNgamma$plot+labs(subtitle = "IFN") + plotBee_ClusAnnot$ODN2006$plot + labs(subtitle = "ODN2006")) /
  (plotBee_ClusAnnot$antiCD3CD28$plot + labs(subtitle = "anti-CD3/CD28") + plotBee_ClusAnnot$MegaCD40L$plot +labs(subtitle = "CD40L"))+
  plot_annotation("8h") & labs(x = "Original clusters") & ylim(-10,10)
```

### Selection of interesting neighbourhoods - soft cut-offs per stimulation

#### Unstimulated neighbourhoods

```{r NBH_Selection_8h_UnstimulatedCells}
CutOff_Frac<-0.5
# Join all data of the beesswarms over the stimulations
plotBee_ClusAnnot.data<-rbind(plotBee_ClusAnnot$IFNgamma$data,
                              plotBee_ClusAnnot$MegaCD40L$data,
                              plotBee_ClusAnnot$antiCD3CD28$data,
                              plotBee_ClusAnnot$ODN2006$data)

### Unstimulated clusters/neighbourhood
# Check minimum and maximum LogFC
NBH_Selection_Unstim.CutOffs <- plotBee_ClusAnnot.data%>%
  filter(B_filtered_CCAInt_clusters == 0)%>%
  group_by(Stimulation, B_filtered_CCAInt_clusters)%>%
  summarise(Min = min(logFC),
            Max = max(logFC),
            CutOff = min(logFC)*(1-CutOff_Frac))

NBH_Selection_Unstim.CutOffs
```
```{r NBH_Overlap_Unstimulated}
# Filtering based on the cutoffs
NBH_Selection_Unstim<-left_join(plotBee_ClusAnnot.data,
                                NBH_Selection_Unstim.CutOffs%>%
                                  select(Stimulation, B_filtered_CCAInt_clusters, CutOff))%>%
  filter(B_filtered_CCAInt_clusters == 0,
         logFC < CutOff)

# Prepare UpsetPlotData
NBH_Selection_Unstim.CombMat<-lapply(unique(NBH_Selection_Unstim$Stimulation),
                                     FUN = function(stim){(NBH_Selection_Unstim%>%filter(Stimulation == stim))[["Nhood"]]})
names(NBH_Selection_Unstim.CombMat)<-unique(NBH_Selection_Unstim$Stimulation)
NBH_Selection_Unstim.CombMat<-make_comb_mat(NBH_Selection_Unstim.CombMat, mode = "distinct")
# Plot upset plot
UpSet(NBH_Selection_Unstim.CombMat,
      comb_order = order(comb_size(NBH_Selection_Unstim.CombMat), decreasing = T),
      top_annotation = HeatmapAnnotation(
                "Intersection size" = anno_barplot(comb_size(NBH_Selection_Unstim.CombMat),
                                                   border = FALSE,
                                                   gp = gpar(fill = "#595B5C",
                                                             lty = "blank"),
                                                   add_numbers = T,
                                                   height = unit(4, "cm")),
                annotation_name_side = "left",
                annotation_name_rot = 90),
      right_annotation = rowAnnotation(
                "DE genes" = anno_barplot(set_size(NBH_Selection_Unstim.CombMat),
                                          border = FALSE,
                                          gp = gpar(fill = "#595B5C",
                                                    lty = "blank"),
                                          add_numbers = T,
                                          width = unit(4, "cm"))))

# Selected NBHs
NBH_Selection_Unstim.Numbers<-sapply(X = c("1110", "1111", "1011", "1101"),
                                     FUN = function(set){extract_comb(NBH_Selection_Unstim.CombMat, set)})%>%
  unlist()
length(NBH_Selection_Unstim.Numbers)
```

#### Stimulated neighbourhoods

```{r NBH_Selection_8h_Stimulations, fig.height=10, fig.width=10}
### IFNgamma and ODN2006
# Number of clusters plotted in the Beeswarm plot
NumClus<-length(levels(plotBee_ClusAnnot.data$B_filtered_CCAInt_clusters))
# Obtain the cutoff values per cluster and per Stimulation
# Only focusing on IFNgamma and ODN2006 stimulation
NBH_Selection_IFNgamma_ODN2006.CutOffs <- plotBee_ClusAnnot.data%>%
  filter(B_filtered_CCAInt_clusters %in% c(1,3,6,12), 
         Stimulation %in% c("IFNgamma", "ODN2006"))%>%
  group_by(Stimulation, B_filtered_CCAInt_clusters)%>%
  summarise(Min = min(logFC),
            Max = max(logFC),
            CutOff = max(logFC)*(1-CutOff_Frac))%>%
  mutate(positionHigh = NumClus - as.numeric(as.character(B_filtered_CCAInt_clusters)) + 0.5,
         positionLow = NumClus - as.numeric(as.character(B_filtered_CCAInt_clusters)) - 0.5)
# Check the obtained cutoffs
NBH_Selection_IFNgamma_ODN2006.CutOffs%>%
  arrange(B_filtered_CCAInt_clusters)

### MegaCD40L and Anti-CD3/C29
# Obtain the cutoff values per cluster and per Stimulation
# Only focusing on IFNgamma and ODN2006 stimulation
NBH_Selection_CD40L_CD3CD28.CutOffs <- plotBee_ClusAnnot.data%>%
  filter(B_filtered_CCAInt_clusters %in% c(2,5,7), 
         Stimulation %in% c("MegaCD40L", "antiCD3CD28"))%>%
  group_by(Stimulation, B_filtered_CCAInt_clusters)%>%
  summarise(Min = min(logFC),
            Max = max(logFC),
            CutOff = max(logFC)*(1-CutOff_Frac))%>%
  mutate(positionHigh = NumClus - as.numeric(as.character(B_filtered_CCAInt_clusters)) + 0.5,
         positionLow = NumClus - as.numeric(as.character(B_filtered_CCAInt_clusters)) - 0.5)
# Check the obtained cutoffs
NBH_Selection_CD40L_CD3CD28.CutOffs%>%
  arrange(B_filtered_CCAInt_clusters)
```

#### Merging of cut-off values and selected NBH indication


```{r NBH_Selection_8h}
NBH_Selection.CutOffs<-rbind(NBH_Selection_Unstim.CutOffs,
                             NBH_Selection_IFNgamma_ODN2006.CutOffs,
                             NBH_Selection_CD40L_CD3CD28.CutOffs)

# Add cutoffs to the data
plotBee_ClusAnnot.data <- left_join(plotBee_ClusAnnot.data,
                                    NBH_Selection.CutOffs%>%
                                      select(CutOff, Stimulation, B_filtered_CCAInt_clusters),
                                    by = c("Stimulation", "B_filtered_CCAInt_clusters"))
# Indicate which neighbourhoods to include in downstream analysis
plotBee_ClusAnnot.data <- plotBee_ClusAnnot.data%>%
  # Neighbourhood is selected if logFC is larger then the cutoff (only for stimulated clusters)
  # OR is present in the selected NBH of the unstimulated cluster
  mutate(ToInclude = (logFC > CutOff) | (Nhood %in% NBH_Selection_Unstim.Numbers))%>%
  mutate(ToInclude = replace_na(ToInclude, FALSE))
```


#### Visualization of cutoffs and selected neighbourhoods

```{r NBH_PlotBeeswarmWithCutoffs_Arrows_8h, fig.height=10, fig.width=10}
plotBee_ClusAnnot$IFNgamma$plot+
  # add line segment
  geom_segment(data = NBH_Selection_IFNgamma_ODN2006.CutOffs%>%
                 filter(Stimulation == "IFNgamma"),
               mapping = aes(x = positionHigh, xend = positionLow,
                             y = CutOff),
               color = "black", linewidth = 1, alpha = 0.5)+
  # add arrow next to the segment
  geom_segment(data = NBH_Selection_IFNgamma_ODN2006.CutOffs%>%
                 filter(Stimulation == "IFNgamma"),
               mapping = aes(x = positionLow, 
                             y = CutOff, yend = CutOff+1),
               color = "black", linewidth = 1, alpha=0.5, 
               arrow = arrow(length = unit(0.03, "npc")))+
(plotBee_ClusAnnot$ODN2006$plot+
  # add line segment
  geom_segment(data = NBH_Selection_IFNgamma_ODN2006.CutOffs%>%
                 filter(Stimulation == "ODN2006"),
               mapping = aes(x = positionHigh, xend = positionLow,
                             y = CutOff),
               color = "black", linewidth = 1, alpha = 0.5)+
  # add arrow next to the segment
  geom_segment(data = NBH_Selection_IFNgamma_ODN2006.CutOffs%>%
                 filter(Stimulation == "ODN2006"),
               mapping = aes(x = positionLow, 
                             y = CutOff, yend = CutOff+1),
               color = "black", linewidth = 1, alpha=0.5, 
               arrow = arrow(length = unit(0.03, "npc"))))+
(plotBee_ClusAnnot$MegaCD40L$plot+
  # add line segment
  geom_segment(data = NBH_Selection_CD40L_CD3CD28.CutOffs%>%
                 filter(Stimulation == "MegaCD40L"),
               mapping = aes(x = positionHigh, xend = positionLow,
                             y = CutOff),
               color = "black", linewidth = 1, alpha = 0.5)+
  # add arrow next to the segment
  geom_segment(data = NBH_Selection_CD40L_CD3CD28.CutOffs%>%
                 filter(Stimulation == "MegaCD40L"),
               mapping = aes(x = positionLow, 
                             y = CutOff, yend = CutOff+1),
               color = "black", linewidth = 1, alpha=0.5, 
               arrow = arrow(length = unit(0.03, "npc"))))+
(plotBee_ClusAnnot$antiCD3CD28$plot+
  # add line segment
  geom_segment(data = NBH_Selection_CD40L_CD3CD28.CutOffs%>%
                 filter(Stimulation == "antiCD3CD28"),
               mapping = aes(x = positionHigh, xend = positionLow,
                             y = CutOff),
               color = "black", linewidth = 1, alpha = 0.5)+
  # add arrow next to the segment
  geom_segment(data = NBH_Selection_CD40L_CD3CD28.CutOffs%>%
                 filter(Stimulation == "antiCD3CD28"),
               mapping = aes(x = positionLow, 
                             y = CutOff, yend = CutOff+1),
               color = "black", linewidth = 1, alpha=0.5, 
               arrow = arrow(length = unit(0.03, "native"))))+
  plot_annotation("8h") & labs(x = "Original clusters") & ylim(-10,10)
```

```{r NBH_PlotBeeswarmWithCutoffs_8h, fig.height=10, fig.width=10}
shapeValues = c("TRUE" = 19, "FALSE" = 1)
sizeValues = c("TRUE" = 1, "FALSE" = 0.5)

plotDAbeeswarm_SubsetIndication(plotBee_ClusAnnot.data%>%filter(Stimulation == "IFNgamma"), 
                                group.by  = "B_filtered_CCAInt_clusters", alpha = FDR_thres, ToInclude = "ToInclude",
                                sizeValues = sizeValues, shapeValues = shapeValues)+
  labs(subtitle = "IFNgamma")+
plotDAbeeswarm_SubsetIndication(plotBee_ClusAnnot.data%>%filter(Stimulation == "ODN2006"), 
                                group.by  = "B_filtered_CCAInt_clusters", alpha = FDR_thres, ToInclude = "ToInclude",
                                sizeValues = sizeValues, shapeValues = shapeValues)+
  labs(subtitle = "ODN2006")+
plotDAbeeswarm_SubsetIndication(plotBee_ClusAnnot.data%>%filter(Stimulation == "antiCD3CD28"), 
                                group.by  = "B_filtered_CCAInt_clusters", alpha = FDR_thres, ToInclude = "ToInclude",
                                sizeValues = sizeValues, shapeValues = shapeValues)+
  labs(subtitle = "Anti-CD3/CD28")+
plotDAbeeswarm_SubsetIndication(plotBee_ClusAnnot.data%>%filter(Stimulation == "MegaCD40L"), 
                                group.by  = "B_filtered_CCAInt_clusters", alpha = FDR_thres, ToInclude = "ToInclude",
                                sizeValues = sizeValues, shapeValues = shapeValues)+
  labs(subtitle = "MegaCD40L")+
  plot_annotation("8h") & labs(x = "Original clusters") & ylim(-10,10)&
  themeG &
  theme(panel.grid.major.y = element_line(color = adjustcolor( "lightgrey", alpha.f = 0.1),
                                          linewidth = 1)) &  NoLegend()
```

```{r NBH_Retained_AbsoluteNumbers_8h}
## Retained number of NBHs per cluster
plotBee_ClusAnnot.data%>%
  group_by(Stimulation, B_filtered_CCAInt_clusters)%>%
  summarise(RetainedNBH = sum(ToInclude))%>%
  filter(RetainedNBH != 0)%>%
  arrange(B_filtered_CCAInt_clusters)

## Retained number of NBHs per Stimulation
plotBee_ClusAnnot.data%>%
  filter(B_filtered_CCAInt_clusters != 0)%>%
  group_by(Stimulation)%>%
  summarise(Retained_StimNBH = sum(ToInclude))
```

#### Extraction of the cell IDs of interest based on the selected NBHs

```{r ExtractionCellIDS_8h, echo=FALSE}
# Unstimulated
# Selection of cells linked to the selected NBHs
NBH_Selection_Unstim.Cells<-sapply(as.numeric(NBH_Selection_Unstim.Numbers),
                                   FUN = function(nhood)
                                     {nhoods(milo.B_8h.FilClus$milo_graph)@Dimnames[[1]][nhoods(milo.B_8h.FilClus$milo_graph)[,nhood]==1]})%>%
                                       unlist()%>%
                                       unique()

# Selection of the Nhoods of interest only linked to a stimulation
NBH_Selection.Nhoods<-plotBee_ClusAnnot.data%>%
  filter(ToInclude)%>%
  select(Nhood, Stimulation)
# Selection of cellIDS
for (nhood in unique(NBH_Selection.Nhoods$Nhood)){
  if (nhood %in% NBH_Selection_Unstim.Numbers){
    # Keep track of unstimulated Nhoods as well
    Condition <- "Unstimulated"
  } else {
    Condition <- "Stimulated"
  }
  ExtractedCells.tmp <- data.frame(Nhood = nhood,
                                   CellID = nhoods(milo.B_8h.FilClus$milo_graph)@Dimnames[[1]][nhoods(milo.B_8h.FilClus$milo_graph)[,nhood]==1],
                                   Condition = Condition)
  if (!exists("NBH_Selection.Cells") || (dim(NBH_Selection.Cells)[1]) == 0){
    message("Creating dataframe")
    NBH_Selection.Cells <- ExtractedCells.tmp
  } else {
    message("\nJoining additional cells")
    NBH_Selection.Cells <- rbind(NBH_Selection.Cells,
                                 ExtractedCells.tmp)
  }
  # Clean environment
  rm(ExtractedCells.tmp)
}

# Joining the Nhood information to the Cell IDs
NBH_Selection.Cells <- left_join(NBH_Selection.Cells,
                                 NBH_Selection.Nhoods,
                                 by = "Nhood")
# Annotate the cells that are overlapping with the unstimulated cells
NBH_Selection.Cells <- NBH_Selection.Cells%>%
  mutate(Condition = case_when(
    CellID %in% NBH_Selection_Unstim.Cells ~ "Unstimulated",
    TRUE ~ "Stimulated"
  ))
```


#### Save selection of cells
```{r SaveSelectedCells_8h}
# Create a blank workbook
WorkBook <- createWorkbook()

# Add some sheets to the workbook
addWorksheet(WorkBook, "CutOffs")
addWorksheet(WorkBook, "Nhoods_Sel")
addWorksheet(WorkBook, "Cells_Sel")
addWorksheet(WorkBook, "UnstimCells")
addWorksheet(WorkBook, "OverviewData")

# Write the data to the sheets
writeData(WorkBook, sheet = "CutOffs", 
          x = NBH_Selection.CutOffs%>%
            select(-positionHigh, -positionLow))
writeData(WorkBook, sheet = "Nhoods_Sel", 
          x = NBH_Selection.Nhoods)
writeData(WorkBook, sheet = "Cells_Sel", 
          x = NBH_Selection.Cells)
writeData(WorkBook, sheet = "UnstimCells",
          x = NBH_Selection_Unstim.Cells)
writeData(WorkBook, sheet = "OverviewData",
          x = plotBee_ClusAnnot.data)

# Export the file
saveWorkbook(WorkBook, 
             # overwrite = T,
             file.path(params$path_Results,
                            "MiloR",
                            paste(params$date2, "CutOffs_Selection_Nhoods_Cells", "8h.xlsx", sep = "_")))
```

```{r readInData_8h, eval=FALSE}
NBH_Selection.Nhoods<-read.xlsx(xlsxFile = file.path(params$path_Results,
                                                         "MiloR",
                                                         paste(params$date2, "CutOffs_Selection_Nhoods_Cells", "8h.xlsx", sep = "_")),
                            sheet = "Nhoods_Sel")
NBH_Selection.Cells<-read.xlsx(xlsxFile = file.path(params$path_Results,
                                                         "MiloR",
                                                         paste(params$date2, "CutOffs_Selection_Nhoods_Cells", "8h.xlsx", sep = "_")),
                            sheet = "Cells_Sel")

NBH_Selection_Unstim.Cells<-read.xlsx(xlsxFile = file.path(params$path_Results,
                                                         "MiloR",
                                                         paste(params$date2, "CutOffs_Selection_Nhoods_Cells", "8h.xlsx", sep = "_")),
                            sheet = "UnstimCells")
plotBee_ClusAnnot.data<-read.xlsx(xlsxFile = file.path(params$path_Results,
                                                         "MiloR",
                                                         paste(params$date2, "CutOffs_Selection_Nhoods_Cells", "8h.xlsx", sep = "_")),
                            sheet = "OverviewData")
```

### Selection of interesting neighbourhoods - hard cutoff over stimulations

#### Unstimulated neighbourhoods

```{r NBH_Selection_8h_UnstimulatedCells_FIXED}
# Define hard cutoff
CutOff.Fixed<-3
# Join all data of the beeswarms over the stimulations
plotBee_ClusAnnot.Fixed.data<-rbind(plotBee_ClusAnnot$IFNgamma$data,
                              plotBee_ClusAnnot$MegaCD40L$data,
                              plotBee_ClusAnnot$antiCD3CD28$data,
                              plotBee_ClusAnnot$ODN2006$data)
```

```{r NBH_Overlap_Unstimulated_FIXED}
# Filtering based on the cutoffs
NBH_Selection_Unstim.Fixed<-plotBee_ClusAnnot.Fixed.data%>%
  mutate(CutOff_Fixed = -CutOff.Fixed)%>%
  filter(B_filtered_CCAInt_clusters %in% c(0, 9, 11, 13),
         logFC < CutOff_Fixed)

# Prepare UpsetPlotData
NBH_Selection_Unstim.Fixed.CombMat<-lapply(unique(NBH_Selection_Unstim.Fixed$Stimulation),
                                           FUN = function(stim){(NBH_Selection_Unstim.Fixed%>%filter(Stimulation == stim))[["Nhood"]]})
names(NBH_Selection_Unstim.Fixed.CombMat)<-unique(NBH_Selection_Unstim.Fixed$Stimulation)
NBH_Selection_Unstim.Fixed.CombMat<-make_comb_mat(NBH_Selection_Unstim.Fixed.CombMat, mode = "distinct")
# Plot upset plot
UpSet(NBH_Selection_Unstim.Fixed.CombMat,
      comb_order = order(comb_size(NBH_Selection_Unstim.Fixed.CombMat), decreasing = T),
      top_annotation = HeatmapAnnotation(
                "Intersection size" = anno_barplot(comb_size(NBH_Selection_Unstim.Fixed.CombMat),
                                                   border = FALSE,
                                                   gp = gpar(fill = "#595B5C",
                                                             lty = "blank"),
                                                   add_numbers = T,
                                                   height = unit(4, "cm")),
                annotation_name_side = "left",
                annotation_name_rot = 90),
      right_annotation = rowAnnotation(
                "DE genes" = anno_barplot(set_size(NBH_Selection_Unstim.Fixed.CombMat),
                                          border = FALSE,
                                          gp = gpar(fill = "#595B5C",
                                                    lty = "blank"),
                                          add_numbers = T,
                                          width = unit(4, "cm"))))

# Selected NBHs - overlap in at least 3 of the stimulatory conditions
NBH_Selection_Unstim.Fixed.Numbers<-sapply(X = c("1110", "1111", "1011", "1101"),
                                     FUN = function(set){extract_comb(NBH_Selection_Unstim.Fixed.CombMat, set)})%>%
  unlist()
length(NBH_Selection_Unstim.Fixed.Numbers)

# Select the neighbourhoods based on the overlap
NBH_Selection_Unstim.Fixed<-NBH_Selection_Unstim.Fixed%>%
  mutate(ToInclude = case_when(
    Nhood %in% NBH_Selection_Unstim.Fixed.Numbers ~ TRUE,
    TRUE ~ FALSE))
```

#### Stimulated neighbourhoods 

```{r NBH_Selection_8h_Stimulations_FIXED, fig.height=10, fig.width=10}
# Filtering based on the cutoffs
NBH_Selection_Stim.Fixed<-plotBee_ClusAnnot.Fixed.data%>%
  mutate(CutOff_Fixed = CutOff.Fixed)%>%
  filter(B_filtered_CCAInt_clusters != 0,
         logFC > CutOff_Fixed)

# Indicate which NBHs to include
NBH_Selection_Stim.Fixed<-NBH_Selection_Stim.Fixed%>%
  mutate(ToInclude = case_when(
    Stimulation %in% c("IFNgamma", "ODN2006") & B_filtered_CCAInt_clusters %in% c(1, 3, 6, 12) ~ TRUE,
    Stimulation %in% c("antiCD3CD28", "MegaCD40L") & B_filtered_CCAInt_clusters %in% c(2, 5, 7) ~TRUE,
    TRUE ~ FALSE
  ))

# Check the number of retained neighbourhoods per stimulation and cluster
table(NBH_Selection_Stim.Fixed$Stimulation,
      NBH_Selection_Stim.Fixed$B_filtered_CCAInt_clusters)
# Check the number of retained neighbourhoods per stimulation
table(NBH_Selection_Stim.Fixed$Stimulation,
      NBH_Selection_Stim.Fixed$ToInclude)
```

#### Merging of cut-off values and selected NBH indication


```{r NBH_Selection_8h}
NBH_Selection.Fixed<-rbind(NBH_Selection_Unstim.Fixed,
                           NBH_Selection_Stim.Fixed)

# Add cutoffs to the data
plotBee_ClusAnnot.Fixed.data <- left_join(plotBee_ClusAnnot.Fixed.data,
                                    NBH_Selection.Fixed%>%
                                      select(CutOff_Fixed, Stimulation, Nhood, ToInclude),
                                    by = c("Stimulation", "Nhood"))
# Indicate which neighbourhoods to include in downstream analysis
plotBee_ClusAnnot.Fixed.data <- plotBee_ClusAnnot.Fixed.data%>%
  # Neighbourhood is selected if logFC is larger then the cutoff (only for stimulated clusters)
  # OR is present in the selected NBH of the unstimulated cluster
  mutate(ToInclude = replace_na(ToInclude, FALSE))
```

#### Visualization of cutoffs and selected neighbourhoods

```{r NBH_PlotBeeswarmWithCutoffs_8h_FIXED, fig.height=10, fig.width=10}
shapeValues = c("TRUE" = 19, "FALSE" = 1)
sizeValues = c("TRUE" = 1, "FALSE" = 0.5)

plotDAbeeswarm_SubsetIndication(plotBee_ClusAnnot.Fixed.data%>%filter(Stimulation == "IFNgamma"), 
                                group.by  = "B_filtered_CCAInt_clusters", alpha = FDR_thres, ToInclude = "ToInclude",
                                sizeValues = sizeValues, shapeValues = shapeValues)+
  labs(subtitle = "IFNgamma")+
plotDAbeeswarm_SubsetIndication(plotBee_ClusAnnot.Fixed.data%>%filter(Stimulation == "ODN2006"), 
                                group.by  = "B_filtered_CCAInt_clusters", alpha = FDR_thres, ToInclude = "ToInclude",
                                sizeValues = sizeValues, shapeValues = shapeValues)+
  labs(subtitle = "ODN2006")+
plotDAbeeswarm_SubsetIndication(plotBee_ClusAnnot.Fixed.data%>%filter(Stimulation == "antiCD3CD28"), 
                                group.by  = "B_filtered_CCAInt_clusters", alpha = FDR_thres, ToInclude = "ToInclude",
                                sizeValues = sizeValues, shapeValues = shapeValues)+
  labs(subtitle = "Anti-CD3/CD28")+
plotDAbeeswarm_SubsetIndication(plotBee_ClusAnnot.Fixed.data%>%filter(Stimulation == "MegaCD40L"), 
                                group.by  = "B_filtered_CCAInt_clusters", alpha = FDR_thres, ToInclude = "ToInclude",
                                sizeValues = sizeValues, shapeValues = shapeValues)+
  labs(subtitle = "MegaCD40L")+
  plot_annotation("8h") & labs(x = "Original clusters") & ylim(-10,10)&
  themeG &
  theme(panel.grid.major.y = element_line(color = adjustcolor( "lightgrey", alpha.f = 0.1),
                                          linewidth = 1)) &  NoLegend()
```

#### Extraction of the cell IDs of interest based on the selected NBHs

```{r ExtractionCellIDS_8h_FIXED, echo=FALSE}
# Unstimulated
# Selection of cells linked to the selected NBHs
NBH_Selection_Unstim.Fixed.Cells<-sapply(as.numeric(NBH_Selection_Unstim.Fixed.Numbers),
                                   FUN = function(nhood){nhoods(milo.B_8h.FilClus$milo_graph)@Dimnames[[1]][nhoods(
                                     milo.B_8h.FilClus$milo_graph)[,nhood]==1]})%>%
                                       unlist()%>%
                                       unique()

# Selection of the Nhoods of interest only linked to a stimulation
NBH_Selection.Fixed.Nhoods<-plotBee_ClusAnnot.Fixed.data%>%
  filter(ToInclude)%>%
  select(Nhood, Stimulation)
# Selection of cellIDS
for (nhood in unique(NBH_Selection.Fixed.Nhoods$Nhood)){
  if (nhood %in% NBH_Selection_Unstim.Fixed.Numbers){
    # Keep track of unstimulated Nhoods as well
    Condition <- "Unstimulated"
  } else {
    Condition <- "Stimulated"
  }
  ExtractedCells.tmp <- data.frame(Nhood = nhood,
                                   CellID = nhoods(milo.B_8h.FilClus$milo_graph)@Dimnames[[1]][nhoods(
                                     milo.B_8h.FilClus$milo_graph)[,nhood]==1],
                                   Condition = Condition)
  if (!exists("NBH_Selection.Fixed.Cells") || (dim(NBH_Selection.Fixed.Cells)[1]) == 0){
    message("Creating dataframe")
    NBH_Selection.Fixed.Cells <- ExtractedCells.tmp
  } else {
    message("\nJoining additional cells")
    NBH_Selection.Fixed.Cells <- rbind(NBH_Selection.Fixed.Cells,
                                 ExtractedCells.tmp)
  }
  # Clean environment
  rm(ExtractedCells.tmp)
}

# Joining the Nhood information to the Cell IDs
NBH_Selection.Fixed.Cells <- left_join(NBH_Selection.Fixed.Cells,
                                 NBH_Selection.Fixed.Nhoods,
                                 by = "Nhood")
# Annotate the cells that are overlapping with the unstimulated cells
NBH_Selection.Fixed.Cells <- NBH_Selection.Fixed.Cells%>%
  mutate(Condition = case_when(
    CellID %in% NBH_Selection_Unstim.Fixed.Cells ~ "Unstimulated",
    TRUE ~ "Stimulated"
  ))
```


#### Save selection of cells
```{r SaveSelectedCells_8h}
# Create a blank workbook
WorkBook <- createWorkbook()

# Add some sheets to the workbook
addWorksheet(WorkBook, "Nhoods_Sel")
addWorksheet(WorkBook, "Cells_Sel")
addWorksheet(WorkBook, "UnstimCells")
addWorksheet(WorkBook, 'OverviewData')

# Write the data to the sheets
writeData(WorkBook, sheet = "Nhoods_Sel", 
          x = NBH_Selection.Fixed.Nhoods)
writeData(WorkBook, sheet = "Cells_Sel", 
          x = NBH_Selection.Fixed.Cells)
writeData(WorkBook, sheet = "UnstimCells",
          x = NBH_Selection_Unstim.Fixed.Cells)
writeData(WorkBook, sheet = "OverviewData",
          x = plotBee_ClusAnnot.Fixed.data)

# Export the file
saveWorkbook(WorkBook, 
             overwrite = T,
             file.path(params$path_Results,
                            "MiloR",
                            paste(params$date3, "HardCutOff3_Selection_Nhoods_Cells", "8h.xlsx", sep = "_")))
```

```{r ReadInData_SelectionCells_Fixed_8h}
NBH_Selection.Fixed.Nhoods <- read.xlsx(file.path(params$path_Results,
                            "MiloR",
                            paste(params$date3, "HardCutOff3_Selection_Nhoods_Cells", "8h.xlsx", sep = "_")),
                            sheet = "Nhoods_Sel")
NBH_Selection.Fixed.Cells <- read.xlsx(file.path(params$path_Results,
                            "MiloR",
                            paste(params$date3, "HardCutOff3_Selection_Nhoods_Cells", "8h.xlsx", sep = "_")),
                            sheet = "Cells_Sel")
NBH_Selection_Unstim.Fixed.Cells <- read.xlsx(file.path(params$path_Results,
                            "MiloR",
                            paste(params$date3, "HardCutOff3_Selection_Nhoods_Cells", "8h.xlsx", sep = "_")),
                            sheet = "UnstimCells")
plotBee_ClusAnnot.Fixed.data <- read.xlsx(file.path(params$path_Results,
                            "MiloR",
                            paste(params$date3, "HardCutOff3_Selection_Nhoods_Cells", "8h.xlsx", sep = "_")),
                            sheet = "OverviewData")
```

#### Comparison to selection based on soft cut-offs

```{r ComparisonCutOffs_8h}
# For stimulated cells
for(stim in unique(plotBee_ClusAnnot.data$Stimulation)){
  # Prepare plotdata
  CompPlotdata<-list("SOFT" = (plotBee_ClusAnnot.data%>%filter(Stimulation == stim,
                                                               B_filtered_CCAInt_clusters != 0,
                                                               ToInclude == TRUE))[["Nhood"]],
                     "HARD" = (plotBee_ClusAnnot.Fixed.data%>%filter(Stimulation == stim,
                                                               B_filtered_CCAInt_clusters != 0,
                                                               ToInclude == TRUE))[["Nhood"]])
  CompPlotdata<-make_comb_mat(CompPlotdata, mode = "distinct")
  
  # Plot upsetplot
  print(UpSet(CompPlotdata,
        comb_order = order(comb_size(CompPlotdata), decreasing = T),
        top_annotation = HeatmapAnnotation(
                  "Intersection size" = anno_barplot(comb_size(CompPlotdata),
                                                     border = FALSE,
                                                     gp = gpar(fill = "#595B5C",
                                                               lty = "blank"),
                                                     add_numbers = T,
                                                     height = unit(4, "cm")),
                  annotation_name_side = "left",
                  annotation_name_rot = 90),
        right_annotation = rowAnnotation(
                  "DE genes" = anno_barplot(set_size(CompPlotdata),
                                            border = FALSE,
                                            gp = gpar(fill = "#595B5C",
                                                      lty = "blank"),
                                            add_numbers = T,
                                            width = unit(4, "cm"))),
        row_title = stim))
  rm(CompPlotdata)
}

# For unstimulated cells
# Prepare plotdata
CompPlotdata<-list("SOFT" = (plotBee_ClusAnnot.data%>%filter(Stimulation == stim,
                                                             B_filtered_CCAInt_clusters == 0,
                                                             ToInclude == TRUE))[["Nhood"]],
                   "HARD" = (plotBee_ClusAnnot.Fixed.data%>%filter(Stimulation == stim,
                                                             B_filtered_CCAInt_clusters == 0,
                                                             ToInclude == TRUE))[["Nhood"]])
CompPlotdata<-make_comb_mat(CompPlotdata, mode = "distinct")

# Plot upsetplot
print(UpSet(CompPlotdata,
      comb_order = order(comb_size(CompPlotdata), decreasing = T),
      top_annotation = HeatmapAnnotation(
                "Intersection size" = anno_barplot(comb_size(CompPlotdata),
                                                   border = FALSE,
                                                   gp = gpar(fill = "#595B5C",
                                                             lty = "blank"),
                                                   add_numbers = T,
                                                   height = unit(4, "cm")),
                annotation_name_side = "left",
                annotation_name_rot = 90),
      right_annotation = rowAnnotation(
                "DE genes" = anno_barplot(set_size(CompPlotdata),
                                          border = FALSE,
                                          gp = gpar(fill = "#595B5C",
                                                    lty = "blank"),
                                          add_numbers = T,
                                          width = unit(4, "cm"))),
      row_title = "Unstimulated"))
rm(CompPlotdata)
```

```{r ComparisonCutOffs_Cells_8h}
# For stimulated cells
for(stim in unique(NBH_Selection.Cells$Stimulation)){
  # Prepare plotdata
  CompPlotdata<-list("SOFT" = (NBH_Selection.Cells%>%filter(Stimulation == stim,
                                                               Condition == "Stimulated"))[["CellID"]]%>%unique(),
                     "HARD" = (NBH_Selection.Fixed.Cells%>%filter(Stimulation == stim,
                                                               Condition == "Stimulated"))[["CellID"]]%>%unique())
  CompPlotdata<-make_comb_mat(CompPlotdata, mode = "distinct")
  
  # Plot upsetplot
  print(UpSet(CompPlotdata,
        comb_order = order(comb_size(CompPlotdata), decreasing = T),
        top_annotation = HeatmapAnnotation(
                  "Intersection size" = anno_barplot(comb_size(CompPlotdata),
                                                     border = FALSE,
                                                     gp = gpar(fill = "#595B5C",
                                                               lty = "blank"),
                                                     add_numbers = T,
                                                     height = unit(4, "cm")),
                  annotation_name_side = "left",
                  annotation_name_rot = 90),
        right_annotation = rowAnnotation(
                  "DE genes" = anno_barplot(set_size(CompPlotdata),
                                            border = FALSE,
                                            gp = gpar(fill = "#595B5C",
                                                      lty = "blank"),
                                            add_numbers = T,
                                            width = unit(4, "cm"))),
        row_title = stim))
  rm(CompPlotdata)
}

# For unstimulated cells
# Prepare plotdata
CompPlotdata<-list("SOFT" = (NBH_Selection.Cells%>%filter(Condition == "Unstimulated"))[["CellID"]]%>%unique(),
                   "HARD" = (NBH_Selection.Fixed.Cells%>%filter(Condition == "Unstimulated"))[["CellID"]]%>%unique())
CompPlotdata<-make_comb_mat(CompPlotdata, mode = "distinct")

# Plot upsetplot
print(UpSet(CompPlotdata,
      comb_order = order(comb_size(CompPlotdata), decreasing = T),
      top_annotation = HeatmapAnnotation(
                "Intersection size" = anno_barplot(comb_size(CompPlotdata),
                                                   border = FALSE,
                                                   gp = gpar(fill = "#595B5C",
                                                             lty = "blank"),
                                                   add_numbers = T,
                                                   height = unit(4, "cm")),
                annotation_name_side = "left",
                annotation_name_rot = 90),
      right_annotation = rowAnnotation(
                "DE genes" = anno_barplot(set_size(CompPlotdata),
                                          border = FALSE,
                                          gp = gpar(fill = "#595B5C",
                                                    lty = "blank"),
                                          add_numbers = T,
                                          width = unit(4, "cm"))),
      row_title = "Unstimulated"))
rm(CompPlotdata)
```

## 30 hours of stimulation

### Data analysis using MiloR

```{r MiloR_DAanalysis_30h}
set.seed(2024)
# Generation of the milo object and graph
# Conform the vignette of MiloR
milo.B_30h.FilClus<-MiloDA(object = IntData.B_30h.FilCus, assay = "RNA", k = 50, d = 20, prop = 0.15,
                          # As recommended by Noah, perform KNN graph generation based on the dimensionality reduction technique
                          # containing the most information
                          RedDim = "B_filtered_CCAInt.rna", sampleID = "PatStim", 
                          design = c("PatStim", "Stimulation"), overlap = 5)
# Save the object for later use
saveRDS(milo.B_30h.FilClus, 
        file = file.path(params$path_Results, "MiloR",
                         paste(params$prefix_output, "MiloRobject_FilteredB_30h.rds", sep = "_")))

milo.B_30h.FilClus$size_histogram
# Set k such that you have a minimum of Sx5 counts per nhood, where S=total number of samples

# Statistical testing
milo.B_30h.FilClus$milo_design$Stimulation <- milo.B_30h.FilClus$milo_design$Stimulation%>%
  gsub(pattern = "CD3/CD28_beads", replacement = "antiCD3CD28")
contrasts<-c("StimulationIFNgamma - StimulationUnstimulated", "StimulationMegaCD40L - StimulationUnstimulated", 
             "StimulationantiCD3CD28 - StimulationUnstimulated", "StimulationODN2006 - StimulationUnstimulated")

# Extract results of differential abundance analysis
# Note: This part is a bit different from what is performed in 006_NoBase_visualization_doubletCheck_ClusterAnalysis_IntegrationV5.Rmd script
DA.B_30h.FilClus<-testNhoods(x = milo.B_30h.FilClus$milo_graph, design = ~0+Stimulation, design.df = milo.B_30h.FilClus$milo_design,
                            # Correct for multiple testing using a spatial FDR correction
                            fdr.weighting = "graph-overlap", 
                            model.contrasts = contrasts, reduced.dim = "B_filtered_CCAInt.rna"%>%toupper())

# Rename the columns of the DA results
names(DA.B_30h.FilClus)
DA.B_30h.FilClus<- DA.B_30h.FilClus%>%
  pivot_longer(cols = starts_with(("logFC")),
               names_to = "Stimulation",
               values_to = "logFC")%>%
  mutate(Stimulation = gsub(pattern = "logFC.Stimulation", 
                            replacement = "",
                            x = Stimulation),
         Stimulation = gsub(pattern = "...S.*", 
                            replacement = "",
                            x = Stimulation))
``` 

```{r ReadIn_MiloObjects_30h}
milo.B_30h.FilClus<-readRDS(file = file.path(params$path_Results, "MiloR",
                                            paste(params$prefix_output, "MiloRobject_FilteredB_30h.rds", sep = "_")))
```

### Neighbourhood graph plotting

```{r MiloDA_BCells_Filtered_Visualisation_30h}
# Visualization 
# Generation of plots
pIFN<-plotNhoodGraphDA(x = milo.B_30h.FilClus$milo_graph, alpha = 0.1,
                   milo_res = DA.B_30h.FilClus%>%filter(Stimulation == "IFNgamma"), layout = "B_filtered_CCAInt.umap.rna"%>%toupper(),
                   size_range = c(0.01,1.5), node_stroke = 0.2)+
    labs(subtitle = paste0("IFN","\U03B3","/R-848"))+
    guides(size=guide_legend("Size of Nhoods"), 
           edge_width = guide_legend("Overlap Nhoods"), 
           fill = guide_colourbar("LogFC", theme = theme(legend.key.width  = unit(0.5, "lines"))))

pCD40L<-plotNhoodGraphDA(x = milo.B_30h.FilClus$milo_graph, alpha = 0.1,
                   milo_res = DA.B_30h.FilClus%>%filter(Stimulation == "MegaCD40L"), layout = "B_filtered_CCAInt.umap.rna"%>%toupper(),
                   size_range = c(0.01,1.5), node_stroke = 0.2)+
    labs(subtitle = "MegaCD40L")+
    guides(size=guide_legend("Size of Nhoods"), 
           edge_width = guide_legend("Overlap Nhoods"), 
           fill = guide_colourbar("LogFC", theme = theme(legend.key.width  = unit(0.5, "lines"))))

pCD3CD28<-plotNhoodGraphDA(x = milo.B_30h.FilClus$milo_graph, alpha = 0.1,
                   milo_res = DA.B_30h.FilClus%>%filter(Stimulation == "antiCD3CD28"), layout = "B_filtered_CCAInt.umap.rna"%>%toupper(),
                   size_range = c(0.01,1.5), node_stroke = 0.2)+
    labs(subtitle = "Anti-CD3/CD28")+
    guides(size=guide_legend("Size of Nhoods"), 
           edge_width = guide_legend("Overlap Nhoods"), 
           fill = guide_colourbar("LogFC", theme = theme(legend.key.width  = unit(0.5, "lines"))))

pODN<-plotNhoodGraphDA(x = milo.B_30h.FilClus$milo_graph, alpha = 0.1,
                   milo_res = DA.B_30h.FilClus%>%filter(Stimulation == "ODN2006"), layout = "B_filtered_CCAInt.umap.rna"%>%toupper(),
                   size_range = c(0.01,1.5), node_stroke = 0.2)+
    labs(subtitle = "ODN2006")+
    guides(size=guide_legend("Size of Nhoods"), 
           edge_width = guide_legend("Overlap Nhoods"), 
           fill = guide_colourbar("LogFC", theme = theme(legend.key.width  = unit(0.5, "lines"))))

pUMAP<-DimPlot(object = IntData.B_30h.FilCus,
        reduction = "B_filtered_CCAInt.umap.rna", group.by = "Stimulation",
        pt.size = 0.001)+
    scale_color_manual(values = colorsStim)+
    labs(title=NULL, x = "UMAP1", y = "UMAP2", color = "Stimulation")+
    theme(axis.line = element_blank(), axis.text = element_blank(),
          axis.ticks = element_blank(), axis.title = element_blank())+
    guides(color = guide_legend(override.aes = list(size=1))) 
# Via patchwork, arranging of plots
pUMAP + plot_spacer() + ( (pIFN + plot_spacer() + pODN + plot_layout(widths = c(1,0,1))) / 
                            (pCD3CD28 + plot_spacer() + pCD40L + plot_layout(widths = c(1,0,1))))+
  plot_layout(guides = "collect", widths = c(1,0,1)) & 
  scale_fill_gradientn(limits = range(min(DA.B_30h.FilClus$logFC), max(DA.B_30h.FilClus$logFC)),
                       colours = c("#8C2F39", "white", "#2D0650"),
                       name = "LogFC") &
  labs(size = "Size of Nhood") &
  guides(fill = guide_colourbar("LogFC", theme = theme(legend.key.width  = unit(0.5, "lines")))) &
  theme(plot.subtitle = element_text(size = 10, color = "gray30"),
        legend.text =element_text(margin = margin(t = 1,l = 3), size = 8, color = "gray30"),
        legend.title = element_text(size = 10, color = "gray30"),
        plot.margin = margin(1,0,0,0)) &
  theme(legend.key.size = unit(1, "lines"),
        legend.spacing.y = unit(0.5, "lines"),
        legend.spacing.x = unit(0.5, "lines"),
        legend.margin = margin(1,0,0,1), 
        legend.key.spacing = unit(0.3, "lines"))
  
ggsave(filename = file.path(params$path_Results,
                            "MiloR",
                            paste(params$prefix_output, "DA_FilteredB_30h_RefUnstim.jpeg",
                                  sep = "_")),
       bg = "white", width = 25, height = 14, device = "jpeg", units = "cm",
       dpi = 400)

# Cleaning workspace
rm(pIFN, pODN,pCD3CD28, pCD40L, pUMAP)
```

### Beeswarm plots based on original clusters

```{r OriginalClustering_Nhoods_30h, fig.height=10, fig.width=10}
# Annotation of the Nhoods per stimulation linking to the original clusters
DA.B_30h.FilClus_ClustIFNG <- annotateNhoods(milo.B_30h.FilClus$milo_graph, 
                                            # Only possible per stimulation
                                            DA.B_30h.FilClus%>%
                                              filter(Stimulation == "IFNgamma"), 
                                            # Linking back to the clusters obtained after filtering and CCA integration
                                            coldata_col = "B_filtered_CCAInt_clusters")
# Convert to dataframe to use as input for the plotDAbeeswarm function
DA.B_30h.FilClus_ClustIFNG <- as.data.frame(DA.B_30h.FilClus_ClustIFNG)

# Check the outcome after Nhood annotation
## Interesting: the fractions of linked clusters to the Nhoods is given as well
head(DA.B_30h.FilClus_ClustIFNG)

# Annotation of the Nhoods per stimulation linking to the original clusters
plotBee_ClusAnnot_30h <- lapply(unique(DA.B_30h.FilClus$Stimulation),
                                     FUN = function(layer){
                                       # Annotation of the Nhoods per stimulation linking to the original clusters
                                       DA.B_30h.FilClus_ClusAnnot.tmp <- annotateNhoods(milo.B_30h.FilClus$milo_graph,
                                                                                       # Only possible per stimulation
                                                                                      DA.B_30h.FilClus%>%
                                                                                        filter(Stimulation == layer), 
                                                                                      # Linking back to the clusters obtained after filtering and CCA integration
                                                                                      coldata_col = "B_filtered_CCAInt_clusters")
                                       # Highligh neighbourhoods that are a mix of clusters
                                       DA.B_30h.FilClus_ClusAnnot.tmp$B_filtered_CCAInt_clusters <- ifelse(
                                         DA.B_30h.FilClus_ClusAnnot.tmp$B_filtered_CCAInt_clusters_fraction < 0.7, "Mixed", 
                                         DA.B_30h.FilClus_ClusAnnot.tmp$B_filtered_CCAInt_clusters)
                                       # Convert to dataframe to use as input for the plotDAbeeswarm function
                                       DA.B_30h.FilClus_ClusAnnot.tmp <- as.data.frame(DA.B_30h.FilClus_ClusAnnot.tmp)%>%
                                         # Ordering the clusters to visualize properly in the graph
                                         mutate(B_filtered_CCAInt_clusters = factor(B_filtered_CCAInt_clusters,
                                                                                    levels = unique(DA.B_30h.FilClus_ClusAnnot.tmp$B_filtered_CCAInt_clusters) %>%
                                                                                      gtools::mixedsort(decreasing = T),
                                                                                    ordered = T))
                                       
                                       # Check the outcome
                                       head(DA.B_30h.FilClus_ClusAnnot.tmp)
                                       return(list(plotDAbeeswarm(DA.B_30h.FilClus_ClusAnnot.tmp,
                                                      group.by = "B_filtered_CCAInt_clusters", alpha = FDR_thres)+
                                         themeG+
                                         theme(panel.grid.major.y = element_line(color = adjustcolor( "lightgrey", alpha.f = 0.1),
                                                                                 linewidth = 1)),
                                         DA.B_30h.FilClus_ClusAnnot.tmp))})
# Change names of the list
names(plotBee_ClusAnnot_30h) <- unique(DA.B_30h.FilClus$Stimulation)
# Change names of the lists in the list, aka data and plot layer
for (layer in names(plotBee_ClusAnnot_30h)){
  names(plotBee_ClusAnnot_30h[[layer]])<-c("plot", "data")
}

# Plotting result
(plotBee_ClusAnnot_30h$IFNgamma$plot+labs(subtitle = "IFN") + plotBee_ClusAnnot_30h$ODN2006$plot + labs(subtitle = "ODN2006")) /
  (plotBee_ClusAnnot_30h$antiCD3CD28$plot + labs(subtitle = "anti-CD3/CD28") + plotBee_ClusAnnot_30h$MegaCD40L$plot +labs(subtitle = "CD40L"))+
  plot_annotation("30h") & labs(x = "Original clusters") & ylim(-10,10)
```

### Selection of interesting neighbourhoods - soft cut-offs per stimulation

#### Unstimulated neighbourhoods

```{r NBH_Selection_30h_UnstimulatedCells}
CutOff_Frac<-0.5
# Join all data of the beesswarms over the stimulations
plotBee_ClusAnnot_30h.data<-rbind(plotBee_ClusAnnot_30h$IFNgamma$data,
                                  plotBee_ClusAnnot_30h$MegaCD40L$data,
                                  plotBee_ClusAnnot_30h$antiCD3CD28$data,
                                  plotBee_ClusAnnot_30h$ODN2006$data)

### Unstimulated clusters/neighbourhood
# Check minimum and maximum LogFC
NBH_Selection_Unstim_30h.CutOffs <- plotBee_ClusAnnot_30h.data%>%
  filter(B_filtered_CCAInt_clusters %in% c(2,4))%>%
  group_by(Stimulation, B_filtered_CCAInt_clusters)%>%
  summarise(Min = min(logFC),
            Max = max(logFC),
            CutOff = min(logFC)*(1-CutOff_Frac))

NBH_Selection_Unstim_30h.CutOffs
```


```{r NBH_Overlap_Unstimulated}
# Filtering based on the cutoffs
NBH_Selection_Unstim_30h<-left_join(plotBee_ClusAnnot_30h.data,
                                NBH_Selection_Unstim_30h.CutOffs%>%
                                  select(Stimulation, B_filtered_CCAInt_clusters, CutOff))%>%
  filter(B_filtered_CCAInt_clusters %in% c(2,4),
         logFC < CutOff)

# Prepare UpsetPlotData
NBH_Selection_Unstim_30h.CombMat<-lapply(unique(NBH_Selection_Unstim_30h$Stimulation),
                                     FUN = function(stim){(NBH_Selection_Unstim_30h%>%filter(Stimulation == stim))[["Nhood"]]})
names(NBH_Selection_Unstim_30h.CombMat)<-unique(NBH_Selection_Unstim_30h$Stimulation)
NBH_Selection_Unstim_30h.CombMat<-make_comb_mat(NBH_Selection_Unstim_30h.CombMat, mode = "distinct")
# Plot upset plot
UpSet(NBH_Selection_Unstim_30h.CombMat,
      comb_order = order(comb_size(NBH_Selection_Unstim_30h.CombMat), decreasing = T),
      top_annotation = HeatmapAnnotation(
                "Intersection size" = anno_barplot(comb_size(NBH_Selection_Unstim_30h.CombMat),
                                                   border = FALSE,
                                                   gp = gpar(fill = "#595B5C",
                                                             lty = "blank"),
                                                   add_numbers = T,
                                                   height = unit(4, "cm")),
                annotation_name_side = "left",
                annotation_name_rot = 90),
      right_annotation = rowAnnotation(
                "DE genes" = anno_barplot(set_size(NBH_Selection_Unstim_30h.CombMat),
                                          border = FALSE,
                                          gp = gpar(fill = "#595B5C",
                                                    lty = "blank"),
                                          add_numbers = T,
                                          width = unit(4, "cm"))))

# Selected NBHs
NBH_Selection_Unstim_30h.Numbers<-sapply(X = c("1110", "1111", "1011", "1101"),
                                     FUN = function(set){extract_comb(NBH_Selection_Unstim_30h.CombMat, set)})%>%
  unlist()
length(NBH_Selection_Unstim_30h.Numbers)
```

#### Stimulated neighbourhoods

```{r NBH_Selection_30h_Stimulations, fig.height=10, fig.width=10}
### IFNgamma 
# Number of clusters plotted in the Beeswarm plot
NumClus_30h<-length(levels(plotBee_ClusAnnot_30h.data$B_filtered_CCAInt_clusters))
# Obtain the cutoff values per cluster and per Stimulation
# Only focusing on IFNgamma and ODN2006 stimulation
NBH_Selection_IFNgamma_30h.CutOffs <- plotBee_ClusAnnot_30h.data%>%
  filter(B_filtered_CCAInt_clusters %in% c(0, 1, 3, 7, 8, 9, 12, 14), 
         Stimulation %in% c("IFNgamma"))%>%
  group_by(Stimulation, B_filtered_CCAInt_clusters)%>%
  summarise(Min = min(logFC),
            Max = max(logFC),
            CutOff = max(logFC)*(1-CutOff_Frac))%>%
  mutate(positionHigh = NumClus_30h - as.numeric(as.character(B_filtered_CCAInt_clusters)) + 0.5,
         positionLow = NumClus_30h - as.numeric(as.character(B_filtered_CCAInt_clusters)) - 0.5)
# Check the obtained cutoffs
NBH_Selection_IFNgamma_30h.CutOffs%>%
  arrange(B_filtered_CCAInt_clusters)

### ODN2006
# Obtain the cutoff values per cluster and per Stimulation
# Only focusing on IFNgamma and ODN2006 stimulation
NBH_Selection_ODN2006_30h.CutOffs <- plotBee_ClusAnnot_30h.data%>%
  filter(B_filtered_CCAInt_clusters %in% c(0, 1, 3, 5, 6, 7, 8, 10, 15), 
         Stimulation %in% c("ODN2006"))%>%
  group_by(Stimulation, B_filtered_CCAInt_clusters)%>%
  summarise(Min = min(logFC),
            Max = max(logFC),
            CutOff = max(logFC)*(1-CutOff_Frac))%>%
  mutate(positionHigh = NumClus_30h - as.numeric(as.character(B_filtered_CCAInt_clusters)) + 0.5,
         positionLow = NumClus_30h - as.numeric(as.character(B_filtered_CCAInt_clusters)) - 0.5)
# Check the obtained cutoffs
NBH_Selection_ODN2006_30h.CutOffs%>%
  arrange(B_filtered_CCAInt_clusters)

### Anti-CD3/CD28
# Obtain the cutoff values per cluster and per Stimulation
# Only focusing on IFNgamma and ODN2006 stimulation
NBH_Selection_CD3CD28_30h.CutOffs <- plotBee_ClusAnnot_30h.data%>%
  filter(B_filtered_CCAInt_clusters %in% c(5, 6, 7, 11, 13), 
         Stimulation %in% c("antiCD3CD28"))%>%
  group_by(Stimulation, B_filtered_CCAInt_clusters)%>%
  summarise(Min = min(logFC),
            Max = max(logFC),
            CutOff = max(logFC)*(1-CutOff_Frac))%>%
  mutate(positionHigh = NumClus_30h - as.numeric(as.character(B_filtered_CCAInt_clusters)) + 0.5,
         positionLow = NumClus_30h - as.numeric(as.character(B_filtered_CCAInt_clusters)) - 0.5)
# Check the obtained cutoffs
NBH_Selection_CD3CD28_30h.CutOffs%>%
  arrange(B_filtered_CCAInt_clusters)

### MegaCD40L
# Obtain the cutoff values per cluster and per Stimulation
# Only focusing on IFNgamma and ODN2006 stimulation
NBH_Selection_MegaCD40L_30h.CutOffs <- plotBee_ClusAnnot_30h.data%>%
  filter(B_filtered_CCAInt_clusters %in% c(6, 11), 
         Stimulation %in% c("MegaCD40L"))%>%
  group_by(Stimulation, B_filtered_CCAInt_clusters)%>%
  summarise(Min = min(logFC),
            Max = max(logFC),
            CutOff = max(logFC)*(1-CutOff_Frac))%>%
  mutate(positionHigh = NumClus_30h - as.numeric(as.character(B_filtered_CCAInt_clusters)) + 0.5,
         positionLow = NumClus_30h - as.numeric(as.character(B_filtered_CCAInt_clusters)) - 0.5)
# Check the obtained cutoffs
NBH_Selection_MegaCD40L_30h.CutOffs%>%
  arrange(B_filtered_CCAInt_clusters)
```

#### Merging of cut-off values and selected NBH indication


```{r NBH_Selection_30h}
NBH_Selection_30h.CutOffs<-rbind(NBH_Selection_Unstim_30h.CutOffs,
                                 NBH_Selection_IFNgamma_30h.CutOffs,
                                 NBH_Selection_ODN2006_30h.CutOffs,
                                 NBH_Selection_MegaCD40L_30h.CutOffs,
                                 NBH_Selection_CD3CD28_30h.CutOffs)

# Add cutoffs to the data
plotBee_ClusAnnot_30h.data <- left_join(plotBee_ClusAnnot_30h.data,
                                    NBH_Selection_30h.CutOffs%>%
                                      select(CutOff, Stimulation, B_filtered_CCAInt_clusters),
                                    by = c("Stimulation", "B_filtered_CCAInt_clusters"))
# Indicate which neighbourhoods to include in downstream analysis
plotBee_ClusAnnot_30h.data <- plotBee_ClusAnnot_30h.data%>%
  # Neighbourhood is selected if logFC is larger then the cutoff (only for stimulated clusters)
  # OR is present in the selected NBH of the unstimulated cluster
  mutate(ToInclude = (logFC > CutOff) | (Nhood %in% NBH_Selection_Unstim_30h.Numbers))%>%
  mutate(ToInclude = replace_na(ToInclude, FALSE))
```


#### Visualization of cutoffs and selected neighbourhoods

```{r NBH_PlotBeeswarmWithCutoffs_Arrows_30h, fig.height=10, fig.width=10}
plotBee_ClusAnnot_30h$IFNgamma$plot+
  # add line segment
  geom_segment(data = NBH_Selection_IFNgamma_30h.CutOffs%>%
                 filter(Stimulation == "IFNgamma"),
               mapping = aes(x = positionHigh, xend = positionLow,
                             y = CutOff),
               color = "black", linewidth = 1, alpha = 0.5)+
  # add arrow next to the segment
  geom_segment(data = NBH_Selection_IFNgamma_30h.CutOffs%>%
                 filter(Stimulation == "IFNgamma"),
               mapping = aes(x = positionLow, 
                             y = CutOff, yend = CutOff+1),
               color = "black", linewidth = 1, alpha=0.5, 
               arrow = arrow(length = unit(0.03, "npc")))+
(plotBee_ClusAnnot_30h$ODN2006$plot+
  # add line segment
  geom_segment(data = NBH_Selection_ODN2006_30h.CutOffs%>%
                 filter(Stimulation == "ODN2006"),
               mapping = aes(x = positionHigh, xend = positionLow,
                             y = CutOff),
               color = "black", linewidth = 1, alpha = 0.5)+
  # add arrow next to the segment
  geom_segment(data = NBH_Selection_ODN2006_30h.CutOffs%>%
                 filter(Stimulation == "ODN2006"),
               mapping = aes(x = positionLow, 
                             y = CutOff, yend = CutOff+1),
               color = "black", linewidth = 1, alpha=0.5, 
               arrow = arrow(length = unit(0.03, "npc"))))+
(plotBee_ClusAnnot_30h$antiCD3CD28$plot+
  # add line segment
  geom_segment(data = NBH_Selection_CD3CD28_30h.CutOffs%>%
                 filter(Stimulation == "antiCD3CD28"),
               mapping = aes(x = positionHigh, xend = positionLow,
                             y = CutOff),
               color = "black", linewidth = 1, alpha = 0.5)+
  # add arrow next to the segment
  geom_segment(data = NBH_Selection_CD3CD28_30h.CutOffs%>%
                 filter(Stimulation == "antiCD3CD28"),
               mapping = aes(x = positionLow, 
                             y = CutOff, yend = CutOff+1),
               color = "black", linewidth = 1, alpha=0.5, 
               arrow = arrow(length = unit(0.03, "npc"))))+
(plotBee_ClusAnnot_30h$MegaCD40L$plot+
  # add line segment
  geom_segment(data = NBH_Selection_MegaCD40L_30h.CutOffs%>%
                 filter(Stimulation == "MegaCD40L"),
               mapping = aes(x = positionHigh, xend = positionLow,
                             y = CutOff),
               color = "black", linewidth = 1, alpha = 0.5)+
  # add arrow next to the segment
  geom_segment(data = NBH_Selection_MegaCD40L_30h.CutOffs%>%
                 filter(Stimulation == "MegaCD40L"),
               mapping = aes(x = positionLow, 
                             y = CutOff, yend = CutOff+1),
               color = "black", linewidth = 1, alpha=0.5, 
               arrow = arrow(length = unit(0.03, "native"))))+
  plot_annotation("30h") & labs(x = "Original clusters") & ylim(-10,10)
```

```{r NBH_PlotBeeswarmWithCutoffs_30h, fig.height=10, fig.width=10}
shapeValues = c("TRUE" = 19, "FALSE" = 1)
sizeValues = c("TRUE" = 1, "FALSE" = 0.5)

plotDAbeeswarm_SubsetIndication(plotBee_ClusAnnot_30h.data%>%filter(Stimulation == "IFNgamma"), 
                                group.by  = "B_filtered_CCAInt_clusters", alpha = FDR_thres, ToInclude = "ToInclude",
                                sizeValues = sizeValues, shapeValues = shapeValues)+
  labs(subtitle = "IFNgamma")+
plotDAbeeswarm_SubsetIndication(plotBee_ClusAnnot_30h.data%>%filter(Stimulation == "ODN2006"), 
                                group.by  = "B_filtered_CCAInt_clusters", alpha = FDR_thres, ToInclude = "ToInclude",
                                sizeValues = sizeValues, shapeValues = shapeValues)+
  labs(subtitle = "ODN2006")+
plotDAbeeswarm_SubsetIndication(plotBee_ClusAnnot_30h.data%>%filter(Stimulation == "antiCD3CD28"), 
                                group.by  = "B_filtered_CCAInt_clusters", alpha = FDR_thres, ToInclude = "ToInclude",
                                sizeValues = sizeValues, shapeValues = shapeValues)+
  labs(subtitle = "Anti-CD3/CD28")+
plotDAbeeswarm_SubsetIndication(plotBee_ClusAnnot_30h.data%>%filter(Stimulation == "MegaCD40L"), 
                                group.by  = "B_filtered_CCAInt_clusters", alpha = FDR_thres, ToInclude = "ToInclude",
                                sizeValues = sizeValues, shapeValues = shapeValues)+
  labs(subtitle = "MegaCD40L")+
  plot_annotation("30h") & labs(x = "Original clusters") & ylim(-10,10)&
  themeG &
  theme(panel.grid.major.y = element_line(color = adjustcolor( "lightgrey", alpha.f = 0.1),
                                          linewidth = 1)) & NoLegend()
```

```{r NBH_Retained_AbsoluteNumbers_30h}
## Retained number of NBHs per cluster
plotBee_ClusAnnot_30h.data%>%
  group_by(Stimulation, B_filtered_CCAInt_clusters)%>%
  summarise(RetainedNBH = sum(ToInclude))%>%
  filter(RetainedNBH != 0)%>%
  arrange(B_filtered_CCAInt_clusters)

## Retained number of NBHs per Stimulation
plotBee_ClusAnnot_30h.data%>%
  filter(!B_filtered_CCAInt_clusters %in% c(2,4))%>%
  group_by(Stimulation)%>%
  summarise(Retained_StimNBH = sum(ToInclude))
```

#### Extraction of the cell IDs of interest based on the selected NBHs

```{r ExtractionCellIDS_8h, echo=FALSE}
# Unstimulated
# Selection of cells linked to the selected NBHs
NBH_Selection_Unstim_30h.Cells<-sapply(as.numeric(NBH_Selection_Unstim_30h.Numbers),
                                   FUN = function(nhood)
                                     {nhoods(milo.B_30h.FilClus$milo_graph)@Dimnames[[1]][nhoods(milo.B_30h.FilClus$milo_graph)[,nhood]==1]})%>%
                                       unlist()%>%
                                       unique()

# Selection of the Nhoods of interest only linked to a stimulation
NBH_Selection_30h.Nhoods<-plotBee_ClusAnnot_30h.data%>%
  filter(ToInclude)%>%
  select(Nhood, Stimulation)
# Selection of cellIDS
for (nhood in unique(NBH_Selection_30h.Nhoods$Nhood)){
  if (nhood %in% NBH_Selection_Unstim_30h.Numbers){
    # Keep track of unstimulated Nhoods as well
    Condition <- "Unstimulated"
  } else {
    Condition <- "Stimulated"
  }
  ExtractedCells.tmp <- data.frame(Nhood = nhood,
                                   CellID = nhoods(milo.B_30h.FilClus$milo_graph)@Dimnames[[1]][nhoods(milo.B_30h.FilClus$milo_graph)[,nhood]==1],
                                   Condition = Condition)
  if (!exists("NBH_Selection_30h.Cells") || (dim(NBH_Selection_30h.Cells)[1]) == 0){
    message("Creating dataframe")
    NBH_Selection_30h.Cells <- ExtractedCells.tmp
  } else {
    message("\nJoining additional cells")
    NBH_Selection_30h.Cells <- rbind(NBH_Selection_30h.Cells,
                                 ExtractedCells.tmp)
  }
  # Clean environment
  rm(ExtractedCells.tmp)
}


# Joining the Nhood information to the Cell IDs
NBH_Selection_30h.Cells <- left_join(NBH_Selection_30h.Cells,
                                 NBH_Selection_30h.Nhoods,
                                 by = "Nhood")
# Annotate the cells that are overlapping with the unstimulated cells
NBH_Selection_30h.Cells <- NBH_Selection_30h.Cells%>%
  mutate(Condition = case_when(
    CellID %in% NBH_Selection_Unstim_30h.Cells ~ "Unstimulated",
    TRUE ~ "Stimulated"
  ))
```


#### Save selection of cells
```{r SaveSelectedCells_8h}
# Create a blank workbook
WorkBook <- createWorkbook()

# Add some sheets to the workbook
addWorksheet(WorkBook, "CutOffs")
addWorksheet(WorkBook, "Nhoods_Sel")
addWorksheet(WorkBook, "Cells_Sel")
addWorksheet(WorkBook, "UnstimCells")
addWorksheet(WorkBook, "OverviewData")

# Write the data to the sheets
writeData(WorkBook, sheet = "CutOffs", 
          x = NBH_Selection_30h.CutOffs%>%
            select(-positionHigh, -positionLow))
writeData(WorkBook, sheet = "Nhoods_Sel", 
          x = NBH_Selection_30h.Nhoods)
writeData(WorkBook, sheet = "Cells_Sel", 
          x = NBH_Selection_30h.Cells)
writeData(WorkBook, sheet = "UnstimCells",
          x = NBH_Selection_Unstim_30h.Cells)
writeData(WorkBook, sheet = "OverviewData",
          x = plotBee_ClusAnnot_30h.data)

# Export the file
saveWorkbook(WorkBook, 
             # overwrite = T,
             file.path(params$path_Results,
                            "MiloR",
                            paste(params$date2, "CutOffs_Selection_Nhoods_Cells", "30h.xlsx", sep = "_")))

``` 

```{r readInData_30h, eval=FALSE}
NBH_Selection_30h.Nhoods<-read.xlsx(xlsxFile = file.path(params$path_Results,
                                                         "MiloR",
                                                         paste(params$date2, "CutOffs_Selection_Nhoods_Cells", "30h.xlsx", sep = "_")),
                            sheet = "Nhoods_Sel")
NBH_Selection_30h.Cells<-read.xlsx(xlsxFile = file.path(params$path_Results,
                                                         "MiloR",
                                                         paste(params$date2, "CutOffs_Selection_Nhoods_Cells", "30h.xlsx", sep = "_")),
                            sheet = "Cells_Sel")
NBH_Selection_Unstim_30h.Cells<-read.xlsx(xlsxFile = file.path(params$path_Results,
                                                         "MiloR",
                                                         paste(params$date2, "CutOffs_Selection_Nhoods_Cells", "30h.xlsx", sep = "_")),
                            sheet = "UnstimCells")
plotBee_ClusAnnot_30h.data<-read.xlsx(xlsxFile = file.path(params$path_Results,
                                                         "MiloR",
                                                         paste(params$date2, "CutOffs_Selection_Nhoods_Cells", "30h.xlsx", sep = "_")),
                            sheet = "OverviewData")
```

### Selection of interesting neighbourhoods - hard cutoff over stimulations

#### Unstimulated neighbourhoods

```{r NBH_Selection_30h_UnstimulatedCells_FIXED}
# Define hard cutoff
CutOff.Fixed<-3
# Join all data of the beesswarms over the stimulations
plotBee_ClusAnnot_30h.Fixed.data<-rbind(plotBee_ClusAnnot_30h$IFNgamma$data,
                                        plotBee_ClusAnnot_30h$MegaCD40L$data,
                                        plotBee_ClusAnnot_30h$antiCD3CD28$data,
                                        plotBee_ClusAnnot_30h$ODN2006$data)
```

```{r NBH_Overlap_Unstimulated_30h_FIXED}
# Filtering based on the cutoffs
NBH_Selection_Unstim_30h.Fixed<-plotBee_ClusAnnot_30h.Fixed.data%>%
  mutate(CutOff_Fixed = -CutOff.Fixed)%>%
  filter(B_filtered_CCAInt_clusters %in% c(2,4),
         logFC < CutOff_Fixed)

# Prepare UpsetPlotData
NBH_Selection_Unstim_30h.Fixed.CombMat<-lapply(unique(NBH_Selection_Unstim_30h.Fixed$Stimulation),
                                           FUN = function(stim){(NBH_Selection_Unstim_30h.Fixed%>%filter(Stimulation == stim))[["Nhood"]]})
names(NBH_Selection_Unstim_30h.Fixed.CombMat)<-unique(NBH_Selection_Unstim_30h.Fixed$Stimulation)
NBH_Selection_Unstim_30h.Fixed.CombMat<-make_comb_mat(NBH_Selection_Unstim_30h.Fixed.CombMat, mode = "distinct")
# Plot upset plot
UpSet(NBH_Selection_Unstim_30h.Fixed.CombMat,
      comb_order = order(comb_size(NBH_Selection_Unstim_30h.Fixed.CombMat), decreasing = T),
      top_annotation = HeatmapAnnotation(
                "Intersection size" = anno_barplot(comb_size(NBH_Selection_Unstim_30h.Fixed.CombMat),
                                                   border = FALSE,
                                                   gp = gpar(fill = "#595B5C",
                                                             lty = "blank"),
                                                   add_numbers = T,
                                                   height = unit(4, "cm")),
                annotation_name_side = "left",
                annotation_name_rot = 90),
      right_annotation = rowAnnotation(
                "DE genes" = anno_barplot(set_size(NBH_Selection_Unstim_30h.Fixed.CombMat),
                                          border = FALSE,
                                          gp = gpar(fill = "#595B5C",
                                                    lty = "blank"),
                                          add_numbers = T,
                                          width = unit(4, "cm"))))

# Selected NBHs - overlap in at least 3 of the stimulatory conditions
NBH_Selection_Unstim_30h.Fixed.Numbers<-sapply(X = c("1110", "1111", "1011", "1101", "0111"),
                                     FUN = function(set){extract_comb(NBH_Selection_Unstim_30h.Fixed.CombMat, set)})%>%
  unlist()
length(NBH_Selection_Unstim_30h.Fixed.Numbers)

# Select the neighbourhoods based on the overlap
NBH_Selection_Unstim_30h.Fixed<-NBH_Selection_Unstim_30h.Fixed%>%
  mutate(ToInclude = case_when(
    Nhood %in% NBH_Selection_Unstim_30h.Fixed.Numbers ~ TRUE,
    TRUE ~ FALSE))
```

#### Stimulated neighbourhoods 

```{r NBH_Selection_30h_Stimulations_FIXED, fig.height=10, fig.width=10}
# Filtering based on the cutoffs
NBH_Selection_Stim_30h.Fixed<-plotBee_ClusAnnot_30h.Fixed.data%>%
  mutate(CutOff_Fixed = CutOff.Fixed)%>%
  filter(!B_filtered_CCAInt_clusters %in% c(2,4),
         logFC > CutOff_Fixed)

# Indicate which NBHs to include
NBH_Selection_Stim_30h.Fixed<-NBH_Selection_Stim_30h.Fixed%>%
  mutate(ToInclude = case_when(
    Stimulation == "IFNgamma" & B_filtered_CCAInt_clusters %in% c(0, 1, 3, 7, 8, 9, 12, 14) ~ TRUE,
    Stimulation == "ODN2006" & B_filtered_CCAInt_clusters %in% c(0, 1, 3, 5, 6, 7, 8, 10, 12, 15)~ TRUE,
    Stimulation == "antiCD3CD28" & B_filtered_CCAInt_clusters %in% c(0, 1, 3, 5, 6, 7, 9, 11, 13) ~ TRUE,
    Stimulation == "MegaCD40L" & B_filtered_CCAInt_clusters %in% c(6, 11)~ TRUE,
    TRUE ~ FALSE
  ))

# Check the number of retained neighbourhoods per stimulation and cluster
table(NBH_Selection_Stim_30h.Fixed$Stimulation,
      NBH_Selection_Stim_30h.Fixed$B_filtered_CCAInt_clusters)
# Check the number of retained neighbourhoods per stimulation
table(NBH_Selection_Stim_30h.Fixed$Stimulation,
      NBH_Selection_Stim_30h.Fixed$ToInclude)
```

#### Merging of cut-off values and selected NBH indication


```{r NBH_Selection_8h}
NBH_Selection_30h.Fixed<-rbind(NBH_Selection_Unstim_30h.Fixed,
                           NBH_Selection_Stim_30h.Fixed)

# Add cutoffs to the data
plotBee_ClusAnnot_30h.Fixed.data <- left_join(plotBee_ClusAnnot_30h.Fixed.data,
                                    NBH_Selection_30h.Fixed%>%
                                      select(CutOff_Fixed, Stimulation, Nhood, ToInclude),
                                    by = c("Stimulation", "Nhood"))
# Indicate which neighbourhoods to include in downstream analysis
plotBee_ClusAnnot_30h.Fixed.data <- plotBee_ClusAnnot_30h.Fixed.data%>%
  # Neighbourhood is selected if logFC is larger then the cutoff (only for stimulated clusters)
  # OR is present in the selected NBH of the unstimulated cluster
  mutate(ToInclude = replace_na(ToInclude, FALSE))
```

#### Visualization of cutoffs and selected neighbourhoods

```{r NBH_PlotBeeswarmWithCutoffs_8h_FIXED, fig.height=10, fig.width=10}
shapeValues = c("TRUE" = 19, "FALSE" = 1)
sizeValues = c("TRUE" = 1, "FALSE" = 0.5)

plotDAbeeswarm_SubsetIndication(plotBee_ClusAnnot_30h.Fixed.data%>%filter(Stimulation == "IFNgamma"), 
                                group.by  = "B_filtered_CCAInt_clusters", alpha = FDR_thres, ToInclude = "ToInclude",
                                sizeValues = sizeValues, shapeValues = shapeValues)+
  labs(subtitle = "IFNgamma")+
plotDAbeeswarm_SubsetIndication(plotBee_ClusAnnot_30h.Fixed.data%>%filter(Stimulation == "ODN2006"), 
                                group.by  = "B_filtered_CCAInt_clusters", alpha = FDR_thres, ToInclude = "ToInclude",
                                sizeValues = sizeValues, shapeValues = shapeValues)+
  labs(subtitle = "ODN2006")+
plotDAbeeswarm_SubsetIndication(plotBee_ClusAnnot_30h.Fixed.data%>%filter(Stimulation == "antiCD3CD28"), 
                                group.by  = "B_filtered_CCAInt_clusters", alpha = FDR_thres, ToInclude = "ToInclude",
                                sizeValues = sizeValues, shapeValues = shapeValues)+
  labs(subtitle = "Anti-CD3/CD28")+
plotDAbeeswarm_SubsetIndication(plotBee_ClusAnnot_30h.Fixed.data%>%filter(Stimulation == "MegaCD40L"), 
                                group.by  = "B_filtered_CCAInt_clusters", alpha = FDR_thres, ToInclude = "ToInclude",
                                sizeValues = sizeValues, shapeValues = shapeValues)+
  labs(subtitle = "MegaCD40L")+
  plot_annotation("8h") & labs(x = "Original clusters") & ylim(-10,10)&
  themeG &
  theme(panel.grid.major.y = element_line(color = adjustcolor( "lightgrey", alpha.f = 0.1),
                                          linewidth = 1)) &  NoLegend()
```

#### Extraction of the cell IDs of interest based on the selected NBHs

```{r ExtractionCellIDS_30h_FIXED, echo=FALSE}
# Unstimulated
# Selection of cells linked to the selected NBHs
NBH_Selection_Unstim_30h.Fixed.Cells<-sapply(as.numeric(NBH_Selection_Unstim_30h.Fixed.Numbers),
                                   FUN = function(nhood){nhoods(milo.B_30h.FilClus$milo_graph)@Dimnames[[1]][nhoods(
                                     milo.B_30h.FilClus$milo_graph)[,nhood]==1]})%>%
                                       unlist()%>%
                                       unique()

# Selection of the Nhoods of interest only linked to a stimulation
NBH_Selection_30h.Fixed.Nhoods<-plotBee_ClusAnnot_30h.Fixed.data%>%
  filter(ToInclude)%>%
  select(Nhood, Stimulation)
# Selection of cellIDS
for (nhood in unique(NBH_Selection_30h.Fixed.Nhoods$Nhood)){
  if (nhood %in% NBH_Selection_Unstim_30h.Fixed.Numbers){
    # Keep track of unstimulated Nhoods as well
    Condition <- "Unstimulated"
  } else {
    Condition <- "Stimulated"
  }
  ExtractedCells.tmp <- data.frame(Nhood = nhood,
                                   CellID = nhoods(milo.B_30h.FilClus$milo_graph)@Dimnames[[1]][nhoods(
                                     milo.B_30h.FilClus$milo_graph)[,nhood]==1],
                                   Condition = Condition)
  if (!exists("NBH_Selection_30h.Fixed.Cells") || (dim(NBH_Selection_30h.Fixed.Cells)[1]) == 0){
    message("Creating dataframe")
    NBH_Selection_30h.Fixed.Cells <- ExtractedCells.tmp
  } else {
    message("\nJoining additional cells")
    NBH_Selection_30h.Fixed.Cells <- rbind(NBH_Selection_30h.Fixed.Cells,
                                 ExtractedCells.tmp)
  }
  # Clean environment
  rm(ExtractedCells.tmp)
}

# Joining the Nhood information to the Cell IDs
NBH_Selection_30h.Fixed.Cells <- left_join(NBH_Selection_30h.Fixed.Cells,
                                 NBH_Selection_30h.Fixed.Nhoods,
                                 by = "Nhood")
# Annotate the cells that are overlapping with the unstimulated cells
NBH_Selection_30h.Fixed.Cells <- NBH_Selection_30h.Fixed.Cells%>%
  mutate(Condition = case_when(
    CellID %in% NBH_Selection_Unstim_30h.Fixed.Cells ~ "Unstimulated",
    TRUE ~ "Stimulated"
  ))
```

#### Save selection of cells
```{r SaveSelectedCells_8h}
# Create a blank workbook
WorkBook <- createWorkbook()

# Add some sheets to the workbook
addWorksheet(WorkBook, "Nhoods_Sel")
addWorksheet(WorkBook, "Cells_Sel")
addWorksheet(WorkBook, "UnstimCells")
addWorksheet(WorkBook, 'OverviewData')

# Write the data to the sheets
writeData(WorkBook, sheet = "Nhoods_Sel", 
          x = NBH_Selection_30h.Fixed.Nhoods)
writeData(WorkBook, sheet = "Cells_Sel", 
          x = NBH_Selection_30h.Fixed.Cells)
writeData(WorkBook, sheet = "UnstimCells",
          x = NBH_Selection_Unstim_30h.Fixed.Cells)
writeData(WorkBook, sheet = "OverviewData",
          x = plotBee_ClusAnnot_30h.Fixed.data)

# Export the file
saveWorkbook(WorkBook, 
             overwrite = T,
             file.path(params$path_Results,
                            "MiloR",
                            paste(params$date3, "HardCutOff3_Selection_Nhoods_Cells", "30h.xlsx", sep = "_")))
```

```{r ReadInData_SelectionCells_Fixed_8h}
NBH_Selection_30h.Fixed.Nhoods <- read.xlsx(file.path(params$path_Results,
                            "MiloR",
                            paste(params$date3, "HardCutOff3_Selection_Nhoods_Cells", "30h.xlsx", sep = "_")),
                            sheet = "Nhoods_Sel")
NBH_Selection_30h.Fixed.Cells <- read.xlsx(file.path(params$path_Results,
                            "MiloR",
                            paste(params$date3, "HardCutOff3_Selection_Nhoods_Cells", "30h.xlsx", sep = "_")),
                            sheet = "Cells_Sel")
NBH_Selection_Unstim_30h.Fixed.Cells <- read.xlsx(file.path(params$path_Results,
                            "MiloR",
                            paste(params$date3, "HardCutOff3_Selection_Nhoods_Cells", "30h.xlsx", sep = "_")),
                            sheet = "UnstimCells")
plotBee_ClusAnnot_30h.Fixed.data <- read.xlsx(file.path(params$path_Results,
                            "MiloR",
                            paste(params$date3, "HardCutOff3_Selection_Nhoods_Cells", "30h.xlsx", sep = "_")),
                            sheet = "OverviewData")
```

#### Comparison to selection based on soft cut-offs

```{r ComparisonCutOffs_30h}
# For stimulated cells
for(stim in unique(plotBee_ClusAnnot_30h.data$Stimulation)){
  # Prepare plotdata
  CompPlotdata<-list("SOFT" = (plotBee_ClusAnnot_30h.data%>%filter(Stimulation == stim,
                                                               !B_filtered_CCAInt_clusters %in% c(2,4),
                                                               ToInclude == TRUE))[["Nhood"]],
                     "HARD" = (plotBee_ClusAnnot_30h.Fixed.data%>%filter(Stimulation == stim,
                                                               !B_filtered_CCAInt_clusters %in% c(2,4),
                                                               ToInclude == TRUE))[["Nhood"]])
  CompPlotdata<-make_comb_mat(CompPlotdata, mode = "distinct")
  
  # Plot upsetplot
  print(UpSet(CompPlotdata,
        comb_order = order(comb_size(CompPlotdata), decreasing = T),
        top_annotation = HeatmapAnnotation(
                  "Intersection size" = anno_barplot(comb_size(CompPlotdata),
                                                     border = FALSE,
                                                     gp = gpar(fill = "#595B5C",
                                                               lty = "blank"),
                                                     add_numbers = T,
                                                     height = unit(4, "cm")),
                  annotation_name_side = "left",
                  annotation_name_rot = 90),
        right_annotation = rowAnnotation(
                  "DE genes" = anno_barplot(set_size(CompPlotdata),
                                            border = FALSE,
                                            gp = gpar(fill = "#595B5C",
                                                      lty = "blank"),
                                            add_numbers = T,
                                            width = unit(4, "cm"))),
        row_title = stim))
  rm(CompPlotdata)
}

# For unstimulated cells
# Prepare plotdata
CompPlotdata<-list("SOFT" = (plotBee_ClusAnnot_30h.data%>%filter(Stimulation == stim,
                                                             B_filtered_CCAInt_clusters %in% c(2,4),
                                                             ToInclude == TRUE))[["Nhood"]],
                   "HARD" = (plotBee_ClusAnnot_30h.Fixed.data%>%filter(Stimulation == stim,
                                                             B_filtered_CCAInt_clusters %in% c(2,4),
                                                             ToInclude == TRUE))[["Nhood"]])
CompPlotdata<-make_comb_mat(CompPlotdata, mode = "distinct")

# Plot upsetplot
print(UpSet(CompPlotdata,
      comb_order = order(comb_size(CompPlotdata), decreasing = T),
      top_annotation = HeatmapAnnotation(
                "Intersection size" = anno_barplot(comb_size(CompPlotdata),
                                                   border = FALSE,
                                                   gp = gpar(fill = "#595B5C",
                                                             lty = "blank"),
                                                   add_numbers = T,
                                                   height = unit(4, "cm")),
                annotation_name_side = "left",
                annotation_name_rot = 90),
      right_annotation = rowAnnotation(
                "DE genes" = anno_barplot(set_size(CompPlotdata),
                                          border = FALSE,
                                          gp = gpar(fill = "#595B5C",
                                                    lty = "blank"),
                                          add_numbers = T,
                                          width = unit(4, "cm"))),
      row_title = "Unstimulated"))
rm(CompPlotdata)
```

```{r ComparisonCutOffs_Cells_30h}
# For stimulated cells
for(stim in unique(NBH_Selection_30h.Cells$Stimulation)){
  # Prepare plotdata
  CompPlotdata<-list("SOFT" = (NBH_Selection_30h.Cells%>%filter(Stimulation == stim,
                                                               Condition == "Stimulated"))[["CellID"]]%>%unique(),
                     "HARD" = (NBH_Selection_30h.Fixed.Cells%>%filter(Stimulation == stim,
                                                               Condition == "Stimulated"))[["CellID"]]%>%unique())
  CompPlotdata<-make_comb_mat(CompPlotdata, mode = "distinct")
  
  # Plot upsetplot
  print(UpSet(CompPlotdata,
        comb_order = order(comb_size(CompPlotdata), decreasing = T),
        top_annotation = HeatmapAnnotation(
                  "Intersection size" = anno_barplot(comb_size(CompPlotdata),
                                                     border = FALSE,
                                                     gp = gpar(fill = "#595B5C",
                                                               lty = "blank"),
                                                     add_numbers = T,
                                                     height = unit(4, "cm")),
                  annotation_name_side = "left",
                  annotation_name_rot = 90),
        right_annotation = rowAnnotation(
                  "DE genes" = anno_barplot(set_size(CompPlotdata),
                                            border = FALSE,
                                            gp = gpar(fill = "#595B5C",
                                                      lty = "blank"),
                                            add_numbers = T,
                                            width = unit(4, "cm"))),
        row_title = stim))
  rm(CompPlotdata)
}

# For unstimulated cells
# Prepare plotdata
CompPlotdata<-list("SOFT" = (NBH_Selection_30h.Cells%>%filter(Condition == "Unstimulated"))[["CellID"]]%>%unique(),
                   "HARD" = (NBH_Selection_30h.Fixed.Cells%>%filter(Condition == "Unstimulated"))[["CellID"]]%>%unique())
CompPlotdata<-make_comb_mat(CompPlotdata, mode = "distinct")

# Plot upsetplot
print(UpSet(CompPlotdata,
      comb_order = order(comb_size(CompPlotdata), decreasing = T),
      top_annotation = HeatmapAnnotation(
                "Intersection size" = anno_barplot(comb_size(CompPlotdata),
                                                   border = FALSE,
                                                   gp = gpar(fill = "#595B5C",
                                                             lty = "blank"),
                                                   add_numbers = T,
                                                   height = unit(4, "cm")),
                annotation_name_side = "left",
                annotation_name_rot = 90),
      right_annotation = rowAnnotation(
                "DE genes" = anno_barplot(set_size(CompPlotdata),
                                          border = FALSE,
                                          gp = gpar(fill = "#595B5C",
                                                    lty = "blank"),
                                          add_numbers = T,
                                          width = unit(4, "cm"))),
      row_title = "Unstimulated"))
rm(CompPlotdata)
```

