---
title: "Filtering of B cells - Visualization - Differential gene expression"
author: "Tine D'hamers"
date: "2025-06-24"
output: html_document
params:
  path_Results: "/kyukon/scratch/gent/vo/002/gvo00224/TOBI/Projects/EProfile/Results/Exp_202211_HCSamples_CEV017-22/002_NoBaseline/202506_Filtered_BCells"
  SeuratObj_Bcells_8h: "/kyukon/scratch/gent/vo/002/gvo00224/TOBI/Projects/EProfile/Results/Exp_202211_HCSamples_CEV017-22/002_NoBaseline/SeuratObjects/seuratObj_NoBase_Integrated_BClusters_8h.rds"
  SeuratObj_Bcells_30h: "/kyukon/scratch/gent/vo/002/gvo00224/TOBI/Projects/EProfile/Results/Exp_202211_HCSamples_CEV017-22/002_NoBaseline/SeuratObjects/seuratObj_NoBase_Integrated_BClusters_30h.rds"
  date: "20250625"
  date2: "20250731"
---

# Introduction

In this RMarkdown file, multiple analysis steps were performed.

1. The final B cell subset obtained in previous scripts (e.g. as loaded in `007_NoBase_DEGeneAnalysis.Rmd`) is filtered to remove the weird blubs on top of the UMAP for 8 hours of stimulation. A similar check was performed for the B cells obtained after 30 hours of incubation.
2. The obtained filtered B cell subset was used for a final differential gene expression analysis

Results were stored in appropriate files

# Preparation

## Library loading

```{r Libraries}
.libPaths("/kyukon/data/gent/vo/000/gvo00027/SingleCell10X/seuratintegrate/Rpackages/R_4.2.1-foss-2022a/")
# Single cell data manipulation
library(Seurat)
# General data manipulation
library(tidyr)
library(dplyr)
library(tibble)
library(xlsx)
library(openxlsx)
# Plot generation
library(ggplot2)
library(ggpubr)
library(ggrepel)
library(ggh4x)
library(RColorBrewer)
library(ComplexHeatmap)
library(circlize)
library(ggridges)
library(patchwork)
# Additional analysis
library(clustree)
# Gene set enrichment
library(msigdbr)
# NCBI scraping
library(rentrez)
``` 

## Aesthetics and function loading

```{r Aesthetics}
# Themes
source("../../PlotThemes_sc_CVID.R")
colorsStim<-c(colorsStim, c("CD3/CD28" = colorsStim[["CD3/CD28_beads"]]))

options(bitmapType='cairo')

source("../../Functions_sc_CVID.R")

# Coloring for the DE genes
colDirection<-c("DOWN" = "darkred",
                "UP" = "darkgreen",
                "MIXED" = "#D5BBB1",
                "NS" = "grey80")
grid.grey = "grey95"
```

## Dataset loading

```{r LoadingData}
# Seurat objects B cells
IntData.B_8h<-readRDS(params$SeuratObj_Bcells_8h)
IntData.B_30h<-readRDS(params$SeuratObj_Bcells_30h)
```

# Visualization of B cell subsets

First, the original B cell dataset is visualized for both 8 and 30 hours of stimulation.

```{r UMAP_Visualization_PreFiltering, fig.width=20, fig.height=6}
DimPlot(object = IntData.B_8h,
        reduction = "B_integrated.umap.rna",
        group.by = "B_clusters",
        raster = F, label = T)+
    themeG+
    labs(x = "RNA_UMAP1", y = "RNA_UMAP2",
         title = "B cell clusters")+
    theme(panel.border = element_rect(colour = "grey60", fill=NA, linewidth=0.5),
          axis.ticks = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank()) + 
  (DimPlot(object = IntData.B_8h,
        reduction = "B_integrated.umap.rna",
        group.by = "predicted.celltype.l1",
        raster = F, label = F)+
    themeG+
    labs(x = "RNA_UMAP1", y = "RNA_UMAP2",
         title = "Cell types predictions")+
    theme(panel.border = element_rect(colour = "grey60", fill=NA, linewidth=0.5),
          axis.ticks = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank()) +
  scale_color_manual(values = colorsCellType)) +
     (DimPlot(object = IntData.B_8h,
        reduction = "B_integrated.umap.rna",
        group.by = "predicted.celltype.l2",
        raster = F, label = F)+
    themeG+
    labs(x = "RNA_UMAP1", y = "RNA_UMAP2",
         title = "Cell types predictions")+
    theme(panel.border = element_rect(colour = "grey60", fill=NA, linewidth=0.5),
          axis.ticks = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank()))

DimPlot(object = IntData.B_30h,
        reduction = "B_integrated.umap.rna",
        group.by = "B_clusters",
        raster = F, label = T)+
    themeG+
    labs(x = "RNA_UMAP1", y = "RNA_UMAP2",
         title = "B cell clusters")+
    theme(panel.border = element_rect(colour = "grey60", fill=NA, linewidth=0.5),
          axis.ticks = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank())+
   (DimPlot(object = IntData.B_30h,
        reduction = "B_integrated.umap.rna",
        group.by = "predicted.celltype.l1",
        raster = F, label = F)+
    themeG+
    labs(x = "RNA_UMAP1", y = "RNA_UMAP2",
         title = "Cell types predictions")+
    theme(panel.border = element_rect(colour = "grey60", fill=NA, linewidth=0.5),
          axis.ticks = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank()) +
  scale_color_manual(values = colorsCellType)) +
   (DimPlot(object = IntData.B_30h,
        reduction = "B_integrated.umap.rna",
        group.by = "predicted.celltype.l2",
        raster = F, label = F)+
    themeG+
    labs(x = "RNA_UMAP1", y = "RNA_UMAP2",
         title = "Cell types predictions")+
    theme(panel.border = element_rect(colour = "grey60", fill=NA, linewidth=0.5),
          axis.ticks = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank()))
```

# Filtering of the confusing clusters

## 8 hours of stimulation

For the B cell subset obtained after 8 hours of stimulation, a confusing blub mainly overlapping with **cluster 17** was removed. Next, the plasmablast cells (**cluster 21**) were removed as well since this a really small cluster in this B cell subset. Furthermore, an additional dimensionality reduction step, generating a new UMAP, was performed. This is the final B cell subset that was used for all the downstream analysis steps.

```{r Filtering_UMAPGeneration_8h, fig.width=6, fig.height=4}
# Change idents needed in the subset function
Idents(IntData.B_8h) <- "B_clusters"
# Subsetting the data
IntData.B_8h.FilCus <- subset(IntData.B_8h,
                              # Only extract the clusters without cluster 17 and 22
                              idents = unique(levels(IntData.B_8h$B_clusters))[!unique(levels(IntData.B_8h$B_clusters)) %in% c("17", "22")])
# Visualization to check if clusters were proberly removed
DimPlot(object = IntData.B_8h.FilCus,
        reduction = "B_integrated.umap.rna",
        group.by = "B_clusters",
        raster = F, label = T)+
    themeG+
    labs(x = "RNA_UMAP1", y = "RNA_UMAP2",
         title = "B cell clusters")+
    theme(panel.border = element_rect(colour = "grey60", fill=NA, linewidth=0.5),
          axis.ticks = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank())
```


### Dimensionality reduction without data integration

In a first step, dimensionality reduction was performe in the standard way, without taking into account any integration of the data. This is performed by generating PC of interest and clustering based on these PCs. Next, a UMAP was generated using the most interesting PCs (using an albow plot to decide on this number).

```{r DimRed_NoIntegration, fig.width=15, fig.height=4}
# Generation of new UMAP 
## Run PCA analysis for dimensionality reduction
IntData.B_8h.FilCus <- Seurat::RunPCA(object = IntData.B_8h.FilCus,
                                      # Based on RNA only
                                      assay = "RNA",
                                      # Name to save
                                      reduction.name = "B_filtered.pca")
## Check interesting number of dimensions in pca dimensionality reduction
ElbowPlot(IntData.B_8h.FilCus, ndims = 50, reduction = "B_filtered.pca")

## Generation of clusters
IntData.B_8h.FilCus <- FindNeighbors(IntData.B_8h.FilCus, 
                                reduction = "B_filtered.pca", 
                                dims = 1:30, graph.name = "B_filtered")
### Clustering using multiple resolution values
for (val_res in seq(0.5,2.5, by=0.5)){
  ResVal<-paste0("B_filtered_res_", val_res)
  message(paste0("Busy working on clustering with resolution ", val_res))
  IntData.B_8h.FilCus<-FindClusters(IntData.B_8h.FilCus,
                                    resolution=val_res,
                        verbose = FALSE, cluster.name = ResVal,
                        graph.name = "B_filtered")}
# Clustree to check the select the appropriate resolution
clustree(IntData.B_8h.FilCus, 
         # Select only the correct clusters based on the resolution
         prefix = "B_filtered_res_",
         # Size of the nodes
         node_size = 6,
         # Edges cross as little a possible
         layout = "sugiyama",
         # Coloring the nodes based on fraction of T cells
         # Per cluster 
         node_colour = "lightblue",
         show_axis = TRUE)+
  scale_color_gradient(low = "lightgrey",
                       high = "#A00D0D")+
  scale_edge_color_continuous(low = "lightgrey",
                             high = "#690F82")+
  labs(edge_alpha = "Proportion moved cells",
       edge_colour = "Counts moved cells",
       colour = "B cell fraction",
       title = "B filtered clusters",
       subtitle = "RNA")+
  ylab("Resolution")+
  theme_clustree

## Resolution of 1.5 seems okay 
## Better to overcluster, merging clusters is possible if necessary
Res_Method<-c("Resolution" = 1.5)
Res_Method[]<-paste0("B_filtered_res_",Res_Method)
IntData.B_8h.FilCus<-SelectMetaData(object = IntData.B_8h.FilCus,
                               to_stay = c("B_filtered_clusters" = Res_Method[["Resolution"]]),
                               pattern_remove = "B_filtered_res_")

## RUN UMAP dimensionality reduction based on the previous PCA analysis
IntData.B_8h.FilCus <- RunUMAP(IntData.B_8h.FilCus, 
                    reduction = "B_filtered.pca", 
                    # Focus on the first 30 PC based on the elbow plot generated above
                    dims = 1:30, reduction.name = "B_filtered.umap")

# Visualization of the final UMAP and its clusters
DimPlot(object = IntData.B_8h.FilCus,
        reduction = "B_filtered.umap",
        group.by = "B_filtered_clusters",
        raster = F, label = T)+
    themeG+
    labs(x = "RNA_UMAP1", y = "RNA_UMAP2",
         title = "B cell clusters")+
    theme(panel.border = element_rect(colour = "grey60", fill=NA, linewidth=0.5),
          axis.ticks = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank()) + 
  (DimPlot(object = IntData.B_8h.FilCus,
        reduction = "B_filtered.umap",
        group.by = "Stimulation",
        raster = F, label = F)+
    themeG+
    labs(x = "RNA_UMAP1", y = "RNA_UMAP2",
         title = "B cell clusters")+
    theme(panel.border = element_rect(colour = "grey60", fill=NA, linewidth=0.5),
          axis.ticks = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank())+
  scale_color_manual(values = colorsStim))+
  (DimPlot(object = IntData.B_8h.FilCus,
        reduction = "B_filtered.umap",
        group.by = "predicted.celltype.l1",
        raster = F, label = F)+
    themeG+
    labs(x = "RNA_UMAP1", y = "RNA_UMAP2",
         title = "B cell clusters")+
    theme(panel.border = element_rect(colour = "grey60", fill=NA, linewidth=0.5),
          axis.ticks = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank())+
     scale_color_manual(values = colorsCellType))

# Save the plot
ggsave(filename = file.path(params$path_Results, paste0(params$date, 
                                                        "_FilteredB_8h_UMAP_clusters_stimulations_annotation.pdf")),
       device = "pdf", height = 4, width = 15,
       dpi = 330)
```

```{r DensityPlots_Stimulations_8h, fig.height=10, fig.width=12}
Idents(IntData.B_8h.FilCus)<-'Stimulation'
# Initialization of list for plots
Dens.plotlist <- list()
# Generate the density plots and save them in the list
for (stim in unique(IntData.B_8h.FilCus$Stimulation)){
  Dens.plotlist[[stim]] <- DimPlot(IntData.B_8h.FilCus, reduction = "B_filtered.umap",
                                 group.by = "Stimulation", cols = rep("lightgrey", times = 5))+
    themeG+
    theme(legend.position = "none")+
    stat_density2d(data = subset(IntData.B_8h.FilCus, idents = stim)@reductions[["B_filtered.umap"]]@cell.embeddings,
                   mapping = aes(x = Bfilteredumap_1,
                                 y = Bfilteredumap_2,
                                 fill = ..level..),
                   geom = "polygon", alpha = 0.6)+
    scale_fill_gradient(low = "lightblue",high = "darkblue", guide = "colourbar",
                        breaks = c(0,0.05, 0.1))+
    labs(title = stim, x = "UMAP1", y = "UMAP2")+
    theme(axis.line = element_blank(),
          axis.ticks = element_blank(),
          axis.text = element_blank(),
          axis.title = element_blank())+
    ylim(-7,7)+
    xlim(-8,8)
}

# Join plots
ggarrange(plotlist = Dens.plotlist, ncol = 3, nrow = 2)
# Save plot
ggsave(filename = file.path(params$path_Results, paste0(params$date, 
                                                        "_FilteredB_8h_UMAP_density_stimulations.pdf")),
       device = "pdf", height = 10, width = 12,
       dpi = 330)

# Clean
rm(Dens.plotlist)
```

### Dimensionality reduction with data integration - CCA

In this part, the filtered B cell subset is first integrated using the CCA approach. Additionally, clusters are generated and a UMAP dimensionality reduction is performed for downstream visualization.

```{r DimRed_CCAintegration, fig.width=15, fig.height=4}
#### INTEGRATION ####
# Set the default assay to RNA
DefaultAssay(IntData.B_8h.FilCus) <- "RNA"

# Split the data according the sampleID (batch effects)
IntData.B_8h.FilCus[["RNA"]]<-split(IntData.B_8h.FilCus[["RNA"]],
                                    IntData.B_8h$samplename)
# Normalization
IntData.B_8h.FilCus<-Seurat::NormalizeData(IntData.B_8h.FilCus)
# Variable Features
IntData.B_8h.FilCus<-Seurat::FindVariableFeatures(IntData.B_8h.FilCus)
# Scaling
IntData.B_8h.FilCus<-Seurat::ScaleData(IntData.B_8h.FilCus, 
                                       features = rownames(IntData.B_8h.FilCus))
# Dimensionality reduction (unintegrated)
IntData.B_8h.FilCus<-Seurat::RunPCA(IntData.B_8h.FilCus, assay = "RNA", 
                                    reduction.name = "B_filtered_pca.priorIntegration.rna")
IntData.B_8h.FilCus<-Seurat::FindNeighbors(IntData.B_8h.FilCus, dims = 1:20, 
                                           reduction = "B_filtered_pca.priorIntegration.rna")
IntData.B_8h.FilCus<-Seurat::FindClusters(IntData.B_8h.FilCus, resolution = 2, 
                                          cluster.name = "B_filtered_pca_clusters.priorIntegration.rna")
IntData.B_8h.FilCus<-Seurat::RunUMAP(IntData.B_8h.FilCus, dims = 1:20, reduction = "B_filtered_pca.priorIntegration.rna", 
                           reduction.name = "B_filtered_umap.priorIntegration.rna")
# CCA integration of RNA layer
IntData.B_8h.FilCus<-IntegrateLayers(IntData.B_8h.FilCus, method = CCAIntegration,
                           orig.reduction = "B_filtered_pca.priorIntegration.rna", 
                           new.reduction = "B_filtered_CCAInt.rna",
                           verbose = F)

# Joining layers
IntData.B_8h.FilCus[["RNA"]] <- JoinLayers(IntData.B_8h.FilCus[["RNA"]])

#### CLUSTERING ####
# Clustering based on the integrated data
IntData.B_8h.FilCus <- FindNeighbors(IntData.B_8h.FilCus, 
                           reduction = "B_filtered_CCAInt.rna", 
                           dims = 1:30, graph.name = "B_filtered_CCAInt.rna_graph")
# Clustering using multiple resolution values
for (val_res in seq(0.5,2.5, by=0.5)){
  ResVal<-paste0("B_filtered_CCAInt_clusters_res_", val_res)
  message(paste0("Busy working on B cell clustering with resolution ", val_res))
  IntData.B_8h.FilCus<-FindClusters(IntData.B_8h.FilCus,resolution=val_res,
                          verbose = FALSE, cluster.name = ResVal,
                          graph.name = "B_filtered_CCAInt.rna_graph")}
# Clustree to check the select the appropriate resolution
clustree(IntData.B_8h.FilCus, 
         # Select only the correct clusters based on the resolution
         prefix = "B_filtered_CCAInt_clusters_res_",
         # Size of the nodes
         node_size = 6,
         # Edges cross as little a possible
         layout = "sugiyama",
         # Coloring the nodes based on fraction of B cells
         # Per cluster 
         node_colour = "lightblue",
         show_axis = TRUE)+
  scale_color_gradient(low = "lightgrey",
                       high = "#A00D0D")+
  scale_edge_color_continuous(low = "lightgrey",
                              high = "#690F82")+
  labs(edge_alpha = "Proportion moved cells",
       edge_colour = "Counts moved cells",
       colour = "B cell fraction",
       title = "Singlets B cells",
       subtitle = "RNA")+
  ylab("Resolution")+
  theme_clustree

# Resolution of 1.5 seems okay 
# Better to overcluster, merging clusters is possible if necessary
Res_Method<-c("Resolution" = 1.5)
Res_Method[]<-paste0("B_filtered_CCAInt_clusters_res_",Res_Method)

IntData.B_8h.FilCus<-SelectMetaData(object = IntData.B_8h.FilCus,
                          to_stay = c("B_filtered_CCAInt_clusters" = Res_Method[["Resolution"]]),
                          pattern_remove = "B_filtered_CCAInt_clusters_res_")

IntData.B_8h.FilCus <- RunUMAP(IntData.B_8h.FilCus,
                               reduction = "B_filtered_CCAInt.rna",
                               dims = 1:30, reduction.name = "B_filtered_CCAInt.umap.rna")

# Visualization of integrated data + clusters
DimPlot(object = IntData.B_8h.FilCus,
        reduction = "B_filtered_CCAInt.umap.rna",
        group.by = "B_filtered_CCAInt_clusters",
        raster = F, label = T)+
    themeG+
    labs(x = "RNA_UMAP1", y = "RNA_UMAP2",
         title = "B cell clusters")+
    theme(panel.border = element_rect(colour = "grey60", fill=NA, linewidth=0.5),
          axis.ticks = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank()) + 
  (DimPlot(object = IntData.B_8h.FilCus,
        reduction = "B_filtered_CCAInt.umap.rna",
        group.by = "Stimulation",
        raster = F, label = F)+
    themeG+
    labs(x = "RNA_UMAP1", y = "RNA_UMAP2",
         title = "Stimulation")+
    theme(panel.border = element_rect(colour = "grey60", fill=NA, linewidth=0.5),
          axis.ticks = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank())+
  scale_color_manual(values = colorsStim))+
  (DimPlot(object = IntData.B_8h.FilCus,
        reduction = "B_filtered_CCAInt.umap.rna",
        group.by = "predicted.celltype.l1",
        raster = F, label = F)+
    themeG+
    labs(x = "RNA_UMAP1", y = "RNA_UMAP2",
         title = "Cell type prediction Azimuth")+
    theme(panel.border = element_rect(colour = "grey60", fill=NA, linewidth=0.5),
          axis.ticks = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank())+
     scale_color_manual(values = colorsCellType)) + 
  plot_annotation(title = "CCA integrated")
```

```{r DensityPlots_Stimulations_8h, fig.height=7, fig.width=12}
Idents(IntData.B_8h.FilCus)<-'Stimulation'
# Initialization of list for plots
Dens.plotlist <- list()
# Generate the density plots and save them in the list
for (stim in unique(IntData.B_8h.FilCus$Stimulation)){
  Dens.plotlist[[stim]] <- DimPlot(IntData.B_8h.FilCus, reduction = "B_filtered_CCAInt.umap.rna",
                                 group.by = "Stimulation", cols = rep("lightgrey", times = 5))+
    themeG+
    theme(legend.position = "none")+
    stat_density2d(data = subset(IntData.B_8h.FilCus, idents = stim)@reductions[["B_filtered_CCAInt.umap.rna"]]@cell.embeddings,
                   mapping = aes(x = BfilteredCCAIntumaprna_1,
                                 y = BfilteredCCAIntumaprna_2,
                                 fill = ..level..),
                   geom = "polygon", alpha = 0.6)+
    scale_fill_gradient(low = "lightblue",high = "darkblue", guide = "colourbar",
                        breaks = c(0,0.05, 0.1))+
    labs(title = stim, x = "UMAP1", y = "UMAP2")+
    theme(axis.line = element_blank(),
          axis.ticks = element_blank(),
          axis.text = element_blank(),
          axis.title = element_blank())+
    ylim(-7,7)+
    xlim(-8,8)
}

# Join plots
ggarrange(plotlist = Dens.plotlist, ncol = 3, nrow = 2)
# Save plot
ggsave(filename = file.path(params$path_Results, paste0(params$date, 
                                                        "_FilteredB_8h_CCAint_UMAP_density_stimulations.pdf")),
       device = "pdf", height = 9, width = 12,
       dpi = 330)

# Clean
rm(Dens.plotlist)
```

### Comparison with or without integration

```{r ComparisonIntegration_DimPlot, fig.height=3, fig.width=8}
#### DIMPLOTs ####
DimPlot(object = IntData.B_8h.FilCus,
        reduction = "B_filtered_CCAInt.umap.rna",
        group.by = "samplename",
        raster = F, label = F)+
    themeG+
    labs(x = "RNA_UMAP1", y = "RNA_UMAP2",
         title = "CCA integration")+
    theme(panel.border = element_rect(colour = "grey60", fill=NA, linewidth=0.5),
          axis.ticks = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank()) +
  scale_color_manual(values = colorsSample) + 
  (DimPlot(object = IntData.B_8h.FilCus,
        reduction = "B_filtered.umap",
        group.by = "samplename",
        raster = F, label = F)+
    themeG+
    labs(x = "RNA_UMAP1", y = "RNA_UMAP2",
         title = "Without integration")+
    theme(panel.border = element_rect(colour = "grey60", fill=NA, linewidth=0.5),
          axis.ticks = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank()) +
  scale_color_manual(values = colorsSample))
```

```{r ComparisonIntegration_DensityPlots, fig.height=7, fig.width=6}
### DENSITY PLOTS ###
Idents(IntData.B_8h.FilCus)<-'samplename'
# Initialization of list for plots
Dens.plotlist <- list()
# Generate the density plots and save them in the list
for (pool in unique(IntData.B_8h.FilCus$samplename)){
  Dens.plotlist[[paste0("CCAInt - ", pool)]] <- DimPlot(IntData.B_8h.FilCus, reduction = "B_filtered_CCAInt.umap.rna",
                                 group.by = "samplename", cols = rep("lightgrey", times = 5))+
    themeG+
    theme(legend.position = "none")+
    stat_density2d(data = subset(IntData.B_8h.FilCus, idents = pool)@reductions[["B_filtered_CCAInt.umap.rna"]]@cell.embeddings,
                   mapping = aes(x = BfilteredCCAIntumaprna_1,
                                 y = BfilteredCCAIntumaprna_2,
                                 fill = ..level..),
                   geom = "polygon", alpha = 0.6)+
    scale_fill_gradient(low = "lightblue",high = "darkblue", guide = "colourbar",
                        breaks = c(0,0.05, 0.1))+
    labs(title = paste0("CCAInt - ", pool), x = "UMAP1", y = "UMAP2")+
    theme(axis.line = element_blank(),
          axis.ticks = element_blank(),
          axis.text = element_blank(),
          axis.title = element_blank(),
          panel.border = element_rect(colour = "grey60", fill=NA, linewidth=0.5))+
    ylim(-7,7)+
    xlim(-8,8)
  
  Dens.plotlist[[paste0("No Integration - ", pool)]] <- DimPlot(IntData.B_8h.FilCus, reduction = "B_filtered.umap",
                                 group.by = "samplename", cols = rep("lightgrey", times = 5))+
    themeG+
    theme(legend.position = "none")+
    stat_density2d(data = subset(IntData.B_8h.FilCus, idents = pool)@reductions[["B_filtered.umap"]]@cell.embeddings,
                   mapping = aes(x = Bfilteredumap_1,
                                 y = Bfilteredumap_2,
                                 fill = ..level..),
                   geom = "polygon", alpha = 0.6)+
    scale_fill_gradient(low = "lightblue",high = "darkblue", guide = "colourbar",
                        breaks = c(0,0.05, 0.1))+
    labs(title = paste0("No Integration - ", pool), x = "UMAP1", y = "UMAP2")+
    theme(axis.line = element_blank(),
          axis.ticks = element_blank(),
          axis.text = element_blank(),
          axis.title = element_blank(),
          panel.border = element_rect(colour = "grey60", fill=NA, linewidth=0.5))+
    ylim(-7,7)+
    xlim(-8,8)
}

# Join plots
ggarrange(plotlist = Dens.plotlist, ncol = 2, nrow = 3)

# Clean
rm(Dens.plotlist)
```

## 30 hours of stimulation

For the B cell subset obtained after 30 hours of stimulation, multiple confusing blubs were observed. However, most of them indeed contained B cells and were retained in the end. 

- **Cluster 24**: This cluster contained plasmablasts and was therefor removed from this B cell subset.
- **Cluster 12**: This cluster contained a long "line" of cells, not really clustering together, and also mainly contained CD4+ T cells.
- **Cluster 19**: This cluster showed high percentages of mitochondrial DNA and was therefore also excluded (see below).
- **Cluster 20** also seemed outliers, but no large reasons could be found (based on annotation and feature plots of the most interestin markers). No bias to any of the samples, conditions or sample pools was observed. Therefore this cluster was retained for downstream analysis. But it also clearly again cluster apart from the other cells when performing the PCA and UMAP analysis again. _Note: I tried with leaving out these clusters separately as well, however this did not improve the findings._

```{r QualityControl_30hClusters, fig.height=10, fig.width=10}
# Change the Idents of the data to the metadata column of interest
Idents(IntData.B_30h)<-"B_clusters"
# Check some quality features
VlnPlot(object = IntData.B_30h, 
        features = c("nFeature_RNA", "nCount_RNA", "percent_mito"),
        alpha = 0.2, sort = "decreasing",
        ncol = 1)

# Check doublets
VlnPlot(object = IntData.B_30h, 
        features = c("scDblFinder.score", "predicted.celltype.l1.score"),
        alpha = 0.2, sort = "decreasing",
        ncol = 1)

# Check specifity for pools or samples
Heatmap(t(as.matrix(table(IntData.B_30h$B_clusters, 
                          IntData.B_30h$samplename))/colSums(as.matrix(table(IntData.B_30h$samplename, 
                                                                             IntData.B_30h$B_clusters)))), 
          name = "Sample pools")
Heatmap(t(as.matrix(table(IntData.B_30h$B_clusters, 
                          IntData.B_30h$PatientID))/colSums(as.matrix(table(IntData.B_30h$PatientID, 
                                                                             IntData.B_30h$B_clusters)))), name = "Healthy donors")
Heatmap(t(as.matrix(table(IntData.B_30h$B_clusters, 
                          IntData.B_30h$PatStim))/colSums(as.matrix(table(IntData.B_30h$PatStim, 
                                                                             IntData.B_30h$B_clusters)))), name = "Samples")
Heatmap(t(as.matrix(table(IntData.B_30h$B_clusters, 
                          IntData.B_30h$Stimulation))/colSums(as.matrix(table(IntData.B_30h$Stimulation, 
                                                                             IntData.B_30h$B_clusters)))), name = "Stimulation")
```



```{r Filtering_UMAPGeneration_30h, fig.width=6, fig.height=4}
# Change idents needed in the subset function
Idents(IntData.B_30h) <- "B_clusters"
# Subsetting the data
IntData.B_30h.FilCus <- subset(IntData.B_30h,
                              # Only extract the clusters without cluster 17 and 22
                              idents = unique(levels(IntData.B_30h$B_clusters))[!unique(levels(IntData.B_30h$B_clusters)) %in% 
                                                                                  c("12", "19", "24")])
# Visualization to check if clusters were proberly removed
DimPlot(object = IntData.B_30h.FilCus,
        reduction = "B_integrated.umap.rna",
        group.by = "B_clusters",
        raster = F, label = T)+
    themeG+
    labs(x = "RNA_UMAP1", y = "RNA_UMAP2",
         title = "B cell clusters")+
    theme(panel.border = element_rect(colour = "grey60", fill=NA, linewidth=0.5),
          axis.ticks = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank())
```

### Dimensionality reduction without data integration

In a first step, dimensionality reduction was performe in the standard way, without taking into account any integration of the data. This is performed by generating PC of interest and clustering based on these PCs. Next, a UMAP was generated using the most interesting PCs (using an albow plot to decide on this number).

```{r DimRed_NoIntegration, fig.width=15, fig.height=4}
# Generation of new UMAP 
## Run PCA analysis for dimensionality reduction
IntData.B_30h.FilCus <- Seurat::RunPCA(object = IntData.B_30h.FilCus,
                                      # Based on RNA only
                                      assay = "RNA",
                                      # Name to save
                                      reduction.name = "B_filtered.pca")
## Check interesting number of dimensions in pca dimensionality reduction
ElbowPlot(IntData.B_30h.FilCus, ndims = 50, reduction = "B_filtered.pca")

## Generation of clusters
IntData.B_30h.FilCus <- FindNeighbors(IntData.B_30h.FilCus, 
                                reduction = "B_filtered.pca", 
                                dims = 1:30, graph.name = "B_filtered")
### Clustering using multiple resolution values
for (val_res in seq(0.5,2.5, by=0.5)){
  ResVal<-paste0("B_filtered_res_", val_res)
  message(paste0("Busy working on clustering with resolution ", val_res))
  IntData.B_30h.FilCus<-FindClusters(IntData.B_30h.FilCus,
                                    resolution=val_res,
                        verbose = FALSE, cluster.name = ResVal,
                        graph.name = "B_filtered")}
# Clustree to check the select the appropriate resolution
clustree(IntData.B_30h.FilCus, 
         # Select only the correct clusters based on the resolution
         prefix = "B_filtered_res_",
         # Size of the nodes
         node_size = 6,
         # Edges cross as little a possible
         layout = "sugiyama",
         # Coloring the nodes based on fraction of T cells
         # Per cluster 
         node_colour = "lightblue",
         show_axis = TRUE)+
  scale_color_gradient(low = "lightgrey",
                       high = "#A00D0D")+
  scale_edge_color_continuous(low = "lightgrey",
                             high = "#690F82")+
  labs(edge_alpha = "Proportion moved cells",
       edge_colour = "Counts moved cells",
       colour = "B cell fraction",
       title = "B filtered clusters",
       subtitle = "RNA")+
  ylab("Resolution")+
  theme_clustree
## Resolution of 1.5 seems okay 
## Better to overcluster, merging clusters is possible if necessary
Res_Method<-c("Resolution" = 1.5)
Res_Method[]<-paste0("B_filtered_res_",Res_Method)
IntData.B_30h.FilCus<-SelectMetaData(object = IntData.B_30h.FilCus,
                               to_stay = c("B_filtered_clusters" = Res_Method[["Resolution"]]),
                               pattern_remove = "B_filtered_res_")

## RUN UMAP dimensionality reduction based on the previous PCA analysis
IntData.B_30h.FilCus <- RunUMAP(IntData.B_30h.FilCus, 
                    reduction = "B_filtered.pca", 
                    # Focus on the first 30 PC based on the elbow plot generated above
                    dims = 1:30, reduction.name = "B_filtered.umap")

# Visualization of the final UMAP and its clusters
DimPlot(object = IntData.B_30h.FilCus,
        reduction = "B_filtered.umap",
        group.by = "B_filtered_clusters",
        raster = F, label = T)+
    themeG+
    labs(x = "RNA_UMAP1", y = "RNA_UMAP2",
         title = "B cell clusters")+
    theme(panel.border = element_rect(colour = "grey60", fill=NA, linewidth=0.5),
          axis.ticks = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank()) + 
  (DimPlot(object = IntData.B_30h.FilCus,
        reduction = "B_filtered.umap",
        group.by = "Stimulation",
        raster = F, label = F)+
    themeG+
    labs(x = "RNA_UMAP1", y = "RNA_UMAP2",
         title = "B cell clusters")+
    theme(panel.border = element_rect(colour = "grey60", fill=NA, linewidth=0.5),
          axis.ticks = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank())+
  scale_color_manual(values = colorsStim))+
  (DimPlot(object = IntData.B_30h.FilCus,
        reduction = "B_filtered.umap",
        group.by = "predicted.celltype.l1",
        raster = F, label = F)+
    themeG+
    labs(x = "RNA_UMAP1", y = "RNA_UMAP2",
         title = "B cell clusters")+
    theme(panel.border = element_rect(colour = "grey60", fill=NA, linewidth=0.5),
          axis.ticks = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank())+
     scale_color_manual(values = colorsCellType))

table(IntData.B_30h.FilCus$B_filtered_clusters, IntData.B_30h.FilCus$B_clusters)

# Save the plot
ggsave(filename = file.path(params$path_Results, paste0(params$date,
                                                        "_FilteredB_30h_UMAP_clusters_stimulations_annotation.pdf")),
       device = "pdf", height = 4, width = 15,
       dpi = 330)
```

```{r FeaturePlot_30hours}
# Check the feature expression typical for B and T cells
# Mainly focussing on CD3 and CD19 as markers (expressed in ADT)
DefaultAssay(IntData.B_30h)<-"ADT"
FeaturePlot(IntData.B_30h, features  = c("CD3", "CD3.1", "CD19.1"),
            reduction = "B_integrated.umap.rna")

# Mainly focussing on CD3 and CD19 as markers (expressed in RNA)
DefaultAssay(IntData.B_30h)<-"RNA"
FeaturePlot(IntData.B_30h, features  = c("CD3E", "CD3D", "CD3G", "CD19"),
            reduction = "B_integrated.umap.rna")

# Check filtered cells as well
# Mainly focussing on CD3 and CD19 as markers (expressed in RNA)
DefaultAssay(IntData.B_30h.FilCus)<-"RNA"
FeaturePlot(IntData.B_30h.FilCus, features  = c("CD3E", "CD3D", "CD3G", "CD19"),
            reduction = "B_filtered.umap")
```

```{r DensityPlots_Stimulations_30h, fig.height=10, fig.width=12}
Idents(IntData.B_30h.FilCus)<-'Stimulation'
# Initialization of list for plots
Dens.plotlist <- list()
# Generate the density plots and save them in the list
for (stim in unique(IntData.B_30h.FilCus$Stimulation)){
  Dens.plotlist[[stim]] <- DimPlot(IntData.B_30h.FilCus, reduction = "B_filtered.umap",
                                 group.by = "Stimulation", cols = rep("lightgrey", times = 5))+
    themeG+
    theme(legend.position = "none")+
    stat_density2d(data = subset(IntData.B_30h.FilCus, idents = stim)@reductions[["B_filtered.umap"]]@cell.embeddings,
                   mapping = aes(x = Bfilteredumap_1,
                                 y = Bfilteredumap_2,
                                 fill = ..level..),
                   geom = "polygon", alpha = 0.6)+
    scale_fill_gradient(low = "lightblue",high = "darkblue", guide = "colourbar",
                        breaks = c(0,0.05, 0.1))+
    labs(title = stim, x = "UMAP1", y = "UMAP2")+
    theme(axis.line = element_blank(),
          axis.ticks = element_blank(),
          axis.text = element_blank(),
          axis.title = element_blank())+
    ylim(-7,7)+
    xlim(-8,8)
}

# Join plots
ggarrange(plotlist = Dens.plotlist, ncol = 3, nrow = 2)
# Save plot
ggsave(filename = file.path(params$path_Results, paste0(params$date, 
                                                        "_FilteredB_30h_UMAP_density_stimulations.pdf")),
       device = "pdf", height = 10, width = 12,
       dpi = 330)

# Clean
rm(Dens.plotlist)
```

### Dimensionality reduction with data integration - CCA

In this part, the filtered B cell subset is first integrated using the CCA approach. Additionally, clusters are generated and a UMAP dimensionality reduction is performed for downstream visualization.

```{r DimRed_CCAintegration, fig.width=15, fig.height=4}
#### INTEGRATION ####
# Set the default assay to RNA
DefaultAssay(IntData.B_30h.FilCus) <- "RNA"

# Split the data according the sampleID (batch effects)
IntData.B_30h.FilCus[["RNA"]]<-split(IntData.B_30h.FilCus[["RNA"]],
                                    IntData.B_30h$samplename)
# Normalization
IntData.B_30h.FilCus<-Seurat::NormalizeData(IntData.B_30h.FilCus)
# Variable Features
IntData.B_30h.FilCus<-Seurat::FindVariableFeatures(IntData.B_30h.FilCus)
# Scaling
IntData.B_30h.FilCus<-Seurat::ScaleData(IntData.B_30h.FilCus, 
                                       features = rownames(IntData.B_30h.FilCus))
# Dimensionality reduction (unintegrated)
IntData.B_30h.FilCus<-Seurat::RunPCA(IntData.B_30h.FilCus, assay = "RNA", 
                                    reduction.name = "B_filtered_pca.priorIntegration.rna")
IntData.B_30h.FilCus<-Seurat::FindNeighbors(IntData.B_30h.FilCus, dims = 1:20, 
                                           reduction = "B_filtered_pca.priorIntegration.rna")
IntData.B_30h.FilCus<-Seurat::FindClusters(IntData.B_30h.FilCus, resolution = 2, 
                                          cluster.name = "B_filtered_pca_clusters.priorIntegration.rna")
IntData.B_30h.FilCus<-Seurat::RunUMAP(IntData.B_30h.FilCus, dims = 1:20, reduction = "B_filtered_pca.priorIntegration.rna", 
                           reduction.name = "B_filtered_umap.priorIntegration.rna")
# CCA integration of RNA layer
IntData.B_30h.FilCus<-IntegrateLayers(IntData.B_30h.FilCus, method = CCAIntegration,
                           orig.reduction = "B_filtered_pca.priorIntegration.rna", 
                           new.reduction = "B_filtered_CCAInt.rna",
                           verbose = F)

# Joining layers
IntData.B_30h.FilCus[["RNA"]] <- JoinLayers(IntData.B_30h.FilCus[["RNA"]])

#### CLUSTERING ####
# Clustering based on the integrated data
IntData.B_30h.FilCus <- FindNeighbors(IntData.B_30h.FilCus, 
                           reduction = "B_filtered_CCAInt.rna", 
                           dims = 1:30, graph.name = "B_filtered_CCAInt.rna_graph")
# Clustering using multiple resolution values
for (val_res in seq(0.5,2.5, by=0.5)){
  ResVal<-paste0("B_filtered_CCAInt_clusters_res_", val_res)
  message(paste0("Busy working on B cell clustering with resolution ", val_res))
  IntData.B_30h.FilCus<-FindClusters(IntData.B_30h.FilCus,resolution=val_res,
                          verbose = FALSE, cluster.name = ResVal,
                          graph.name = "B_filtered_CCAInt.rna_graph")}

# Clustree to check the select the appropriate resolution
clustree(IntData.B_30h.FilCus, 
         # Select only the correct clusters based on the resolution
         prefix = "B_filtered_CCAInt_clusters_res_",
         # Size of the nodes
         node_size = 6,
         # Edges cross as little a possible
         layout = "sugiyama",
         # Coloring the nodes based on fraction of B cells
         # Per cluster 
         node_colour = "lightblue",
         show_axis = TRUE)+
  scale_color_gradient(low = "lightgrey",
                       high = "#A00D0D")+
  scale_edge_color_continuous(low = "lightgrey",
                              high = "#690F82")+
  labs(edge_alpha = "Proportion moved cells",
       edge_colour = "Counts moved cells",
       colour = "B cell fraction",
       title = "Singlets B cells",
       subtitle = "RNA")+
  ylab("Resolution")+
  theme_clustree

# Resolution of 1.5 seems okay 
# Better to overcluster, merging clusters is possible if necessary
Res_Method<-c("Resolution" = 1.5)
Res_Method[]<-paste0("B_filtered_CCAInt_clusters_res_",Res_Method)

IntData.B_30h.FilCus<-SelectMetaData(object = IntData.B_30h.FilCus,
                          to_stay = c("B_filtered_CCAInt_clusters" = Res_Method[["Resolution"]]),
                          pattern_remove = "B_filtered_CCAInt_clusters_res_")

IntData.B_30h.FilCus <- RunUMAP(IntData.B_30h.FilCus,
                               reduction = "B_filtered_CCAInt.rna",
                               dims = 1:30, reduction.name = "B_filtered_CCAInt.umap.rna")

# Visualization of integrated data + clusters
DimPlot(object = IntData.B_30h.FilCus,
        reduction = "B_filtered_CCAInt.umap.rna",
        group.by = "B_filtered_CCAInt_clusters",
        raster = F, label = T)+
    themeG+
    labs(x = "RNA_UMAP1", y = "RNA_UMAP2",
         title = "B cell clusters")+
    theme(panel.border = element_rect(colour = "grey60", fill=NA, linewidth=0.5),
          axis.ticks = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank()) + 
  (DimPlot(object = IntData.B_30h.FilCus,
        reduction = "B_filtered_CCAInt.umap.rna",
        group.by = "Stimulation",
        raster = F, label = F)+
    themeG+
    labs(x = "RNA_UMAP1", y = "RNA_UMAP2",
         title = "Stimulation")+
    theme(panel.border = element_rect(colour = "grey60", fill=NA, linewidth=0.5),
          axis.ticks = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank())+
  scale_color_manual(values = colorsStim))+
  (DimPlot(object = IntData.B_30h.FilCus,
        reduction = "B_filtered_CCAInt.umap.rna",
        group.by = "predicted.celltype.l1",
        raster = F, label = F)+
    themeG+
    labs(x = "RNA_UMAP1", y = "RNA_UMAP2",
         title = "Cell type prediction Azimuth")+
    theme(panel.border = element_rect(colour = "grey60", fill=NA, linewidth=0.5),
          axis.ticks = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank())+
     scale_color_manual(values = colorsCellType)) + 
  plot_annotation(title = "CCA integrated")
```

```{r DensityPlots_Stimulations_30h, fig.height=10, fig.width=12}
Idents(IntData.B_30h.FilCus)<-'Stimulation'
# Initialization of list for plots
Dens.plotlist <- list()
# Generate the density plots and save them in the list
for (stim in unique(IntData.B_30h.FilCus$Stimulation)){
  Dens.plotlist[[stim]] <- DimPlot(IntData.B_30h.FilCus, reduction = "B_filtered_CCAInt.umap.rna",
                                 group.by = "Stimulation", cols = rep("lightgrey", times = 5))+
    themeG+
    theme(legend.position = "none")+
    stat_density2d(data = subset(IntData.B_30h.FilCus, idents = stim)@reductions[["B_filtered_CCAInt.umap.rna"]]@cell.embeddings,
                   mapping = aes(x = BfilteredCCAIntumaprna_1,
                                 y = BfilteredCCAIntumaprna_2,
                                 fill = ..level..),
                   geom = "polygon", alpha = 0.6)+
    scale_fill_gradient(low = "lightblue",high = "darkblue", guide = "colourbar",
                        breaks = c(0,0.05, 0.1))+
    labs(title = stim, x = "UMAP1", y = "UMAP2")+
    theme(axis.line = element_blank(),
          axis.ticks = element_blank(),
          axis.text = element_blank(),
          axis.title = element_blank())+
    ylim(-8,8)+
    xlim(-8,8)
}

# Join plots
ggarrange(plotlist = Dens.plotlist, ncol = 3, nrow = 2)
# Save plot
ggsave(filename = file.path(params$path_Results, paste0(params$date, 
                                                        "_FilteredB_30h_CCAint_UMAP_density_stimulations.pdf")),
       device = "pdf", height = 9, width = 12,
       dpi = 330)

# Clean
rm(Dens.plotlist)
```

### Comparison with or without integration

```{r ComparisonIntegration_DimPlot, fig.height=3, fig.width=8}
#### DIMPLOTs ####
DimPlot(object = IntData.B_30h.FilCus,
        reduction = "B_filtered_CCAInt.umap.rna",
        group.by = "samplename",
        raster = F, label = F)+
    themeG+
    labs(x = "RNA_UMAP1", y = "RNA_UMAP2",
         title = "CCA integration")+
    theme(panel.border = element_rect(colour = "grey60", fill=NA, linewidth=0.5),
          axis.ticks = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank()) +
  scale_color_manual(values = colorsSample) + 
  (DimPlot(object = IntData.B_30h.FilCus,
        reduction = "B_filtered.umap",
        group.by = "samplename",
        raster = F, label = F)+
    themeG+
    labs(x = "RNA_UMAP1", y = "RNA_UMAP2",
         title = "Without integration")+
    theme(panel.border = element_rect(colour = "grey60", fill=NA, linewidth=0.5),
          axis.ticks = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank()) +
  scale_color_manual(values = colorsSample))
```

```{r ComparisonIntegration_DensityPlots, fig.height=7, fig.width=6}
### DENSITY PLOTS ###
Idents(IntData.B_30h.FilCus)<-'samplename'
# Initialization of list for plots
Dens.plotlist <- list()
# Generate the density plots and save them in the list
for (pool in unique(IntData.B_30h.FilCus$samplename)){
  Dens.plotlist[[paste0("CCAInt - ", pool)]] <- DimPlot(IntData.B_30h.FilCus, reduction = "B_filtered_CCAInt.umap.rna",
                                 group.by = "samplename", cols = rep("lightgrey", times = 5))+
    themeG+
    theme(legend.position = "none")+
    stat_density2d(data = subset(IntData.B_30h.FilCus, idents = pool)@reductions[["B_filtered_CCAInt.umap.rna"]]@cell.embeddings,
                   mapping = aes(x = BfilteredCCAIntumaprna_1,
                                 y = BfilteredCCAIntumaprna_2,
                                 fill = ..level..),
                   geom = "polygon", alpha = 0.6)+
    scale_fill_gradient(low = "lightblue",high = "darkblue", guide = "colourbar",
                        breaks = c(0,0.05, 0.1))+
    labs(title = paste0("CCAInt - ", pool), x = "UMAP1", y = "UMAP2")+
    theme(axis.line = element_blank(),
          axis.ticks = element_blank(),
          axis.text = element_blank(),
          axis.title = element_blank(),
          panel.border = element_rect(colour = "grey60", fill=NA, linewidth=0.5))+
    ylim(-7,7)+
    xlim(-8,8)
  
  Dens.plotlist[[paste0("No Integration - ", pool)]] <- DimPlot(IntData.B_30h.FilCus, reduction = "B_filtered.umap",
                                 group.by = "samplename", cols = rep("lightgrey", times = 5))+
    themeG+
    theme(legend.position = "none")+
    stat_density2d(data = subset(IntData.B_30h.FilCus, idents = pool)@reductions[["B_filtered.umap"]]@cell.embeddings,
                   mapping = aes(x = Bfilteredumap_1,
                                 y = Bfilteredumap_2,
                                 fill = ..level..),
                   geom = "polygon", alpha = 0.6)+
    scale_fill_gradient(low = "lightblue",high = "darkblue", guide = "colourbar",
                        breaks = c(0,0.05, 0.1))+
    labs(title = paste0("No Integration - ", pool), x = "UMAP1", y = "UMAP2")+
    theme(axis.line = element_blank(),
          axis.ticks = element_blank(),
          axis.text = element_blank(),
          axis.title = element_blank(),
          panel.border = element_rect(colour = "grey60", fill=NA, linewidth=0.5))+
    ylim(-7,7)+
    xlim(-8,8)
}

# Join plots
ggarrange(plotlist = Dens.plotlist, ncol = 2, nrow = 3)

# Clean
rm(Dens.plotlist)
```


## Save final B cell subsets
```{r SaveRDS_BSubsets}
saveRDS(IntData.B_8h.FilCus,
        file = file.path(params$path_Results,
                         "SeuratObjects",
                         paste0(params$date, "_seuratObj_NoBase_Integrated_Bclusters_Filtered_8h.rds")))
saveRDS(IntData.B_30h.FilCus,
        file = file.path(params$path_Results,
                         "SeuratObjects",
                         paste0(params$date, "_seuratObj_NoBase_Integrated_Bclusters_Filtered_30h.rds")))
```

## Read in final B cell subsets
```{r SaveRDS_BSubsets, eval=FALSE}
IntData.B_8h.FilCus<-readRDS(file = file.path(params$path_Results,
                                             "SeuratObjects",
                                             paste0(params$date, "_seuratObj_NoBase_Integrated_Bclusters_Filtered_8h.rds")))
IntData.B_30h.FilCus<-readRDS(file = file.path(params$path_Results,
                                               "SeuratObjects",
                                               paste0(params$date, "_seuratObj_NoBase_Integrated_Bclusters_Filtered_30h.rds")))
```

# Differential gene expression between stimulated and unstimulated cells 

This section is almost entirely copy-pasted as coded in the script `011_DEGenes_Visualization_GSEAOverlap.Rmd`. The only difference is the input B cell subset, using the filtered subset this time, and an adaptation was perform to perform the HGT analysis more correctly.

This section focuses on the differential gene expression analysis solely based on the metadata annotation of the cells: either unstimulated or stimulated with a specific stimulatory condition.

```{r DEGeneAnalysis_Stimulation}
set.seed(2024)
# Change idents of cells
Idents(IntData.B_8h.FilCus)<-"Stimulation"
Idents(IntData.B_30h.FilCus)<-"Stimulation"

#Initialization of data frame
DEG_Stim_Unstim_8h_FilClus<-sapply(unique(IntData.B_8h.FilCus$Stimulation),
                        function(x) NULL)
DEG_Stim_Unstim_30h_FilClus<-sapply(unique(IntData.B_30h.FilCus$Stimulation),
                        function(x) NULL)

# Remove unstimulated (is used as a reference)
DEG_Stim_Unstim_8h_FilClus[["Unstimulated"]]<-NULL
DEG_Stim_Unstim_30h_FilClus[["Unstimulated"]]<-NULL

# Perform DEG analysis
for (stim in names(DEG_Stim_Unstim_8h_FilClus)){
  DEG_Stim_Unstim_8h_FilClus[[stim]]<-FindMarkers(object = IntData.B_8h.FilCus,
                                       min.pct = 0, logfc.threshold = 0,
                                       ident.1 = stim, ident.2 = "Unstimulated")%>%
    mutate(Score = (pct.1/(pct.2 + 0.01))*avg_log2FC)%>%
    rownames_to_column(var = "Gene")
}

for (stim in names(DEG_Stim_Unstim_30h_FilClus)){
  DEG_Stim_Unstim_30h_FilClus[[stim]]<-FindMarkers(object = IntData.B_30h.FilCus,
                                       min.pct = 0, logfc.threshold = 0,
                                       ident.1 = stim, ident.2 = "Unstimulated")%>%
    mutate(Score = (pct.1/(pct.2 + 0.01))*avg_log2FC)%>%
    rownames_to_column(var = "Gene")
}

# Save list of DE genes (all genes are scored here)
saveRDS(DEG_Stim_Unstim_8h_FilClus, 
        file = file.path(params$path_Results, 
                         "DEGAnalysis",
                         paste(params$date,
                               params$prefix_output, "DE_AllGenes_Stim_vs_Unstim_8h.rds", sep = "_")))
saveRDS(DEG_Stim_Unstim_30h_FilClus, 
        file = file.path(params$path_Results, 
                         "DEGAnalysis",
                         paste(params$date,
                               params$prefix_output, "DE_AllGenes_Stim_vs_Unstim_30h.rds", sep = "_")))
``` 

To check if these genes are up or downregulated, an additional annotation is added to the dataframe.

```{r DEGenesAnnot}
for (stim in names(DEG_Stim_Unstim_8h_FilClus)){
  # Add additional information to the dataset
  DEG_Stim_Unstim_8h_FilClus[[stim]]<-DEG_Stim_Unstim_8h_FilClus[[stim]]%>%
  mutate(Annot=case_when(
    avg_log2FC > 1 & p_val_adj <0.05 ~ "UP",
    avg_log2FC < -1 & p_val_adj <0.05 ~ "DOWN",
    TRUE ~ "NS"))
  # Select genes to be plotted as label on the volcano graph
  top20genes <- head(DEG_Stim_Unstim_8h_FilClus[[stim]]%>%
                       filter(Annot != 'NS')%>%
                       arrange(p_val_adj, descending = TRUE), 20)[,'Gene']
  DEG_Stim_Unstim_8h_FilClus[[stim]]<-DEG_Stim_Unstim_8h_FilClus[[stim]]%>%
    mutate(GeneLabels=ifelse(DEG_Stim_Unstim_8h_FilClus[[stim]]$Gene %in% top20genes,
                             DEG_Stim_Unstim_8h_FilClus[[stim]]$Gene,NA))
}

for (stim in names(DEG_Stim_Unstim_30h_FilClus)){
  # Add additional information to the dataset
  DEG_Stim_Unstim_30h_FilClus[[stim]]<-DEG_Stim_Unstim_30h_FilClus[[stim]]%>%
  mutate(Annot=case_when(
    avg_log2FC > 1 & p_val_adj <0.05 ~ "UP",
    avg_log2FC < -1 & p_val_adj <0.05 ~ "DOWN",
    TRUE ~ "NS"))
  # Select genes to be plotted as label on the volcano graph
  top20genes <- head(DEG_Stim_Unstim_30h_FilClus[[stim]]%>%
                       filter(Annot != 'NS')%>%
                       arrange(p_val_adj, descending = TRUE), 20)[,'Gene']
  DEG_Stim_Unstim_30h_FilClus[[stim]]<-DEG_Stim_Unstim_30h_FilClus[[stim]]%>%
    mutate(GeneLabels=ifelse(DEG_Stim_Unstim_30h_FilClus[[stim]]$Gene %in% top20genes,
                             DEG_Stim_Unstim_30h_FilClus[[stim]]$Gene,NA))
}
rm(top20genes)
```

## Visualization with Volcano plot

```{r Volcano_8h, fig.height=4, fig.width=6}
ggplot(DEG_Stim_Unstim_8h_FilClus$IFNgamma, mapping = aes(x = avg_log2FC, y = ifelse(p_val_adj == 0,
                                                                              -log10(.Machine$double.xmin),
                                                                              -log10(p_val_adj)),
                                        color = Annot, label = GeneLabels))+
  geom_point(size = 0.7)+
  geom_vline(xintercept = c(1, -1), color = "grey60", linetype = "dashed")+
  geom_hline(yintercept = -log10(0.05), color = "grey60", linetype = "dashed")+
  themeG+
  theme(panel.grid.major = element_line(color = "grey90", 
                                  size = 0.4, 
                                  linetype = 1),
        axis.line = element_blank(),
        axis.ticks = element_blank())+
  coord_cartesian(expand = TRUE)+
  scale_color_manual(values = colDirection)+
  ggtitle("IFNg/R-848")+
  ylim(0,400)+xlim(-10,10)+
  geom_text_repel(max.overlaps = Inf)+
  labs(x = "Average Log2FC", y = "-Log10 of adjusted p-value")

ggplot(DEG_Stim_Unstim_8h_FilClus$MegaCD40L, mapping = aes(x = avg_log2FC, y = ifelse(p_val_adj == 0,
                                                                              -log10(.Machine$double.xmin),
                                                                              -log10(p_val_adj)),
                                        color = Annot, label = GeneLabels))+
  geom_point(size = 0.7)+
  geom_vline(xintercept = c(1, -1), color = "grey60", linetype = "dashed")+
  geom_hline(yintercept = -log10(0.05), color = "grey60", linetype = "dashed")+
  themeG+
  theme(panel.grid.major = element_line(color = "grey90", 
                                  size = 0.4, 
                                  linetype = 1),
        axis.line = element_blank(),
        axis.ticks = element_blank())+
  coord_cartesian(expand = TRUE)+
  scale_color_manual(values = colDirection)+
  ggtitle("MegaCD40L")+
  ylim(0,310)+xlim(-10,10)+
  geom_text_repel(max.overlaps = Inf)+
  labs(x = "Average Log2FC", y = "-Log10 of adjusted p-value")

ggplot(DEG_Stim_Unstim_8h_FilClus$`CD3/CD28_beads`, mapping = aes(x = avg_log2FC, y = ifelse(p_val_adj == 0,
                                                                              -log10(.Machine$double.xmin),
                                                                              -log10(p_val_adj)),
                                        color = Annot, label = GeneLabels))+
  geom_point(size = 0.7)+
  geom_vline(xintercept = c(1, -1), color = "grey60", linetype = "dashed")+
  geom_hline(yintercept = -log10(0.05), color = "grey60", linetype = "dashed")+
  themeG+
  theme(panel.grid.major = element_line(color = "grey90", 
                                  size = 0.4, 
                                  linetype = 1),
        axis.line = element_blank(),
        axis.ticks = element_blank())+
  coord_cartesian(expand = TRUE)+
  scale_color_manual(values = colDirection)+
  ggtitle("Anti-CD3/CD28")+
  ylim(0,310)+xlim(-10,10)+
  geom_text_repel(max.overlaps = Inf)+
  labs(x = "Average Log2FC", y = "-Log10 of adjusted p-value")

ggplot(DEG_Stim_Unstim_8h_FilClus$ODN2006, mapping = aes(x = avg_log2FC, y = ifelse(p_val_adj == 0,
                                                                              -log10(.Machine$double.xmin),
                                                                              -log10(p_val_adj)),
                                        color = Annot, label = GeneLabels))+
  geom_point(size = 0.7)+
  geom_vline(xintercept = c(1, -1), color = "grey60", linetype = "dashed")+
  geom_hline(yintercept = -log10(0.05), color = "grey60", linetype = "dashed")+
  themeG+
  theme(panel.grid.major = element_line(color = "grey90", 
                                  size = 0.4, 
                                  linetype = 1),
        axis.line = element_blank(),
        axis.ticks = element_blank())+
  coord_cartesian(expand = TRUE)+
  scale_color_manual(values = colDirection)+
  ggtitle("ODN2006")+
  ylim(0,310)+xlim(-10,10)+
  geom_text_repel(max.overlaps = Inf)+
  labs(x = "Average Log2FC", y = "-Log10 of adjusted p-value")
```

```{r Volcano_30h, fig.height=4, fig.width=6}
ggplot(DEG_Stim_Unstim_30h_FilClus$IFNgamma, mapping = aes(x = avg_log2FC, y = ifelse(p_val_adj == 0,
                                                                              -log10(.Machine$double.xmin),
                                                                              -log10(p_val_adj)),
                                        color = Annot, label = GeneLabels))+
  geom_point(size = 0.7)+
  geom_vline(xintercept = c(1, -1), color = "grey60", linetype = "dashed")+
  geom_hline(yintercept = -log10(0.05), color = "grey60", linetype = "dashed")+
  themeG+
  theme(panel.grid.major = element_line(color = "grey90", 
                                  size = 0.4, 
                                  linetype = 1),
        axis.line = element_blank(),
        axis.ticks = element_blank())+
  coord_cartesian(expand = TRUE)+
  scale_color_manual(values = colDirection)+
  ggtitle("IFNg/R-848")+
  ylim(0,400)+xlim(-10,10)+
  geom_text_repel(max.overlaps = Inf)+
  labs(x = "Average Log2FC", y = "-Log10 of adjusted p-value")

ggplot(DEG_Stim_Unstim_30h_FilClus$MegaCD40L, mapping = aes(x = avg_log2FC, y = ifelse(p_val_adj == 0,
                                                                              -log10(.Machine$double.xmin),
                                                                              -log10(p_val_adj)),
                                        color = Annot, label = GeneLabels))+
  geom_point(size = 0.7)+
  geom_vline(xintercept = c(1, -1), color = "grey60", linetype = "dashed")+
  geom_hline(yintercept = -log10(0.05), color = "grey60", linetype = "dashed")+
  themeG+
  theme(panel.grid.major = element_line(color = "grey90", 
                                  size = 0.4, 
                                  linetype = 1),
        axis.line = element_blank(),
        axis.ticks = element_blank())+
  coord_cartesian(expand = TRUE)+
  scale_color_manual(values = colDirection)+
  ggtitle("MegaCD40L")+
  ylim(0,310)+xlim(-10,10)+
  geom_text_repel(max.overlaps = Inf)+
  labs(x = "Average Log2FC", y = "-Log10 of adjusted p-value")

ggplot(DEG_Stim_Unstim_30h_FilClus$`CD3/CD28_beads`, mapping = aes(x = avg_log2FC, y = ifelse(p_val_adj == 0,
                                                                              -log10(.Machine$double.xmin),
                                                                              -log10(p_val_adj)),
                                        color = Annot, label = GeneLabels))+
  geom_point(size = 0.7)+
  geom_vline(xintercept = c(1, -1), color = "grey60", linetype = "dashed")+
  geom_hline(yintercept = -log10(0.05), color = "grey60", linetype = "dashed")+
  themeG+
  theme(panel.grid.major = element_line(color = "grey90", 
                                  size = 0.4, 
                                  linetype = 1),
        axis.line = element_blank(),
        axis.ticks = element_blank())+
  coord_cartesian(expand = TRUE)+
  scale_color_manual(values = colDirection)+
  ggtitle("Anti-CD3/CD28")+
  ylim(0,310)+xlim(-10,10)+
  geom_text_repel(max.overlaps = Inf)+
  labs(x = "Average Log2FC", y = "-Log10 of adjusted p-value")

ggplot(DEG_Stim_Unstim_30h_FilClus$ODN2006, mapping = aes(x = avg_log2FC, y = ifelse(p_val_adj == 0,
                                                                              -log10(.Machine$double.xmin),
                                                                              -log10(p_val_adj)),
                                        color = Annot, label = GeneLabels))+
  geom_point(size = 0.7)+
  geom_vline(xintercept = c(1, -1), color = "grey60", linetype = "dashed")+
  geom_hline(yintercept = -log10(0.05), color = "grey60", linetype = "dashed")+
  themeG+
  theme(panel.grid.major = element_line(color = "grey90", 
                                  size = 0.4, 
                                  linetype = 1),
        axis.line = element_blank(),
        axis.ticks = element_blank())+
  coord_cartesian(expand = TRUE)+
  scale_color_manual(values = colDirection)+
  ggtitle("ODN2006")+
  ylim(0,310)+xlim(-10,10)+
  geom_text_repel(max.overlaps = Inf)+
  labs(x = "Average Log2FC", y = "-Log10 of adjusted p-value")
```


## Total gene per stimulation

```{r DEGenesTotal}
TotalDEGene_Time<-rbind(
  # Extract and combine both time points
  lapply(X = DEG_Stim_Unstim_8h_FilClus, 
       FUN = function(x){table(x$Annot)})%>%
    unlist()%>%
    data.frame()%>%
    mutate(Time = "8h")%>%
  rownames_to_column(var = "Stimulation"),
  lapply(X = DEG_Stim_Unstim_30h_FilClus,
         FUN = function(x){table(x$Annot)})%>%
    unlist()%>%
    data.frame()%>%
    mutate(Time = "30h")%>%
  rownames_to_column(var = "Stimulation"))

# Rename the columns
names(TotalDEGene_Time)[names(TotalDEGene_Time)=="."]<-"Counts"

# Reorder the data frame
TotalDEGene_Time<-TotalDEGene_Time%>%
  mutate(Direction = gsub(x = Stimulation, pattern = ".*\\.", replacement = ""),
         Stimulation = gsub(x = Stimulation, pattern = "\\..*", replacement = ""),
         Time = factor(x = Time,
                       levels = c("8h", "30h"),
                       ordered = T),
         Stimulation = factor(x = Stimulation,
                              levels = c("ODN2006", "IFNgamma", "MegaCD40L", "CD3/CD28_beads"),
                              ordered = T),
         Direction = factor(x = Direction,
                            levels = c("UP", "DOWN"),
                            ordered = T))%>%
  filter(Direction != "NS")
TotalDEGene_Time
```

```{r DEGenesTotal_Barplot, fig.height=3, fig.width=6}
# Visualization using bar plots
ggplot(TotalDEGene_Time%>%
         mutate(Stimulation = factor(Stimulation,
                                     levels = c("IFNgamma", "ODN2006", "CD3/CD28_beads", "MegaCD40L"),
                                     ordered = T)), 
       mapping = aes(y = Counts, x = Time, fill = Direction))+
  geom_col(position = 'dodge')+
  scale_fill_manual(values = colDirection)+
  facet_wrap(facets = vars(Stimulation),
             ncol = 4)+
  themeG+
  theme(axis.line = element_blank(),
        axis.ticks = element_blank(),
        panel.grid.major.y = element_line(color = "grey95"))+
  labs(y = "# Differentially expressed genes",
       x = NULL)
```

```{r DEGenesTotal_Dotplot}
# Visualization
ggplot(TotalDEGene_Time, mapping = aes(x = Time, y = Counts, color = Direction,
                                       group = Direction))+
  geom_point()+
  geom_line(alpha = 0.25)+
  scale_color_manual(values = colDirection)+
  facet_wrap(facets = vars(Stimulation),
             ncol = 4)+
  themeG+
  theme(axis.ticks = element_blank(),
        axis.line = element_blank(),
        panel.grid.minor = element_line(colour = "grey95"),
        panel.grid.major = element_line(colour = "grey95"))+
  ylim(0,NA)
```


## Overlap DEgenes - Upset plot
An upset plot is also generated to visualize the overlap of DE genes between stimulations better. This is performed using the Complex Heatmap package. Additional information can be found [here](https://jokergoo.github.io/ComplexHeatmap-reference/book/upset-plot.html). In this upset plot, `mode = "distinct'` is used to visualize the overlap between stimulatory conditions as neat as possible.

### 8h stimulation

```{r UpsetPlotVisualzation_OverlapDEGenes_8h, fig.height=4, fig.width=6}
# Only focus on upregulated genes
UpsetData_UP_8h.FilClus <- list("IFN" = (DEG_Stim_Unstim_8h_FilClus$IFNgamma%>%
                                           filter(Annot =="UP"))[, "Gene"], 
                               "MegaCD40L" = (DEG_Stim_Unstim_8h_FilClus$MegaCD40L%>%
                                                filter(Annot =="UP"))[, "Gene"], 
                               "ODN2006" = (DEG_Stim_Unstim_8h_FilClus$ODN2006%>%
                                              filter(Annot =="UP"))[, "Gene"],
                               "CD3CD28" = (DEG_Stim_Unstim_8h_FilClus$`CD3/CD28_beads`%>%
                                              filter(Annot =="UP"))[, "Gene"]) 
# Generation of list of upregulated genes per 
UpsetData_UP_8h.FilClus <- list_to_matrix(UpsetData_UP_8h.FilClus)
# Generation of comb_mat object
UpsetData_UP_8h.FilClus <- make_comb_mat(UpsetData_UP_8h.FilClus, mode = "distinct")
# Generation of the plot
UpSet(UpsetData_UP_8h.FilClus, comb_order = order(comb_size(UpsetData_UP_8h.FilClus), decreasing = T),
      comb_col = "#595B5C", bg_pt_col = "white",
      column_title = "Overlap upregulated DE genes",
      right_annotation = rowAnnotation(
                "DE genes" = anno_barplot(set_size(UpsetData_UP_8h.FilClus),
                                          border = FALSE,
                                          gp = gpar(fill = "#595B5C",
                                                    lty = "blank"), 
                                          add_numbers = T,
                                          width = unit(4, "cm"))),
      top_annotation = HeatmapAnnotation(
                "Intersection size" = anno_barplot(comb_size(UpsetData_UP_8h.FilClus),
                                                   border = FALSE,
                                                   gp = gpar(fill = "#595B5C",
                                                             lty = "blank"),
                                                   add_numbers = T,
                                                   height = unit(4, "cm")),
                annotation_name_side = "left",
                annotation_name_rot = 90))

# Only focus on downregulated genes
UpsetData_DOWN_8h.FilClus <- list("IFN" = (DEG_Stim_Unstim_8h_FilClus$IFNgamma%>%
                                             filter(Annot =="DOWN"))[, "Gene"], 
                                 "MegaCD40L" = (DEG_Stim_Unstim_8h_FilClus$MegaCD40L%>%
                                                  filter(Annot =="DOWN"))[, "Gene"], 
                                 "ODN2006" = (DEG_Stim_Unstim_8h_FilClus$ODN2006%>%
                                                filter(Annot =="DOWN"))[, "Gene"],
                                 "CD3CD28" = (DEG_Stim_Unstim_8h_FilClus$`CD3/CD28_beads`%>%
                                                filter(Annot =="DOWN"))[, "Gene"]) 
# Generation of list of downregulated genes per 
UpsetData_DOWN_8h.FilClus <- list_to_matrix(UpsetData_DOWN_8h.FilClus)
# Generation of comb_mat object
UpsetData_DOWN_8h.FilClus <- make_comb_mat(UpsetData_DOWN_8h.FilClus, mode = "distinct")
# Generation of the plot
UpSet(UpsetData_DOWN_8h.FilClus, comb_order = order(comb_size(UpsetData_DOWN_8h.FilClus), decreasing = T),
      comb_col = "#595B5C", bg_pt_col = "white",
      column_title = "Overlap downregulated DE genes",
      right_annotation = rowAnnotation(
                "DE genes" = anno_barplot(set_size(UpsetData_DOWN_8h.FilClus),
                                          border = FALSE,
                                          gp = gpar(fill = "#595B5C",
                                                    lty = "blank"), 
                                          add_numbers = T,
                                          width = unit(4, "cm"))),
      top_annotation = HeatmapAnnotation(
                "Intersection size" = anno_barplot(comb_size(UpsetData_DOWN_8h.FilClus),
                                                   border = FALSE,
                                                   gp = gpar(fill = "#595B5C",
                                                             lty = "blank"), 
                                                   add_numbers = T,
                                                   height = unit(4, "cm")),
                annotation_name_side = "left",
                annotation_name_rot = 90))

# Only focus on significantly different genes
UpsetData_SIGN_8h_FilClus <- list("IFN" = (DEG_Stim_Unstim_8h_FilClus$IFNgamma%>%
                                           filter(Annot !="NS"))[, "Gene"], 
                               "MegaCD40L" = (DEG_Stim_Unstim_8h_FilClus$MegaCD40L%>%
                                                filter(Annot !="NS"))[, "Gene"], 
                               "ODN2006" = (DEG_Stim_Unstim_8h_FilClus$ODN2006%>%
                                              filter(Annot !="NS"))[, "Gene"],
                               "CD3CD28" = (DEG_Stim_Unstim_8h_FilClus$`CD3/CD28_beads`%>%
                                              filter(Annot !="NS"))[, "Gene"]) 
# Generation of list of downregulated genes per 
UpsetData_SIGN_8h_FilClus <- list_to_matrix(UpsetData_SIGN_8h_FilClus)
# Generation of comb_mat object
UpsetData_SIGN_8h_FilClus <- make_comb_mat(UpsetData_SIGN_8h_FilClus, mode = "distinct")
# Generation of the plot
UpSet(UpsetData_SIGN_8h_FilClus, comb_order = order(comb_size(UpsetData_SIGN_8h_FilClus), decreasing = T),
      comb_col = "#595B5C", bg_pt_col = "white",
      column_title = "Overlap significant DE genes",
      right_annotation = rowAnnotation(
                "DE genes" = anno_barplot(set_size(UpsetData_SIGN_8h_FilClus),
                                          border = FALSE,
                                          gp = gpar(fill = "#595B5C",
                                                    lty = "blank"), 
                                          add_numbers = T,
                                          width = unit(4, "cm"))),
      top_annotation = HeatmapAnnotation(
                "Intersection size" = anno_barplot(comb_size(UpsetData_SIGN_8h_FilClus),
                                                   border = FALSE,
                                                   gp = gpar(fill = "#595B5C",
                                                             lty = "blank"), 
                                                   add_numbers = T,
                                                   height = unit(4, "cm")),
                annotation_name_side = "left",
                annotation_name_rot = 90))
```


```{r UpsetPlotVisualzation_OverlapDEGenes_Advanced, fig.height=4, fig.width=6}
### GENERATE DATA FRAMES NEEDED TO PLOT DATA ###
# Filter out the genes of interest in 2 different sets
UpsetData_COMB_8h_FilClus <- c(lapply(DEG_Stim_Unstim_8h_FilClus, FUN = function(stim){(stim%>%
                                                                          filter(Annot == "UP"))[, "Gene"]}),
                       lapply(DEG_Stim_Unstim_8h_FilClus, FUN = function(stim){(stim%>%
                                                                          filter(Annot == "DOWN"))[, "Gene"]}))
names(UpsetData_COMB_8h_FilClus)<- paste(names(UpsetData_COMB_8h_FilClus), rep(c("UP", "DOWN"), each = 4), sep = "_")


# Generation of list of downregulated genes per 
UpsetData_COMB_8h_FilClus <- as.data.frame(list_to_matrix(UpsetData_COMB_8h_FilClus))

# Check if the expression change is consistent between stimulations
UpsetData_COMB_8h_FilClus<-UpsetData_COMB_8h_FilClus%>%
  mutate_at(vars(ends_with("UP")), function(x){ifelse(x==1, "UP", NA)})%>%
  mutate_at(vars(ends_with("DOWN")), function(x){ifelse(x==1, "DOWN", NA)})

# Check if the genes were upregulated/downregulated/mixed over the distinct subsets 
UpsetData_COMB_8h_FilClus$Direction<-apply(UpsetData_COMB_8h_FilClus, 1, function(col) {
  u <- unique(na.omit(col))
  if (length(u) == 1) {
    # either "UP" or "DOWN"
    return(u)
  } else if (length(u) > 1) {
    return("MIXED")
  } else {
    # no non-NA values
    return(NA_character_)
  }
})

# Store the genes linked to specific subsets
# stack function creates a data frame from a list
Subsets<-stack(sapply(X = comb_name(UpsetData_SIGN_8h_FilClus),
                      FUN = function(subset){extract_comb(UpsetData_SIGN_8h_FilClus, subset)}))
colnames(Subsets) <- c("Genes", "Subset")
UpsetData_COMB_8h_FilClus <- left_join(x = UpsetData_COMB_8h_FilClus%>%
                              rownames_to_column(var = "Genes"),
                            y = Subsets,
                            by = "Genes")%>%
  mutate(Direction = factor(Direction, levels = c("MIXED", "UP", "DOWN"),
                               ordered = T))

# Store the table of the distinct directions
UpsetData_COMB_8h_FilClus.table<-table(UpsetData_COMB_8h_FilClus$Subset,
                            UpsetData_COMB_8h_FilClus$Direction)
```

Generation figure of manuscript figure. 

```{r PlotUpset_DEGenes, fig.height=3, fig.width=7}
### PLOTTING THE DATA USING UPSET PLOT OF COMPLEX HEATMAP ###
# Generate upset plot
upPlot<-
  UpSet(UpsetData_SIGN_8h_FilClus, comb_order = order(comb_size(UpsetData_SIGN_8h_FilClus), decreasing = T),
      #  Colors of the dots and balkjes onderaan 
      comb_col = "#595B5C", bg_pt_col = "white",
      # Add plot title
      column_title = "Overlap significant DE genes",
      # Update column annotation (entire set of DEgenes)
      right_annotation = rowAnnotation(
                "DE genes" = anno_barplot(set_size(UpsetData_SIGN_8h_FilClus),
                                          border = FALSE,
                                          gp = gpar(fill = "#595B5C",
                                                    lty = "blank"), 
                                          add_numbers = T,
                                          width = unit(4, "cm"))),
      # Update row annotation on top of the Upset plot
      top_annotation = HeatmapAnnotation(
                                      # Add barplot based on the direction of the DE genes
                "Intersection size" = anno_barplot(UpsetData_COMB_8h_FilClus.table,
                                                   border = FALSE,
                                                   # Store correct colors
                                                   gp = gpar(fill = colDirection[colnames(
                                                     UpsetData_COMB_8h_FilClus.table)],
                                                             lty = "blank"),
                                                   add_numbers = T,
                                                   height = unit(3, "cm")),
                # Annotation name shown at the left side
                annotation_name_side = "left",
                # Rotation of the annotation at the left side
                annotation_name_rot = 90))

# Generate legend
lgd = Legend(
  labels    = colnames(UpsetData_COMB_8h_FilClus.table),
  legend_gp = gpar(fill = colDirection[colnames(UpsetData_COMB_8h_FilClus.table)]),
  title     = "Direction"
)

# Visualise the upset plot
pdf(file.path(params$path_Results, 
              "DEGAnalysis", 
              paste0(params$date, "Upset_SignDEGenes_Stim_vs_Unstim_8h.pdf")), width=7, height=4)
draw(upPlot,
     annotation_legend_list = list(lgd),
     annotation_legend_side = "left")
dev.off()
```

### 30h stimulation

```{r UpsetPlotVisualzation_OverlapDEGenes_30h, fig.height=4, fig.width=6}
# Only focus on upregulated genes
UpsetData_UP_30h.FilClus <- list("IFN" = (DEG_Stim_Unstim_30h_FilClus$IFNgamma%>%
                                           filter(Annot =="UP"))[, "Gene"], 
                               "MegaCD40L" = (DEG_Stim_Unstim_30h_FilClus$MegaCD40L%>%
                                                filter(Annot =="UP"))[, "Gene"], 
                               "ODN2006" = (DEG_Stim_Unstim_30h_FilClus$ODN2006%>%
                                              filter(Annot =="UP"))[, "Gene"],
                               "CD3CD28" = (DEG_Stim_Unstim_30h_FilClus$`CD3/CD28_beads`%>%
                                              filter(Annot =="UP"))[, "Gene"]) 
# Generation of list of upregulated genes per 
UpsetData_UP_30h.FilClus <- list_to_matrix(UpsetData_UP_30h.FilClus)
# Generation of comb_mat object
UpsetData_UP_30h.FilClus <- make_comb_mat(UpsetData_UP_30h.FilClus, mode = "distinct")
# Generation of the plot
UpSet(UpsetData_UP_30h.FilClus, comb_order = order(comb_size(UpsetData_UP_30h.FilClus), decreasing = T),
      comb_col = "#595B5C", bg_pt_col = "white",
      column_title = "Overlap upregulated DE genes",
      right_annotation = rowAnnotation(
                "DE genes" = anno_barplot(set_size(UpsetData_UP_30h.FilClus),
                                          border = FALSE,
                                          gp = gpar(fill = "#595B5C",
                                                    lty = "blank"), 
                                          add_numbers = T,
                                          width = unit(4, "cm"))),
      top_annotation = HeatmapAnnotation(
                "Intersection size" = anno_barplot(comb_size(UpsetData_UP_30h.FilClus),
                                                   border = FALSE,
                                                   gp = gpar(fill = "#595B5C",
                                                             lty = "blank"),
                                                   add_numbers = T,
                                                   height = unit(4, "cm")),
                annotation_name_side = "left",
                annotation_name_rot = 90))

# Only focus on downregulated genes
UpsetData_DOWN_30h.FilClus <- list("IFN" = (DEG_Stim_Unstim_30h_FilClus$IFNgamma%>%
                                             filter(Annot =="DOWN"))[, "Gene"], 
                                 "MegaCD40L" = (DEG_Stim_Unstim_30h_FilClus$MegaCD40L%>%
                                                  filter(Annot =="DOWN"))[, "Gene"], 
                                 "ODN2006" = (DEG_Stim_Unstim_30h_FilClus$ODN2006%>%
                                                filter(Annot =="DOWN"))[, "Gene"],
                                 "CD3CD28" = (DEG_Stim_Unstim_30h_FilClus$`CD3/CD28_beads`%>%
                                                filter(Annot =="DOWN"))[, "Gene"]) 
# Generation of list of downregulated genes per 
UpsetData_DOWN_30h.FilClus <- list_to_matrix(UpsetData_DOWN_30h.FilClus)
# Generation of comb_mat object
UpsetData_DOWN_30h.FilClus <- make_comb_mat(UpsetData_DOWN_30h.FilClus, mode = "distinct")
# Generation of the plot
UpSet(UpsetData_DOWN_30h.FilClus, comb_order = order(comb_size(UpsetData_DOWN_30h.FilClus), decreasing = T),
      comb_col = "#595B5C", bg_pt_col = "white",
      column_title = "Overlap downregulated DE genes",
      right_annotation = rowAnnotation(
                "DE genes" = anno_barplot(set_size(UpsetData_DOWN_30h.FilClus),
                                          border = FALSE,
                                          gp = gpar(fill = "#595B5C",
                                                    lty = "blank"), 
                                          add_numbers = T,
                                          width = unit(4, "cm"))),
      top_annotation = HeatmapAnnotation(
                "Intersection size" = anno_barplot(comb_size(UpsetData_DOWN_30h.FilClus),
                                                   border = FALSE,
                                                   gp = gpar(fill = "#595B5C",
                                                             lty = "blank"), 
                                                   add_numbers = T,
                                                   height = unit(4, "cm")),
                annotation_name_side = "left",
                annotation_name_rot = 90))

# Only focus on significantly different genes
UpsetData_SIGN_30h_FilClus <- list("IFN" = (DEG_Stim_Unstim_30h_FilClus$IFNgamma%>%
                                           filter(Annot !="NS"))[, "Gene"], 
                               "MegaCD40L" = (DEG_Stim_Unstim_30h_FilClus$MegaCD40L%>%
                                                filter(Annot !="NS"))[, "Gene"], 
                               "ODN2006" = (DEG_Stim_Unstim_30h_FilClus$ODN2006%>%
                                              filter(Annot !="NS"))[, "Gene"],
                               "CD3CD28" = (DEG_Stim_Unstim_30h_FilClus$`CD3/CD28_beads`%>%
                                              filter(Annot !="NS"))[, "Gene"]) 
# Generation of list of downregulated genes per 
UpsetData_SIGN_30h_FilClus <- list_to_matrix(UpsetData_SIGN_30h_FilClus)
# Generation of comb_mat object
UpsetData_SIGN_30h_FilClus <- make_comb_mat(UpsetData_SIGN_30h_FilClus, mode = "distinct")
# Generation of the plot
UpSet(UpsetData_SIGN_30h_FilClus, comb_order = order(comb_size(UpsetData_SIGN_30h_FilClus), decreasing = T),
      comb_col = "#595B5C", bg_pt_col = "white",
      column_title = "Overlap significant DE genes",
      right_annotation = rowAnnotation(
                "DE genes" = anno_barplot(set_size(UpsetData_SIGN_30h_FilClus),
                                          border = FALSE,
                                          gp = gpar(fill = "#595B5C",
                                                    lty = "blank"), 
                                          add_numbers = T,
                                          width = unit(4, "cm"))),
      top_annotation = HeatmapAnnotation(
                "Intersection size" = anno_barplot(comb_size(UpsetData_SIGN_30h_FilClus),
                                                   border = FALSE,
                                                   gp = gpar(fill = "#595B5C",
                                                             lty = "blank"), 
                                                   add_numbers = T,
                                                   height = unit(4, "cm")),
                annotation_name_side = "left",
                annotation_name_rot = 90))
```


```{r UpsetPlotVisualzation_OverlapDEGenes_Advanced, fig.height=4, fig.width=6}
### GENERATE DATA FRAMES NEEDED TO PLOT DATA ###
# Filter out the genes of interest in 2 different sets
UpsetData_COMB_30h_FilClus <- c(lapply(DEG_Stim_Unstim_30h_FilClus, FUN = function(stim){(stim%>%
                                                                          filter(Annot == "UP"))[, "Gene"]}),
                       lapply(DEG_Stim_Unstim_30h_FilClus, FUN = function(stim){(stim%>%
                                                                          filter(Annot == "DOWN"))[, "Gene"]}))
names(UpsetData_COMB_30h_FilClus)<- paste(names(UpsetData_COMB_30h_FilClus), rep(c("UP", "DOWN"), each = 4), sep = "_")


# Generation of list of downregulated genes per 
UpsetData_COMB_30h_FilClus <- as.data.frame(list_to_matrix(UpsetData_COMB_30h_FilClus))

# Check if the expression change is consistent between stimulations
UpsetData_COMB_30h_FilClus<-UpsetData_COMB_30h_FilClus%>%
  mutate_at(vars(ends_with("UP")), function(x){ifelse(x==1, "UP", NA)})%>%
  mutate_at(vars(ends_with("DOWN")), function(x){ifelse(x==1, "DOWN", NA)})

# Check if the genes were upregulated/downregulated/mixed over the distinct subsets 
UpsetData_COMB_30h_FilClus$Direction<-apply(UpsetData_COMB_30h_FilClus, 1, function(col) {
  u <- unique(na.omit(col))
  if (length(u) == 1) {
    # either "UP" or "DOWN"
    return(u)
  } else if (length(u) > 1) {
    return("MIXED")
  } else {
    # no non-NA values
    return(NA_character_)
  }
})

# Store the genes linked to specific subsets
# stack function creates a data frame from a list
Subsets<-stack(sapply(X = comb_name(UpsetData_SIGN_30h_FilClus),
                      FUN = function(subset){extract_comb(UpsetData_SIGN_30h_FilClus, subset)}))
colnames(Subsets) <- c("Genes", "Subset")
UpsetData_COMB_30h_FilClus <- left_join(x = UpsetData_COMB_30h_FilClus%>%
                              rownames_to_column(var = "Genes"),
                            y = Subsets,
                            by = "Genes")%>%
  mutate(Direction = factor(Direction, levels = c("MIXED", "UP", "DOWN"),
                               ordered = T))

# Store the table of the distinct directions
UpsetData_COMB_30h_FilClus.table<-table(UpsetData_COMB_30h_FilClus$Subset,
                            UpsetData_COMB_30h_FilClus$Direction)
```

Generation figure of manuscript figure. 

```{r PlotUpset_DEGenes, fig.height=3, fig.width=7}
### PLOTTING THE DATA USING UPSET PLOT OF COMPLEX HEATMAP ###
# Generate upset plot
upPlot<-
  UpSet(UpsetData_SIGN_30h_FilClus, comb_order = order(comb_size(UpsetData_SIGN_30h_FilClus), decreasing = T),
      #  Colors of the dots and balkjes onderaan 
      comb_col = "#595B5C", bg_pt_col = "white",
      # Add plot title
      column_title = "Overlap significant DE genes",
      # Update column annotation (entire set of DEgenes)
      right_annotation = rowAnnotation(
                "DE genes" = anno_barplot(set_size(UpsetData_SIGN_30h_FilClus),
                                          border = FALSE,
                                          gp = gpar(fill = "#595B5C",
                                                    lty = "blank"), 
                                          add_numbers = T,
                                          width = unit(4, "cm"))),
      # Update row annotation on top of the Upset plot
      top_annotation = HeatmapAnnotation(
                                      # Add barplot based on the direction of the DE genes
                "Intersection size" = anno_barplot(UpsetData_COMB_30h_FilClus.table,
                                                   border = FALSE,
                                                   # Store correct colors
                                                   gp = gpar(fill = colDirection[colnames(
                                                     UpsetData_COMB_30h_FilClus.table)],
                                                             lty = "blank"),
                                                   add_numbers = T,
                                                   height = unit(3, "cm")),
                # Annotation name shown at the left side
                annotation_name_side = "left",
                # Rotation of the annotation at the left side
                annotation_name_rot = 90))

# Generate legend
lgd = Legend(
  labels    = colnames(UpsetData_COMB_30h_FilClus.table),
  legend_gp = gpar(fill = colDirection[colnames(UpsetData_COMB_30h_FilClus.table)]),
  title     = "Direction"
)

# Visualise the upset plot
pdf(file.path(params$path_Results, 
              "DEGAnalysis", 
              paste0(params$date, "Upset_SignDEGenes_Stim_vs_Unstim_30h.pdf")), width=7, height=4)
draw(upPlot,
     annotation_legend_list = list(lgd),
     annotation_legend_side = "left")
dev.off()
```

## Hypergeometric testing in pathway enrichment analysis

# Differential gene expression between incubation times

```{r DEGeneAnalysis_IncubationTime}
set.seed(2024)
# Change idents of cells
Idents(IntData.B_8h.FilCus)<-"Stimulation"
Idents(IntData.B_30h.FilCus)<-"Stimulation"

#Initialization of data frame
DEG_Time_Stim_FilClus<-sapply(unique(IntData.B_8h.FilCus$Stimulation),
                              function(x) NULL)

# Perform DEG analysis
for (stim in names(DEG_Time_Stim_FilClus)){
  message(paste0("Working on stimulation ", stim))
  MergedSeurat_Stim<-merge(x = subset(IntData.B_8h.FilCus,
                                             idents = stim),
                                   y = subset(IntData.B_30h.FilCus,
                                              idents = stim),
                           merge.data = TRUE,
                           merge.dr = TRUE)
  MergedSeurat_Stim[["RNA"]] <- JoinLayers(MergedSeurat_Stim[["RNA"]])
  Idents(MergedSeurat_Stim)<-"TimePoint"
  DEG_Time_Stim_FilClus[[stim]]<-FindMarkers(object = MergedSeurat_Stim,
                                       min.pct = 0, logfc.threshold = 0,
                                       # 8 hours of stimulation as reference
                                       ident.1 = "30", ident.2 = 8)%>%
    mutate(Score = (pct.1/(pct.2 + 0.01))*avg_log2FC)%>%
    rownames_to_column(var = "Gene")
}

# Save list of DE genes (all genes are scored here)
saveRDS(DEG_Time_Stim_FilClus, 
        file = file.path(params$path_Results, 
                         "DEGAnalysis",
                         paste(params$date2, "DE_AllGenes_8h_vs_30h_PerStimulation.rds", sep = "_")))
```

```{r DEGeneAnalysis_IncubationTime_ReadIn}
DEG_Time_Stim_FilClus<-readRDS(file = file.path(params$path_Results,
                                                "DEGAnalysis",
                                                 paste(params$date2, "DE_AllGenes_8h_vs_30h_PerStimulation.rds", sep = "_")))
```

```{r DEGeneAnalysis_InucubationTime_SignGenes}
for (stim in names(DEG_Time_Stim_FilClus)){
  # Add additional information to the dataset
  DEG_Time_Stim_FilClus[[stim]]<-DEG_Time_Stim_FilClus[[stim]]%>%
  mutate(Annot=case_when(
    avg_log2FC > 1 & p_val_adj <0.05 ~ "UP",
    avg_log2FC < -1 & p_val_adj <0.05 ~ "DOWN",
    TRUE ~ "NS"))
  # Select genes to be plotted as label on the volcano graph
  top20genes <- head(DEG_Time_Stim_FilClus[[stim]]%>%
                       filter(Annot != 'NS')%>%
                       arrange(p_val_adj, descending = TRUE), 20)[,'Gene']
  DEG_Time_Stim_FilClus[[stim]]<-DEG_Time_Stim_FilClus[[stim]]%>%
    mutate(GeneLabels=ifelse(DEG_Time_Stim_FilClus[[stim]]$Gene %in% top20genes,
                             DEG_Time_Stim_FilClus[[stim]]$Gene,NA))
}
```

## Visualization DEGens between incubation times
 
```{r UpsetPlot_IncuvbationTime_DataPreparation}
# Only focus on significantly different genes
UpsetData_SIGN_Time_FilClus <- list("IFN" = (DEG_Time_Stim_FilClus$IFNgamma%>%
                                           filter(Annot !="NS"))[, "Gene"], 
                               "MegaCD40L" = (DEG_Time_Stim_FilClus$MegaCD40L%>%
                                                filter(Annot !="NS"))[, "Gene"], 
                               "ODN2006" = (DEG_Time_Stim_FilClus$ODN2006%>%
                                              filter(Annot !="NS"))[, "Gene"],
                               "CD3CD28" = (DEG_Time_Stim_FilClus$`CD3/CD28_beads`%>%
                                              filter(Annot !="NS"))[, "Gene"],
                               "Unstimulated" = (DEG_Time_Stim_FilClus$Unstimulated%>%
                                                   filter(Annot != "NS"))[, "Gene"]) 
# Generation of list of significant genes per stimulation
UpsetData_SIGN_Time_FilClus <- list_to_matrix(UpsetData_SIGN_Time_FilClus)
# Generation of comb_mat object
UpsetData_SIGN_Time_FilClus <- make_comb_mat(UpsetData_SIGN_Time_FilClus, mode = "distinct")

### GENERATE DATA FRAMES NEEDED TO PLOT DATA ###
# Filter out the genes of interest in 2 different sets
UpsetData_COMB_Time_FilClus <- c(lapply(DEG_Time_Stim_FilClus, FUN = function(stim){(stim%>%
                                                                          filter(Annot == "UP"))[, "Gene"]}),
                                 lapply(DEG_Time_Stim_FilClus, FUN = function(stim){(stim%>%
                                                                          filter(Annot == "DOWN"))[, "Gene"]}))
names(UpsetData_COMB_Time_FilClus)<- paste(names(UpsetData_COMB_Time_FilClus), rep(c("UP", "DOWN"), each = 5), sep = "_")


# Generation of list of downregulated genes per 
UpsetData_COMB_Time_FilClus <- as.data.frame(list_to_matrix(UpsetData_COMB_Time_FilClus))

# Check if the expression change is consistent between stimulations
UpsetData_COMB_Time_FilClus<-UpsetData_COMB_Time_FilClus%>%
  mutate_at(vars(ends_with("UP")), function(x){ifelse(x==1, "UP", NA)})%>%
  mutate_at(vars(ends_with("DOWN")), function(x){ifelse(x==1, "DOWN", NA)})

# Check if the genes were upregulated/downregulated/mixed over the distinct subsets 
UpsetData_COMB_Time_FilClus$Direction<-apply(UpsetData_COMB_Time_FilClus, 1, function(col) {
  u <- unique(na.omit(col))
  if (length(u) == 1) {
    # either "UP" or "DOWN"
    return(u)
  } else if (length(u) > 1) {
    return("MIXED")
  } else {
    # no non-NA values
    return(NA_character_)
  }
})

# Store the genes linked to specific subsets
# stack function creates a data frame from a list
Subsets<-stack(sapply(X = comb_name(UpsetData_SIGN_Time_FilClus),
                      FUN = function(subset){extract_comb(UpsetData_SIGN_Time_FilClus, subset)}))
colnames(Subsets) <- c("Genes", "Subset")
UpsetData_COMB_Time_FilClus <- left_join(x = UpsetData_COMB_Time_FilClus%>%
                              rownames_to_column(var = "Genes"),
                            y = Subsets,
                            by = "Genes")%>%
  mutate(Direction = factor(Direction, levels = c("MIXED", "UP", "DOWN"),
                               ordered = T))

# Store the table of the distinct directions
UpsetData_COMB_Time_FilClus.table<-table(UpsetData_COMB_Time_FilClus$Subset,
                            UpsetData_COMB_Time_FilClus$Direction)

# Generation of the plot
UpSet(UpsetData_SIGN_Time_FilClus, comb_order = order(comb_size(UpsetData_SIGN_Time_FilClus), decreasing = T),
      comb_col = "#595B5C", bg_pt_col = "white",
      column_title = "Overlap significant DE genes",
      right_annotation = rowAnnotation(
                "DE genes" = anno_barplot(set_size(UpsetData_SIGN_Time_FilClus),
                                          border = FALSE,
                                          gp = gpar(fill = "#595B5C",
                                                    lty = "blank"), 
                                          add_numbers = T,
                                          width = unit(4, "cm"))),
      top_annotation = HeatmapAnnotation(
                "Intersection size" = anno_barplot(comb_size(UpsetData_SIGN_Time_FilClus),
                                                   border = FALSE,
                                                   gp = gpar(fill = "#595B5C",
                                                             lty = "blank"), 
                                                   add_numbers = T,
                                                   height = unit(4, "cm")),
                annotation_name_side = "left",
                annotation_name_rot = 90))
```

```{r UpsetPlot_IncubationTime_PlotGeneration}
### PLOTTING THE DATA USING UPSET PLOT OF COMPLEX HEATMAP ###
# Generate upset plot
upPlot_Time<-
  UpSet(UpsetData_SIGN_Time_FilClus, comb_order = order(comb_size(UpsetData_SIGN_Time_FilClus), decreasing = T),
      #  Colors of the dots and balkjes onderaan 
      comb_col = "#595B5C", bg_pt_col = "white",
      # Add plot title
      column_title = "Overlap significant DE genes",
      # Update column annotation (entire set of DEgenes)
      right_annotation = rowAnnotation(
                "DE genes" = anno_barplot(set_size(UpsetData_SIGN_Time_FilClus),
                                          border = FALSE,
                                          gp = gpar(fill = "#595B5C",
                                                    lty = "blank"), 
                                          add_numbers = T,
                                          width = unit(4, "cm"))),
      # Update row annotation on top of the Upset plot
      top_annotation = HeatmapAnnotation(
                                      # Add barplot based on the direction of the DE genes
                "Intersection size" = anno_barplot(UpsetData_COMB_Time_FilClus.table,
                                                   border = FALSE,
                                                   # Store correct colors
                                                   gp = gpar(fill = colDirection[colnames(
                                                     UpsetData_COMB_Time_FilClus.table)],
                                                             lty = "blank"),
                                                   add_numbers = T,
                                                   height = unit(3, "cm")),
                # Annotation name shown at the left side
                annotation_name_side = "left",
                # Rotation of the annotation at the left side
                annotation_name_rot = 90))

# Generate legend
lgd = Legend(
  labels    = colnames(UpsetData_COMB_Time_FilClus.table),
  legend_gp = gpar(fill = colDirection[colnames(UpsetData_COMB_Time_FilClus.table)]),
  title     = "Direction"
)

# Visualise the upset plot
pdf(file.path(params$path_Results, 
              "DEGAnalysis", 
              paste0(params$date, "Upset_SignDEGenes_Stim_vs_Unstim_8h.pdf")), width=7, height=4)
draw(upPlot_Time,
     annotation_legend_list = list(lgd),
     annotation_legend_side = "left")
dev.off()
```

# Hypergeometric testing in pathway enrichment

Statistical testing of the enrichment of specific gene sets within the unique or overlapping genes will be performed using a hypergeometric statistical testing (HGT) approach. Under the hood, it is actually similar to the one-sided Fisher's exact test based on the contingency table between genes present or absent in the gene set of interest. More information in this [video](https://www.youtube.com/watch?v=bN2OakOqOQI&list=PL1TwkRAGtllLe_bhend9-dzKToOxLuObm).

The approach followed is described [here](https://montilab.github.io/BS831/articles/docs/HyperEnrichment.html) and written down in more detail in my notes (see Remarkable). This part is adapted from the original code written in script `011_DEGenes_Visualisation_GSEAOverlap.Rmd`. The tested subset of significant genes are split into a down- and a upregulated subset to test for pathway down or upregulation, respectively.

## Interesting gene sets
The gene sets of interest were obtained via the MSig database and based on multiple articles such as

- https://www.frontiersin.orghttps://www.frontiersin.org/journals/immunology/articles/10.3389/fimmu.2021.602539/full
- https://www.gsea-msigdb.org/gsea/msigdb/human/collections.jsp#C8

Two gene sets are used that could be of interest for this analysis

- MSigDB, category 5: Gene ontology gene sets (subcategory Biological Processes)
- MSigDB, category 7: Immunologic signature gene sets (not included here yet)

**Note**: Only genes that are present in our dataset can be present in the gene sets which we are using as a reference for the hypergeometric testing. Therefore, genes that are in the subset but not in our database are also deleted from the gene subset.

```{r GeneSetExtraction}
# Gene ontology - Biological processes
GeneSets_GO_BP<-msigdbr(species = "Homo sapiens", category = "C5", subcategory = "BP")%>%
  # Select only the columns of interest for downstream analysis
  select(gs_name, gene_symbol, gs_description)

# Select only the genes that are present in our dataset
GeneSets_GO_BP<-GeneSets_GO_BP%>%
  filter(gene_symbol %in% DEG_Time_Stim_FilClus$IFNgamma[["Gene"]])

dim(GeneSets_GO_BP)
```

## Final testing per subset of genes based on upset plot

In this part, the final gene set enrichment of the distinct DEG subsets using the Gene Ontology Biological Processes is performed using the hypergeometric testing approach.

### Genes related to incubation time per stimulation

A distinguishment is made between upregulated and downregulated genes to obtain pathways that are upregulated and downregulated in specific conditions.
```{r IncubTime_HGT_ExtractGenesOverlap_UpReg}
DEG_IncubTime<-list()

# UPREGULATED GENES
# Unique subsets
DEG_IncubTime[["IFNG_UP"]]<-(UpsetData_COMB_Time_FilClus%>%filter(Subset == "10000", !is.na(IFNgamma_UP)))[,"Genes"]
DEG_IncubTime[["MegaCD40L_UP"]]<-(UpsetData_COMB_Time_FilClus%>%filter(Subset == "01000", !is.na(MegaCD40L_UP)))[,"Genes"]
DEG_IncubTime[["ODN2006_UP"]]<-(UpsetData_COMB_Time_FilClus%>%filter(Subset == "00100", !is.na(ODN2006_UP)))[,"Genes"]
DEG_IncubTime[["CD3CD28_UP"]]<-(UpsetData_COMB_Time_FilClus%>%filter(Subset == "00010", !is.na(`CD3/CD28_beads_UP`)))[,"Genes"]
DEG_IncubTime[["Unstimulated_UP"]]<-(UpsetData_COMB_Time_FilClus%>%filter(Subset == "00001", !is.na(`Unstimulated_UP`)))[,"Genes"]

# Overlapping subsets of interest (Only the upregulated genes are taken into account)
DEG_IncubTime[["IFNG_ODN_CD3CD8_UP"]]<-(UpsetData_COMB_Time_FilClus%>%filter(Subset == "10110", Direction == "UP"))[,"Genes"]
DEG_IncubTime[["ALL_UP"]]<-(UpsetData_COMB_Time_FilClus%>%filter(Subset == "11110", Direction == "UP"))[,"Genes"]
DEG_IncubTime[["ALL_Unstim_UP"]]<-(UpsetData_COMB_Time_FilClus%>%filter(Subset == "11111", Direction == "UP"))[,"Genes"]

# DOWNREGULATED GENES
# Unique subsets
DEG_IncubTime[["IFNG_DOWN"]]<-(UpsetData_COMB_Time_FilClus%>%filter(Subset == "10000", !is.na(IFNgamma_DOWN)))[,"Genes"]
DEG_IncubTime[["MegaCD40L_DOWN"]]<-(UpsetData_COMB_Time_FilClus%>%filter(Subset == "01000", !is.na(MegaCD40L_DOWN)))[,"Genes"]
DEG_IncubTime[["ODN2006_DOWN"]]<-(UpsetData_COMB_Time_FilClus%>%filter(Subset == "00100", !is.na(ODN2006_DOWN)))[,"Genes"]
DEG_IncubTime[["CD3CD28_DOWN"]]<-(UpsetData_COMB_Time_FilClus%>%filter(Subset == "00010", !is.na(`CD3/CD28_beads_DOWN`)))[,"Genes"]
DEG_IncubTime[["Unstimulated_DOWN"]]<-(UpsetData_COMB_Time_FilClus%>%filter(Subset == "00001", !is.na(`Unstimulated_DOWN`)))[,"Genes"]
# Overlapping subsets of interest (Only the upregulated genes are taken into account)
DEG_IncubTime[["IFNG_ODN_CD3CD8_DOWN"]]<-(UpsetData_COMB_Time_FilClus%>%filter(Subset == "10110", Direction == "DOWN"))[,"Genes"]
DEG_IncubTime[["ALL_DOWN"]]<-(UpsetData_COMB_Time_FilClus%>%filter(Subset == "11110", Direction == "DOWN"))[,"Genes"]
DEG_IncubTime[["ALL_Unstim_DOWN"]]<-(UpsetData_COMB_Time_FilClus%>%filter(Subset == "11111", Direction == "DOWN"))[,"Genes"]
```

```{r IncubTime_Phyper_HGT_GOSets}
# Save the total number of genes in the entire dataset
totalGeneNumber<-length(DEG_Time_Stim_FilClus$IFNgamma$Gene)

# Initialization of dataframe
GSE_HGT_GO_BP_IncubTime <- as.data.frame(matrix(nrow=length(unique(GeneSets_GO_BP$gs_name)),ncol=1))
GSE_HGT_GO_BP_IncubTime$GeneSet <-unique(GeneSets_GO_BP$gs_name)
GSE_HGT_GO_BP_IncubTime <- GSE_HGT_GO_BP_IncubTime%>%column_to_rownames(var = "GeneSet")


# For-loop over the pathways of interest
for (pathway in GSE_HGT_GO_BP_IncubTime%>%rownames()){
  # Extract the genes of the pathway of interest
  pathway.genes<-(GeneSets_GO_BP%>%filter(gs_name == pathway))[['gene_symbol']]
  
  # For-loop over the distinct subsets
  for(subset.name in names(DEG_IncubTime)){
           # Initialization of names
           overlap.var<-paste("Overlap",subset.name, sep = "_")
           overlap.pval<-paste("pval", subset.name, sep = "_")
           # Extract subset of interest
           subset<-DEG_IncubTime[[subset.name]]
           # Check overlap + store in the final dataframe
           overlap<-length(intersect(pathway.genes, subset))
           GSE_HGT_GO_BP_IncubTime[pathway, overlap.var]<-overlap
           
           # Perform hypergeometric testing
           GSE_HGT_GO_BP_IncubTime[pathway, overlap.pval]<-phyper(q = overlap - 1,
                                                       m = length(pathway.genes),
                                                       n = totalGeneNumber-length(pathway.genes),
                                                       k = length(subset),
                                                       lower.tail = F)
           }
}

# Remove the unnecessary columns
GSE_HGT_GO_BP_IncubTime[["V1"]]<-NULL
```

```{r IncubTime_Phyper_HGT_MultipleTestingCorrection_GOSets}
# Multiple testing correction with Benjamini Hochberg
for(col in grep(pattern = "^pval", x = names(GSE_HGT_GO_BP_IncubTime), value = T)){
  new_colname <- paste0('adj', col)
  GSE_HGT_GO_BP_IncubTime[,new_colname] <- p.adjust(GSE_HGT_GO_BP_IncubTime[,col], method = "BH")
}

GSE_HGT_GO_BP_IncubTime%>%head()
```

```{r IncubTime_Phyper_HGT_Vis_Pval_Overview, fig.width=8, fig.height=7}
print("GO gene sets")
sapply(X = grep(pattern = "adjpval", x = names(GSE_HGT_GO_BP_IncubTime), value = T),
       FUN = function(subset){
         sum(GSE_HGT_GO_BP_IncubTime[[subset]]<=0.05)
       })
```

Visualization of the interesting enriched gene sets per stimulation (combination). 

Bar plot of the adjusted p-values for the gene sets tested against the DE genes observed in all stimulatory conditions.

```{r IncubTime_Phyper_HGT_Vis_Pval_Barplots, fig.width=15, fig.height=4}
for (col in grep(pattern = "adjpval", x = names(GSE_HGT_GO_BP_IncubTime), value = T)){
  plotData<-GSE_HGT_GO_BP_IncubTime%>%
                 rownames_to_column(var = "GeneSet")%>%
         select(GeneSet, col)%>%
         filter(.data[[col]] <= 0.05)%>%
         arrange(col)%>%
         mutate(GeneSet=gsub(pattern = "GOBP_", replacement = "", x = GeneSet))

  if(dim(plotData)[1] != 0){
    plot<-ggplot(data = plotData%>%head(20),
       mapping = aes(x = -log10(.data[[col]]), y = reorder(GeneSet, -.data[[col]]), 
                     label = format(.data[[col]], scientific = T, digits = 3)))+
    geom_col()+
    xlim(c(0,NA))+
    themeG+
    theme(panel.grid.major = element_line(color = "grey95",
                                    size = 0.5,
                                    linetype = 1),
          panel.grid.minor = element_line(color = "grey95",
                                          size = 0.5,
                                          linetype = 1),
          axis.ticks = element_blank(),
          axis.line = element_blank())+
    geom_text(x = 0.05,hjust = 0, color = "white", size = 3)+
    labs(y = "GO Gene sets - Biological processes",
         x = "-log10(adj. p-value)",
         title = paste0("Enriched pathways based on subset ", gsub(pattern = "adjpval_", replacement = "", x = col)))
   show(plot)
  }
}
rm(plotData, plot)
```


### Top 20 DEGenes over the distinct approaches

```{r Top20Genes}
for (stim in names(DEG_Time_Stim_FilClus)){
  top20genes <- head(DEG_Time_Stim_FilClus[[stim]]%>%
                       filter(Annot != 'NS')%>%
                         # The absolute value of the score is used in this case to also take into account possible downregulated genes
                       arrange(abs(Score), descending = TRUE), 20)[,'Gene']
  DEG_Time_Stim_FilClus[[stim]]<-DEG_Time_Stim_FilClus[[stim]]%>%
    mutate(Top20=ifelse(DEG_Time_Stim_FilClus[[stim]]$Gene %in% top20genes,
                        TRUE,FALSE))
}
rm(top20genes)

# Overlap in top20 genes
Top20_Time_Stim <- list("IFNgamma" = (DEG_Time_Stim_FilClus$IFNgamma%>%
                                           filter(Top20))[, "Gene"], 
                       "MegaCD40L" = (DEG_Time_Stim_FilClus$MegaCD40L%>%
                                        filter(Top20))[, "Gene"], 
                       "ODN2006" = (DEG_Time_Stim_FilClus$ODN2006%>%
                                      filter(Top20))[, "Gene"],
                       "CD3/CD28" = (DEG_Time_Stim_FilClus$`CD3/CD28_beads`%>%
                                      filter(Top20))[, "Gene"])
```


```{r Top20SharedGenes_VennDiagram, fig.height=5, fig.width=15}
# Extract the top 20 DEgenes, taking into account the direction of gene expression
for (stim in names(DEG_Time_Stim_FilClus)){
  top10Genes_UP <- head(DEG_Time_Stim_FilClus[[stim]]%>%
                          # Only extract the top 10 of the upregulated genes
                       filter(Annot == 'UP')%>%
                       arrange(-Score), 10)[,'Gene']
  top10Genes_DOWN <- head(DEG_Time_Stim_FilClus[[stim]]%>%
                       filter(Annot == 'DOWN')%>%
                         # Only extract the top 10 of the downregulated genes
                       arrange(Score), 10)[,'Gene']
  DEG_Time_Stim_FilClus[[stim]]<-DEG_Time_Stim_FilClus[[stim]]%>%
    # Correctly annotate the genes
    mutate(Top10=ifelse(DEG_Time_Stim_FilClus[[stim]]$Gene %in% top10Genes_UP,
                        "UP",ifelse(DEG_Time_Stim_FilClus[[stim]]$Gene %in% top10Genes_DOWN, "DOWN", NA)))
  
  # Store only the top 10 down/upregulated genes in a dataframe
  # First check if the dataframe is does exist or is empyt
  if (!exists("DEG_Time_Stim_FilClus_Top10.df") || (dim(DEG_Time_Stim_FilClus_Top10.df)[1]) == 0){
    message("Creating dataframe")
    DEG_Time_Stim_FilClus_Top10.df <- DEG_Time_Stim_FilClus[[stim]] |>
      filter(!is.na(Top10)) |>
      # Chance the column name using the `stim` variable
      mutate(!!stim := Top10) |>
      select(Gene, !!stim)
  } else {
    message("\nJoining additional genes")
    # Using full_join to add the other genes that were found in other stimulations
    DEG_Time_Stim_FilClus_Top10.df <- full_join(x = DEG_Time_Stim_FilClus_Top10.df,
                                                y = DEG_Time_Stim_FilClus[[stim]] |>
                                                  filter(!is.na(Top10)) |>
                                                  # Chance the column name using the `stim` variable
                                                  mutate(!!stim := Top10) |>
                                                  select(Gene, !!stim),
                                                by = "Gene")
  }
}

# Store as matrix as input for complex heatmap
DEG_Time_Stim_FilClus_Top10.mat<-as.matrix(DEG_Time_Stim_FilClus_Top10.df|>
                                             # Add rownames
                                             column_to_rownames(var = "Gene")|>
                                             # Change order of the columns
                                             select(IFNgamma, ODN2006, MegaCD40L, `CD3/CD28_beads`,Unstimulated)|>
                                             # Sort based on the conditions and gene expression direction
                                             arrange(desc(IFNgamma), desc(ODN2006), desc(MegaCD40L), desc(`CD3/CD28_beads`),
                                                     desc(Unstimulated)))
                                             
# Generate heatmap of top 20 genes
Heatmap(t(DEG_Time_Stim_FilClus_Top10.mat), 
        name = "Expression direction", 
        col = colDirection,
        column_title = "Top20Genes per condition - 8h vs 30h",
        row_names_side = "left", column_names_side = "top",
        na_col = "grey95")
```

#### NCBI information scraping

To link information to the gene of interest based on the NCBI database, a data scraping step is performed using the `rentrez` R package. The following links contain necessary examples, vignettes and explanation on how this is properly performed

- https://bioconnector.github.io/workshops/r-ncbi.html

```{r FunctionExtractData_Entrez}
for (gene in rownames(DEG_Time_Stim_FilClus_Top10.mat)){
  # Get the gene ids based on the NCBI database
  gene_ids<-entrez_search(db = "gene", term  = paste0(gene, "[Gene] AND Homo Sapiens[Organism]"))
  # Fetch the necessary data based on the extracted gene_ids
  for (ids in gene_ids$ids){
    # Stored as docsum
    gene_info.docsum<-entrez_fetch(db = "gene", id = ids, rettype = "docsum")
    # Get the data in an interesting format
    gene_info.df<-data.frame(OGInfo=strsplit(x = gene_info.docsum, split = '\n')|>
      unlist())|>
      filter(grepl(OGInfo, pattern = paste(c("<Name>", "<Description>", "<Summary>"), collapse = "|")))|>
      # Remove the tabs
      mutate(OGInfo = gsub(pattern = "\t", replacement = "", x = OGInfo))|>
      mutate(Gene = gene,
             Key = lapply(X = OGInfo, FUN = function(x){strsplit(x = x, split = "<|>|</")[[1]][2]})%>%
               unlist(),
             Value = lapply(X = OGInfo, FUN = function(x){strsplit(x = x, split = "<|>|</")[[1]][3]})%>%
               unlist())|>
      dplyr::select(-OGInfo)|>
      pivot_wider(names_from = Key,
                  values_from = Value)
    
    # Concatenate all the information on the genes of interest
    if (gene_info.df$Summary=="") {
      # If no additional information is added, the data is not concatenated to the dataframe
      break
      } else if (!exists("DEG_Time_Stim_FilClus_Top10.info") || (dim(DEG_Time_Stim_FilClus_Top10.info)[1]) == 0){
      message("Creating dataframe")
      DEG_Time_Stim_FilClus_Top10.info <- gene_info.df
      } else {
      message("\nJoining dataframes")
      # Using full_join to add the other genes that were found in other stimulations
      DEG_Time_Stim_FilClus_Top10.info <- rbind(DEG_Time_Stim_FilClus_Top10.info,
                                                y = gene_info.df)
    }
  }
}

# Saving the final dataset
write.xlsx(DEG_Time_Stim_FilClus_Top10.info, file.path(params$path_Results, "DEGAnalysis", paste0(params$date2, 
                                                        "_Info_Top20DEGenes_IncubTime_8vs30.xlsx")))
```

#### Dotplot average expression of the top 20 genes

```{r IncubTime_Heatmap_Top20Genes, fig.width=15, fig.height=5}
# Grouping based on Stimulation
Idents(IntData.B_8h.FilCus)<-"Stimulation"
Idents(IntData.B_30h.FilCus)<-"Stimulation"

DotPlot_MultipleFeat(object = IntData.B_8h.FilCus, metaData = "Stimulation",
                     numSorting = F, features = rownames(DEG_Time_Stim_FilClus_Top10.mat))+
  ggtitle("8 hours of stimulation")

DotPlot_MultipleFeat(object = IntData.B_30h.FilCus, metaData = "Stimulation",
                     numSorting = F, features = rownames(DEG_Time_Stim_FilClus_Top10.mat))+
  ggtitle("30 hours of stimulation")
```


## Final testing per stimulation

```{r IncubTime_Phyper_HGT_GOSets_PerStim}
# Save the total number of genes in the entire dataset
totalGeneNumber<-length(DEG_Time_Stim_FilClus$IFNgamma$Gene)

# Initialization of dataframe
GSE_HGT_GO_BP_IncubTime.Stim <- as.data.frame(matrix(nrow=length(unique(GeneSets_GO_BP$gs_name)),ncol=1))
GSE_HGT_GO_BP_IncubTime.Stim$GeneSet <-unique(GeneSets_GO_BP$gs_name)
GSE_HGT_GO_BP_IncubTime.Stim <- GSE_HGT_GO_BP_IncubTime.Stim%>%column_to_rownames(var = "GeneSet")


# For-loop over the pathways of interest
for (pathway in GSE_HGT_GO_BP_IncubTime.Stim%>%rownames()){
  # Extract the genes of the pathway of interest
  pathway.genes<-(GeneSets_GO_BP%>%filter(gs_name == pathway))[['gene_symbol']]
  
  # For-loop over the distinct subsets
  for(stim in names(DEG_Time_Stim_FilClus)){
           # Initialization of names
           overlapUP.var<-paste("Overlap",stim, "UP", sep = "_")
           overlapUP.pval<-paste("pval", stim, "UP", sep = "_")
           overlapDOWN.var<-paste("Overlap",stim, "DOWN", sep = "_")
           overlapDOWN.pval<-paste("pval", stim, "DOWN", sep = "_")
           # Extract subset of interest
           subsetUP<-(DEG_Time_Stim_FilClus[[stim]]%>%
             filter(Annot == "UP"))[["Gene"]]
           subsetDOWN<-(DEG_Time_Stim_FilClus[[stim]]%>%
             filter(Annot == "DOWN"))[["Gene"]]
           # Check overlap + store in the final dataframe
           overlapUP<-length(intersect(pathway.genes, subsetUP))
           GSE_HGT_GO_BP_IncubTime.Stim[pathway, overlapUP.var]<-overlapUP
           overlapDOWN<-length(intersect(pathway.genes, subsetDOWN))
           GSE_HGT_GO_BP_IncubTime.Stim[pathway, overlapDOWN.var]<-overlapDOWN
           
           # Perform hypergeometric testing for upregulated genes
           GSE_HGT_GO_BP_IncubTime.Stim[pathway, overlapUP.pval]<-phyper(q = overlapUP - 1,
                                                       m = length(pathway.genes),
                                                       n = totalGeneNumber-length(pathway.genes),
                                                       k = length(subsetUP),
                                                       lower.tail = F)
           # Perform hypergeometric testing for downregulated genes
           GSE_HGT_GO_BP_IncubTime.Stim[pathway, overlapDOWN.pval]<-phyper(q = overlapDOWN - 1,
                                                       m = length(pathway.genes),
                                                       n = totalGeneNumber-length(pathway.genes),
                                                       k = length(subsetDOWN),
                                                       lower.tail = F)
           }
}

# Remove the unnecessary columns
GSE_HGT_GO_BP_IncubTime.Stim[["V1"]]<-NULL
```

```{r IncubTime_Phyper_HGT_MultipleTestingCorrection_GOSets_PerStim}
# Multiple testing correction with Benjamini Hochberg
for(col in grep(pattern = "^pval", x = names(GSE_HGT_GO_BP_IncubTime.Stim), value = T)){
  new_colname <- paste0('adj', col)
  GSE_HGT_GO_BP_IncubTime.Stim[,new_colname] <- p.adjust(GSE_HGT_GO_BP_IncubTime.Stim[,col], method = "BH")
}

GSE_HGT_GO_BP_IncubTime.Stim%>%head()
```

```{r IncubTime_Phyper_HGT_Vis_Pval_Overview, fig.width=8, fig.height=7}
print("GO gene sets")
sapply(X = grep(pattern = "adjpval", x = names(GSE_HGT_GO_BP_IncubTime.Stim), value = T),
       FUN = function(subset){
         sum(GSE_HGT_GO_BP_IncubTime.Stim[[subset]]<=0.05)
       })
```

Visualization of the interesting enriched gene sets per stimulation (combination). 

Bar plot of the adjusted p-values for the gene sets tested against the DE genes observed in all stimulatory conditions.

```{r IncubTime_Phyper_HGT_Vis_Pval_Barplots, fig.width=15, fig.height=4}
for (col in grep(pattern = "adjpval", x = names(GSE_HGT_GO_BP_IncubTime.Stim), value = T)){
  plotData<-GSE_HGT_GO_BP_IncubTime.Stim%>%
                 rownames_to_column(var = "GeneSet")%>%
         select(GeneSet, col)%>%
         filter(.data[[col]] <= 0.05)%>%
         arrange(col)%>%
         mutate(GeneSet=gsub(pattern = "GOBP_", replacement = "", x = GeneSet))

  if(dim(plotData)[1] != 0){
    plot<-ggplot(data = plotData%>%head(20),
       mapping = aes(x = -log10(.data[[col]]), y = reorder(GeneSet, -.data[[col]]), 
                     label = format(.data[[col]], scientific = T, digits = 3)))+
    geom_col()+
    xlim(c(0,NA))+
    themeG+
    theme(panel.grid.major = element_line(color = "grey95",
                                    size = 0.5,
                                    linetype = 1),
          panel.grid.minor = element_line(color = "grey95",
                                          size = 0.5,
                                          linetype = 1),
          axis.ticks = element_blank(),
          axis.line = element_blank())+
    geom_text(x = 0.05,hjust = 0, color = "white", size = 3)+
    labs(y = "GO Gene sets - Biological processes",
         x = "-log10(adj. p-value)",
         title = paste0("Enriched pathways based on subset ", gsub(pattern = "adjpval_", replacement = "", x = col)))
   show(plot)
  }
}
rm(plotData, plot)
```


```{r IncubTime_GSEA_PerStim_Heatmap, fig.height=20, fig.width=15}
# Extracting the most interesting gene sets based on the top 20 per subset
for (stim_dir in grep(pattern = "adjpval", x = names(GSE_HGT_GO_BP_IncubTime.Stim), value = T)){
  top20GSE.Stim <- head(GSE_HGT_GO_BP_IncubTime.Stim%>%
                          # Only extract the top 10 of the upregulated genes
                       arrange(.data[[stim_dir]]), n = 10)%>%rownames
  
  # Store only the top 10 down/upregulated genes in a dataframe
  # First check if the dataframe is does exist or is empyt
  if (!exists("top20GSE.total") || length(top20GSE.total) == 0){
    message("Creating vector")
    top20GSE.total <- top20GSE.Stim
  } else {
    message("\nJoining additional gene sets")
    # Adding additional genes if they are not yet present in the top20GSE list
   top20GSE.total <- unique(c(top20GSE.total, top20GSE.Stim))
  }
}

# Extracting plot matrix necessary for heatmap generation
plotData <- (GSE_HGT_GO_BP_IncubTime.Stim[top20GSE.total,]%>%
  select(starts_with("adjpval"))%>%
  as.matrix()%>%
  log10())*(-1)
rownames(plotData)<-gsub(x = rownames(plotData), pattern = "GOBP_", replacement = "")

# Plot the heatmap
Heatmap(matrix = plotData,
        row_names_side = "left", column_names_side = "top",
        show_row_dend = F, show_column_dend = F, 
        row_dend_reorder = T, column_dend_reorder = T,
        row_names_gp = gpar(fontsize = 7),
        row_names_rot = -45)
```
