---
title: "004_MultiNicheNet_AdaptedApproach_CorrelationBasedFiltering"
author: "Tine D'hamers"
date: "2026-04-07"
output: html_document
params: 
  prefix_output: "PatStim_NoBase"
  path_input: "/kyukon/scratch/gent/vo/002/gvo00224/TOBI/Projects/EProfile/Results/Exp_202211_HCSamples_CEV017-22/004_MultiNicheNetR/Objects"
  date: "20260407"
  path_output_circos: "/kyukon/scratch/gent/vo/002/gvo00224/TOBI/Projects/EProfile/Results/Exp_202211_HCSamples_CEV017-22/004_MultiNicheNetR/CircosDiagrams"
  path_output: "/kyukon/scratch/gent/vo/002/gvo00224/TOBI/Projects/EProfile/Results/Exp_202211_HCSamples_CEV017-22/004_MultiNicheNetR/"
---

# Rationale

This file will be used to perform the downstream data analysis based on the MultiNicheNet objects generated in the `001_MultiNicheNetObjectInitialization.Rmd` script. The link to an overview of the distinct functions used in this package is availabe [here](https://rdrr.io/github/saeyslab/multinichenetr/api/). The starting part (preprocessing, ...) is similar to what was performed in the `002_MultiNicheNet_DownstreamAnalysis.Rmd` file. However, updates in the analysis were performed based on the inputs of Ruth Seurinck during a meeting at 03/04/2026. Meeting notes can be found in my Drylab Notebook on scCITE-seq. The object initialization using the adapted approach are performed in the `003_MultiNicheNetObjectInitialization_LowerDEScenario.Rmd` script

The adaptation included:

- Switching the prioritization scenario from 'regular' to 'lower DE', in which the ligand actvitiy gets a highter weight and all the expression based features a lower weight in prioritization interactions
- Prioritization based on correlations between ligand activity and up/down-regulation of the target genes in the receiver cell type


# Start-up

## Load libraries

```{r Libraries}
.libPaths("/kyukon/data/gent/vo/002/gvo00224/TOBI/Projects/SingleCellPipeline/resultsCeline/seuratintegrate/Rpackages/R_4.4.1-foss_2023b/")
library(dplyr)
library(tidyverse)
options(bitmapType='cairo')
library(nichenetr)
library(multinichenetr)
library(SingleCellExperiment)

# Plots
library(ggplot2)
library(ggrepel)
library(ggpubr)
library(ggridges)

library(ComplexHeatmap)
```

## Plotting Themes

```{r PlotThemes}
# Data visualization themes
source("/kyukon/data/gent/vo/002/gvo00224/TOBI/Projects/EProfile/Scripts/PlotThemes_sc_CVID.R")
# Store colors for the distinct directions
colDirection<-c("UP" = "darkgreen",
                "DOWN" = "darkred",
                "MIXED" = "#D5BBB1",
                "NS" = "grey80")
grid.grey = "grey95"
```

## Preparatory steps

```{r Network and matrix loading}
# Specify the organism
organism <- "human"

# Time out option
options(timeout = 120)

# Extract the necessary data from Zenodo
# Start with the ligand-receptor network
lr_network_all <- readRDS(url( "https://zenodo.org/record/10229222/files/lr_network_human_allInfo_30112033.rds" )) %>%
  # Converting aliases used in the reference to gene symbols, matching our data
  mutate( ligand = convert_alias_to_symbols(ligand, organism = organism), 
          receptor = convert_alias_to_symbols(receptor, organism = organism)) 

lr_network_all <- lr_network_all %>% 
  # Convert the nqmes in ligand names into valid names 
  # Ths implies e.g. "-" are transformed in a "."
  mutate(ligand = make.names(ligand), receptor = make.names(receptor)) 

lr_network <- lr_network_all %>% 
  # Only keep unique/distinct rows from a data frame to avoid the loss of genes in downstream visualizations
  # Taking into account the combination of ligand and receptor
  distinct(ligand, receptor) 

# Ligand-target matrix (LARGE MATRIX!!)
ligand_target_matrix <- readRDS(url( "https://zenodo.org/record/7074291/files/ligand_target_matrix_nsga2r_final.rds" )) 

# Convert colnames (ligands) into usable names
colnames(ligand_target_matrix) <- colnames(ligand_target_matrix) %>% 
  convert_alias_to_symbols(organism = organism) %>% 
  make.names() 

# Convert rownames (targets) into usable names
rownames(ligand_target_matrix) <- rownames(ligand_target_matrix) %>% 
  convert_alias_to_symbols(organism = organism) %>% 
  make.names() 

# Only use the ligands that are present in the ligand-target matrix for downstream analysis
lr_network <- lr_network %>% 
  filter(ligand %in% colnames(ligand_target_matrix)) 
# Only use the ligands that are present in the ligand-receptor matrix for downstream analysis
ligand_target_matrix <- ligand_target_matrix[, lr_network$ligand %>% unique()]

# Clean a bit
rm(ligand_target_matrix)
```

```{r Contrasts}
# Contrast definition
# contrast_oi <- paste0("'", sort(unique(IntData_NoDbl.sce$Stimulation))[-5], "-Unstimulated", "'", collapse =",")
contrast_oi <- c("'CD3CD28_beads-Unstimulated','IFNgamma-Unstimulated','MegaCD40L-Unstimulated','ODN2006-Unstimulated'")

contrast_tbl <-tibble(contrast =
                        c("CD3CD28_beads-Unstimulated","IFNgamma-Unstimulated",
                          "MegaCD40L-Unstimulated","ODN2006-Unstimulated"),
                      group = c("CD3CD28_beads", "IFNgamma", "MegaCD40L",  "ODN2006"))

contrast_oi
contrast_tbl

# Top N targets per ligand to be shown
top_n_target <- 250
```

## Reading in the necessary data

```{r ReadInRDSObject}
multinichenet_output_8h<-readRDS(file = file.path(params$path_output,
                                                  "Objects",
                                                  paste0("MultiNicheNet_Output_8h_LowerDEScenario_", params$date, ".rds")))
multinichenet_output_30h<-readRDS(file = file.path(params$path_output,
                                                  "Objects",
                                                  paste0("MultiNicheNet_Output_30h_LowerDEScenario_", params$date, ".rds")))
```

# MultiNicheNet downstream data analysis

## 1. Visualization of differential cell-cell interactions

### 1.1. Summarizing ChordDiagram circos plots

Just to get a general overview on the trends throughout the distinct conditions, I generated the circos plots to get a general idea of top priorizited interactions in the overall network of interactions. I will look into the top 50 interactions for now.

```{r ChordDiagrams}
# Extraction of the top interactons overall
## 8h
prioritized_tbl_oi_all_top50.8h <- get_top_n_lr_pairs(
  multinichenet_output_8h$prioritization_tables,
  top_n=50,
  # Top is given overall
  rank_per_group=FALSE
)
table(prioritized_tbl_oi_all_top50.8h$group)

## 30h
prioritized_tbl_oi_all_top50.30h <- get_top_n_lr_pairs(
  multinichenet_output_30h$prioritization_tables,
  top_n=50,
  # Top is given overall
  rank_per_group=FALSE
)
table(prioritized_tbl_oi_all_top50.30h$group)

# Extract the data
prioritized_tbl_oi_all_top50.8h.data<-
  multinichenet_output_8h$prioritization_tables$group_prioritization_tbl%>%
  # Filter out the prioritized top 50 interactions
  filter(id %in% prioritized_tbl_oi_all_top50.8h$id) %>%
  distinct(id, sender, receiver, ligand, receptor, group)%>%
  left_join(prioritized_tbl_oi_all_top50.8h)
prioritized_tbl_oi_all_top50.30h.data<-
  multinichenet_output_30h$prioritization_tables$group_prioritization_tbl%>%
  # Filter out the prioritized top 50 interactions
  filter(id %in% prioritized_tbl_oi_all_top50.30h$id) %>%
  distinct(id, sender, receiver, ligand, receptor, group)%>%
  left_join(prioritized_tbl_oi_all_top50.30h)

# The interactions that are not present (as NA in the table) needs to be set to zero
prioritized_tbl_oi_all_top50.8h.data$prioritization_score[is.na(prioritized_tbl_oi_all_top50.8h.data$prioritization_score)]<-0
prioritized_tbl_oi_all_top50.30h.data$prioritization_score[is.na(prioritized_tbl_oi_all_top50.30h.data$prioritization_score)]<-0

# Extract the sender and receiver cell types
# These are the same for both incubation time points
senders_receivers<-union(prioritized_tbl_oi_all_top50.8h.data$sender %>% unique(),
                         prioritized_tbl_oi_all_top50.8h.data$receiver %>% unique())%>%
  sort()

# Use the same colors as used in previous analysis
colorsCellType<-colorsCellType %>% 
  magrittr::set_names(make.names(names(colorsCellType)))
colors_sender<- colorsCellType %>%
  magrittr::extract(names(colorsCellType) %in% senders_receivers)
colors_receiver<- colorsCellType %>%
  magrittr::extract(names(colorsCellType) %in% senders_receivers)

# Generate the circos diagram
Circos_top50.8h<- 
  make_circos_group_comparison(prioritized_tbl_oi_all_top50.8h.data,
                             colors_sender = colors_sender,
                             colors_receiver = colors_receiver)
Circos_top50.30h<-
  make_circos_group_comparison(prioritized_tbl_oi_all_top50.30h.data,
                             colors_sender = colors_sender,
                             colors_receiver = colors_receiver)
```

```{r SaveCircosPlots_Top50AllInteractions}
# Save plots
# 8h
for (condition in names(Circos_top50.8h)){
  pdf(file = file.path(params$path_output_circos,
                           paste0("Circos_", condition,
                                  "_Top50Interqctions_AllInteractions_8h_", params$date2, ".pdf")),
    width =7, height=8)
  print(Circos_top50.8h[[condition]])
  dev.off()
}

# 30h
for (condition in names(Circos_top50.30h)){
  pdf(file = file.path(params$path_output_circos,
                           paste0("Circos_", condition,
                                  "_Top50Interqctions_AllInteractions_30h_", params$date2, ".pdf")),
    width =7, height=8)
  print(Circos_top50.30h[[condition]])
  dev.off()
}
```

### 1.2. Bubble plot visualization of top 50 interactions per group (stimulatory condition)

```{r BubblePlot_Top50_PerCondition}
BubblePlot_top50_PerCond.8h<-lapply(
  multinichenet_output_8h$prioritization_tables$group_prioritization_table_source$group%>%
    unique,
  FUN = function(group_oi){
    tmp_prior_tbl <- get_top_n_lr_pairs(
      multinichenet_output_8h$prioritization_tables,
      top_n = 50,
      groups_oi = group_oi)
    plot_oi<-make_sample_lr_prod_activity_plots_Omnipath(
      multinichenet_output_8h$prioritization_tables,
      tmp_prior_tbl %>% 
        inner_join(lr_network_all)
    )
    plot_oi + labs(title = group_oi)
  })
BubblePlot_top50_PerCond.30h<-lapply(
  multinichenet_output_30h$prioritization_tables$group_prioritization_table_source$group%>%
    unique,
  FUN = function(group_oi){
    tmp_prior_tbl <- get_top_n_lr_pairs(
      multinichenet_output_30h$prioritization_tables,
      top_n = 50,
      groups_oi = group_oi)
    plot_oi<-make_sample_lr_prod_activity_plots_Omnipath(
      multinichenet_output_30h$prioritization_tables,
      tmp_prior_tbl %>% 
        inner_join(lr_network_all)
    )
    plot_oi + labs(title = group_oi)
  })

names(BubblePlot_top50_PerCond.8h)<-unique(multinichenet_output_8h$prioritization_tables$group_prioritization_table_source$group)
names(BubblePlot_top50_PerCond.30h)<-unique(multinichenet_output_30h$prioritization_tables$group_prioritization_table_source$group)
```

```{r SaveBubblePlots_Top50_PerCondition}
# Save plots
# 8h
for (condition in names(BubblePlot_top50_PerCond.8h)){
  pdf(file = file.path(params$path_output,
                           paste0("BubblePlot_", condition,
                                  "_Top50Interactions_8h_", params$date, ".pdf")),
    width = 17, height=15)
  print(BubblePlot_top50_PerCond.8h[[condition]])
  dev.off()
}

# 30h
for (condition in names(BubblePlot_top50_PerCond.30h)){
  pdf(file = file.path(params$path_output,
                           paste0("BubblePlot_", condition,
                                  "_Top50Interactions_30h_", params$date, ".pdf")),
    width = 17, height=15)
  print(BubblePlot_top50_PerCond.30h[[condition]])
  dev.off()
}
```

## 2. Selection of LR-interactions based on intercellular regulatory network

### 2.1. Filtering of target genes based on LR-target expression correlation 

The ligand-receptor interactions are filtered out based on correlation value. Downstream target genes linked to the LR-interactions of interaction that both show high expression correlation (pearson correlation > 0.33) and have some prior knowledge to support their link are retained to filter out more accurate LR-interactions.

```{r CorFiltering_LRInteractions_8h}
# Define threshold for correlation filtering
CorThresh<-0.80

lr_target_prior_cor_filtered.8h<-
  # Extract all groups
  multinichenet_output_8h$prioritization_tables$group_prioritization_tbl$group %>% 
  unique() %>%
  # Perform the analysis for each of the groups separately
  lapply(function(group_oi){
    lr_target_prior_cor_filtered.8h <- inner_join(
      multinichenet_output_8h$lr_target_prior_cor,
      # Extract the ligand activities of the interactions
      multinichenet_output_8h$ligand_activities_targets_DEgenes$ligand_activities%>%
        # Make sure the combinations made by the folowing features all have a unique line
        # The ligand activity scores are not taken along in this line
        distinct(ligand, target, direction_regulation, contrast)
      ) %>%
      # Adding group information next to the contrast information
      inner_join(contrast_tbl)%>%
      # Only select the group of interest
      filter(group == group_oi)
    
    # Extract upregulated targets meeting the cutoff for the pearson correlation
    lr_target_prior_cor_filtered_up.8h<-lr_target_prior_cor_filtered.8h%>%
      filter(direction_regulation == "up") %>%
      filter((rank_of_target < top_n_target) & (pearson > CorThresh))
    
    # Extract the downregulated targets meeting the cutoff for the pearson correlation
    lr_target_prior_cor_filtered_down.8h<-lr_target_prior_cor_filtered.8h%>%
      filter(direction_regulation == "down") %>%
      filter((rank_of_target < top_n_target) & (pearson < -CorThresh))
    
    # Only retain the selection of up- and downregulated genes
    lr_target_prior_cor_filtered.8h<-bind_rows(
      lr_target_prior_cor_filtered_up.8h,
      lr_target_prior_cor_filtered_down.8h
    )
    }) %>% 
  # Place them in one data frame
  bind_rows()

lr_target_df.8h <- lr_target_prior_cor_filtered.8h %>%
  distinct(group, sender, receiver, ligand, receptor, id, target, direction_regulation)

lr_target_df.8h%>%head(n=20)
```

```{r CorFiltering_LRInteractions_30h}
lr_target_prior_cor_filtered.30h<-
  # Extract all groups
  multinichenet_output_30h$prioritization_tables$group_prioritization_tbl$group %>% 
  unique() %>%
  # Perform the analysis for each of the groups separately
  lapply(function(group_oi){
    lr_target_prior_cor_filtered.30h <- inner_join(
      multinichenet_output_30h$lr_target_prior_cor,
      # Extract the ligand activities of the interactions
      multinichenet_output_30h$ligand_activities_targets_DEgenes$ligand_activities%>%
        # Make sure the combinations made by the folowing features all have a unique line
        # The ligand activity scores are not taken along in this line
        distinct(ligand, target, direction_regulation, contrast)
      ) %>%
      # Adding group information next to the contrast information
      inner_join(contrast_tbl)%>%
      # Only select the group of interest
      filter(group == group_oi)
    
    # Extract upregulated targets meeting the cutoff for the pearson correlation
    lr_target_prior_cor_filtered_up.30h<-lr_target_prior_cor_filtered.30h%>%
      filter(direction_regulation == "up") %>%
      filter((rank_of_target < top_n_target) & (pearson > CorThresh))
    
    # Extract the downregulated targets meeting the cutoff for the pearson correlation
    lr_target_prior_cor_filtered_down.30h<-lr_target_prior_cor_filtered.30h%>%
      filter(direction_regulation == "down") %>%
      filter((rank_of_target < top_n_target) & (pearson < -CorThresh))
    
    # Only retain the selection of up- and downregulated genes
    lr_target_prior_cor_filtered.30h<-bind_rows(
      lr_target_prior_cor_filtered_up.30h,
      lr_target_prior_cor_filtered_down.30h
    )
    }) %>% 
  # Place them in one data frame
  bind_rows()

lr_target_df.30h <- lr_target_prior_cor_filtered.30h %>%
  distinct(group, sender, receiver, ligand, receptor, id, target, direction_regulation)

lr_target_df.30h%>%head(n=20)
```

```{r Network_CorFiltering}
# Infer the interaction network including the DE target genes
network.8h <- infer_intercellular_regulatory_network(
  lr_target_df.8h, prioritized_tbl_oi_all_top50.8h)
network.30h <- infer_intercellular_regulatory_network(
  lr_target_df.30h, prioritized_tbl_oi_all_top50.30h)

# Check the links that are present in the network
network.8h$links %>%
  head(n=20)
sprintf("Number of links prioritized in 8h of stimulation: %i", dim(network.8h$links)[1])
network.30h$links %>%
  head(n=20)
sprintf("Number of links prioritized in 30h of stimulation: %i", dim(network.30h$links)[1])

# Check the nodes that are present in the network
network.8h$nodes %>%
  head(n=20)
sprintf("Number of nodes prioritized in 8h of stimulation: %i", dim(network.8h$nodes)[1])
network.30h$nodes %>%
  head(n=20)
sprintf("Number of nodes prioritized in 30h of stimulation: %i", dim(network.30h$nodes)[1])
``` 


```{r PlotNetwork, fig.height=10, fig.width=30}
# Generate the plot for the network visualization
network_graph.8h <- visualize_network(network.8h, colors_sender)
network_graph.30h <- visualize_network(network.30h, colors_sender)

# Show graph plots
network_graph.8h$plot + labs(title = "8h")
network_graph.30h$plot + labs(title = "30h")
``` 

### 2.2. Selection based on top interactions and correlation cut-off

```{r FilterTop50_CorPrior_Interactions}
# Extract the top 50 interactions that are also present in the prioritization network based on the pearson correlation filtering
prioritized_tbl_oi_network.8h<- prioritized_tbl_oi_all_top50.8h %>%
  inner_join(network.8h$prioritized_lr_interactions)
prioritized_tbl_oi_network.30h<- prioritized_tbl_oi_all_top50.30h %>%
  inner_join(network.30h$prioritized_lr_interactions)

# Retained interactions
sprintf("Final number of filtered interactions - 8h")
table(prioritized_tbl_oi_network.8h$group)
print("Final number of filtered interactions - 30h")
table(prioritized_tbl_oi_network.30h$group)
``` 

It might seem interesting to increase the number of retained top interactions, however this is not the limiting factor in this inner_join. The number of final interactions is the same as what was observed as unique interactions in specific conditions focusing on the network (including target genes as well). 

```{r BubblePlot_FilterInteractions, fig.height = 10, fig.width = 10}
BubblePlot_prior_network.8h<-lapply(unique(network.8h$prioritized_lr_interactions$group),
       function(group_oi){
         tmp_prior_tbl <- prioritized_tbl_oi_network.8h%>%
           filter(group == group_oi)
         
         plot_oi <- make_sample_lr_prod_activity_plots_Omnipath(
           multinichenet_output_8h$prioritization_tables,
           tmp_prior_tbl %>% inner_join(lr_network_all)
         )
         plot_oi + 
           labs(title = group_oi)
       })
BubblePlot_prior_network.30h<-lapply(unique(network.30h$prioritized_lr_interactions$group),
       function(group_oi){
         tmp_prior_tbl <- prioritized_tbl_oi_network.30h%>%
           filter(group == group_oi)
         
         plot_oi <- make_sample_lr_prod_activity_plots_Omnipath(
           multinichenet_output_30h$prioritization_tables,
           tmp_prior_tbl %>% inner_join(lr_network_all)
         )
         plot_oi + 
           labs(title = group_oi)
       })
```

```{r BubblePlot_FilteredInteractions_OverAllStim, fig.height=10, fig.width=10}
# 8h
BubblePlot_prior_network.8h.Merged<-make_sample_lr_prod_activity_plots_Omnipath(
  multinichenet_output_8h$prioritization_tables,
  prioritized_tbl_oi_network.8h %>% 
    inner_join(lr_network_all))+
  labs(title = "8h")

pdf(file = file.path(params$path_output,
                           paste0("BubblePlot_NetworkBasedPrior_8h_", params$date, ".pdf")),
    width =17, height=10)
BubblePlot_prior_network.8h.Merged
dev.off()

# 30h
BubblePlot_prior_network.30h.Merged<-make_sample_lr_prod_activity_plots_Omnipath(
  multinichenet_output_30h$prioritization_tables,
  prioritized_tbl_oi_network.30h %>% 
    inner_join(lr_network_all))+
  labs(title = "30h")

pdf(file = file.path(params$path_output,
                           paste0("BubblePlot_NetworkBasedPrior_30h_", params$date, ".pdf")),
    width =17, height=10)
BubblePlot_prior_network.30h.Merged
dev.off()
```


## 3. Selection of LR-interactions based on intercellular regulatory network for each stimulatory condition separately

### 3.1. Filtering of target genes based on LR-target expression correlation 

The ligand-receptor interactions are filtered out based on correlation value. Downstream target genes linked to the LR-interactions of interaction that both show high expression correlation (pearson correlation > 0.33) and have some prior knowledge to support their link are retained to filter out more accurate LR-interactions.

This step is similar to what was performed in paragraph 2.1, but now we only focus on the top 50 interactions of each of the stimulatory conditions separately, to get a prioritization per condition.

```{r CorFiltering_LRInteractions_8h_PerCond}
# Top n interactions that need to be prioritized
Sel_top<-50
# Extract names of the distinct groups of interest
groups<-multinichenet_output_8h$prioritization_tables$group_prioritization_table_source$group%>%
    unique
## 8h
# Selection of the top 50 interactions per stimulatory condition
prioritized_tbl_oi_PerCond_top50.8h<- lapply(
  groups,
  FUN = function(group_oi){
    get_top_n_lr_pairs(
      multinichenet_output_8h$prioritization_tables,
      top_n = Sel_top,
      groups_oi = group_oi)})
# Make sure the naming of the list is correct
names(prioritized_tbl_oi_PerCond_top50.8h)<-groups

## 30h
# Selection of the top 50 interactions per stimulatory condition
prioritized_tbl_oi_PerCond_top50.30h<- lapply(
  groups,
  FUN = function(group_oi){
    get_top_n_lr_pairs(
      multinichenet_output_30h$prioritization_tables,
      top_n = Sel_top,
      groups_oi = group_oi)})
# Make sure the naming of the list is correct
names(prioritized_tbl_oi_PerCond_top50.30h)<-groups
```

In the following step, instead of using the overall top 50, the top 50 of each of the conditions is used to check for interesting interactions.

```{r Network_CorFiltering_PriorPerCond}
# Infer the interaction network including the DE target genes
network_PerCond.8h <- lapply(X = groups, FUN = function(group_oi){
  infer_intercellular_regulatory_network(
  lr_target_df.8h, prioritized_tbl_oi_PerCond_top50.8h[[group_oi]])})
names(network_PerCond.8h)<-groups
network_PerCond.30h <- lapply(X = groups, FUN = function(group_oi){
  infer_intercellular_regulatory_network(
  lr_target_df.30h, prioritized_tbl_oi_PerCond_top50.30h[[group_oi]])})
names(network_PerCond.30h)<-groups

# Check the links that are present in the network
network_PerCond.8h$IFNgamma$links %>%
  head(n=20)
sprintf("Number of links prioritized in 8h of stimulation: %i", dim(network_PerCond.8h$IFNgamma$links)[1])
network_PerCond.30h$IFNgamma$links %>%
  head(n=20)
sprintf("Number of links prioritized in 30h of stimulation: %i", dim(network_PerCond.30h$IFNgamma$links)[1])

# Check the nodes that are present in the network
network_PerCond.8h$IFNgamma$nodes %>%
  head(n=20)
sprintf("Number of nodes prioritized in 8h of stimulation: %i", dim(network_PerCond.8h$IFNgamma$nodes)[1])
network_PerCond.30h$IFNgamma$nodes %>%
  head(n=20)
sprintf("Number of nodes prioritized in 30h of stimulation: %i", dim(network_PerCond.30h$IFNgamma$nodes)[1])
``` 


### 3.2. Selection based on top interactions and correlation cut-off

```{r FilterTop50_CorPrior_Interactions_PriorPerCond}
# Extract the top 50 interactions that are also present in the prioritization network based on the pearson correlation filtering
prioritized_tbl_oi_PerCond_network.8h<-lapply(X=groups,FUN=function(groups_oi){
  prioritized_tbl_oi_PerCond_top50.8h[[groups_oi]] %>%
  inner_join(network_PerCond.8h[[groups_oi]]$prioritized_lr_interactions)
})
names(prioritized_tbl_oi_PerCond_network.8h)<-groups

prioritized_tbl_oi_PerCond_network.30h<-lapply(X=groups,FUN=function(groups_oi){
  prioritized_tbl_oi_PerCond_top50.30h[[groups_oi]] %>%
  inner_join(network_PerCond.30h[[groups_oi]]$prioritized_lr_interactions)
})
names(prioritized_tbl_oi_PerCond_network.30h)<-groups

# Retained interactions
sprintf("Final number of filtered interactions - 8h")
sapply(X=groups, FUN=function(group_oi){
  table(prioritized_tbl_oi_PerCond_network.8h[[group_oi]]$group)})
print("Final number of filtered interactions - 30h")
sapply(X=groups, FUN=function(group_oi){
  table(prioritized_tbl_oi_PerCond_network.30h[[group_oi]]$group)})
``` 

### 3.3. Visualization of bubble plot for prioritized interactions
```{r BubblePlot_FilterInteractions_PriorPerCond, fig.height = 10, fig.width = 20}
# Generate bubble plot only of prioritized interactions
BubblePlot_prior_PerCond_network.8h<-lapply(X=groups, FUN=function(group_oi){
         tmp_prior_tbl <- prioritized_tbl_oi_PerCond_network.8h[[group_oi]]
         plot_oi <- make_sample_lr_prod_activity_plots_Omnipath(
           multinichenet_output_8h$prioritization_tables,
           tmp_prior_tbl %>% inner_join(lr_network_all)
         )
         plot_oi + 
           labs(title = group_oi)
       })
BubblePlot_prior_PerCond_network.30h<-lapply(X=groups, FUN=function(group_oi){
         tmp_prior_tbl <- prioritized_tbl_oi_PerCond_network.30h[[group_oi]]
         plot_oi <- make_sample_lr_prod_activity_plots_Omnipath(
           multinichenet_output_30h$prioritization_tables,
           tmp_prior_tbl %>% inner_join(lr_network_all)
         )
         plot_oi + 
           labs(title = group_oi)
       })

# Rename lists
names(BubblePlot_prior_PerCond_network.8h)<-groups
names(BubblePlot_prior_PerCond_network.30h)<-groups

# Visualize
BubblePlot_prior_PerCond_network.8h
BubblePlot_prior_PerCond_network.30h
```


### 3.4. Comparison with prioritized interactions

```{r Comparison_PriorInt}
# Extrqct the interaction IDs, including the LR interaction AND interacting cell types
prioritized_tbl_oi_PerCond_network.8h.idList<-lapply(X=groups, FUN=function(group_oi){
  prioritized_tbl_oi_PerCond_network.8h[[group_oi]][['id']]
})
names(prioritized_tbl_oi_PerCond_network.8h.idList)<-paste('8h', groups, sep = "_")

prioritized_tbl_oi_PerCond_network.30h.idList<-lapply(X=groups, FUN=function(group_oi){
  prioritized_tbl_oi_PerCond_network.30h[[group_oi]][['id']]
})
names(prioritized_tbl_oi_PerCond_network.30h.idList)<-paste('30h', groups, sep = "_")

# Join the lists
prioritized_tbl_oi_PerCond_network.idList<-c(prioritized_tbl_oi_PerCond_network.8h.idList,
                                             prioritized_tbl_oi_PerCond_network.30h.idList)

# Generation of combinations
prioritized_tbl_oi_PerCond_network.combMat<-make_comb_mat(prioritized_tbl_oi_PerCond_network.idList)

# Show upsetplot
UpSet(prioritized_tbl_oi_PerCond_network.combMat,
      top_annotation = HeatmapAnnotation(
        "Intersection size" = anno_barplot(comb_size(prioritized_tbl_oi_PerCond_network.combMat),
                                           border = FALSE,
                                           gp = gpar(fill = "#595B5C",
                                                     lty = "blank"), 
                                           add_numbers = TRUE,
                                           height = unit(4, "cm")),
        annotation_name_side = "left",
        annotation_name_rot = 90),
      right_annotation = rowAnnotation(
        "Set size" = anno_barplot(set_size(prioritized_tbl_oi_PerCond_network.combMat),
                                  border = FALSE,
                                  gp = gpar(fill = "#595B5C",
                                            lty = "blank"), 
                                  add_numbers = T,
                                  width = unit(4, "cm")))
      )
```

### 3.5. Final bubble plot generation
```{r Fig4_FinalBubblePlotGeneration,fig.height=10, fig.width=15}
# Merge all the dataframes with the prioritized interactions in one data frame
tmp_AllPrior.8h<-do.call("rbind", prioritized_tbl_oi_PerCond_network.8h)
tmp_AllPrior.30h<-do.call("rbind", prioritized_tbl_oi_PerCond_network.30h)

# Save plots
# 8h
for (condition in names(BubblePlot_top50_PerCond.8h)){
  pdf(file = file.path(params$path_output,
                           paste0("BubblePlot_", condition,
                                  "_Top50Interactions_8h_", params$date, ".pdf")),
    width = 17, height=15)
  print(BubblePlot_top50_PerCond.8h[[condition]])
  dev.off()
}

# Visualize in a BubblePlot
# 8h
s
# 30h
pdf(file = file.path(params$path_output,
                     paste0("BubblePlot_PriorInt_Top50PerStim_30h_", params$date, ".pdf")),
    width=17, height=12)
make_sample_lr_prod_activity_plots_Omnipath(
           multinichenet_output_30h$prioritization_tables,
           tmp_AllPrior.30h %>% inner_join(lr_network_all))
dev.off()
``` 
